### 2019-07-03 README.Rmd
commit

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# ggmatplot

<!-- badges: start -->
<!-- badges: end -->

The goal of ggmatplot is to plot the columns of one matrix against the columns of another.

## Installation

You can install the released version of ggmatplot by the following commands:

```r
library(devtools)
install_github("xuan-liang/ggmatplot")
```

## Example

This is a basic example which shows you how to use this package:

```{r example}
library(ggmatplot)

iris_sub <- subset(iris, Species == "setosa")

# Original way by matplot function
matplot(iris_sub[,c(1,3)], iris_sub[,c(2,4)])

# Use default legend and axis
ggmatplot(iris_sub[,c(1,3)], iris_sub[,c(2,4)])

# Modify legend label and axis
ggmatplot(iris_sub[,c(1,3)], iris_sub[,c(2,4)], shape = c("s","S"), legend_label =  c("Sepal","Petal"), legend_title = "", xlab = "Length", ylab="Width")
```

### 2019-07-03 R/ggmatplot-package.R

@@ -1,0 +1,7 @@
+#' ggmatplot
+#'
+#' This package was created as a way to make a matplot with ggplot version
+#'
+#' @docType ggmatplot
+#' @name ggmatplot
+NULL

### 2019-07-03 R/ggmatplot.R

@@ -1,0 +1,267 @@
+#' Quick matplot function
+#'
+#' `ggmatplot` is a quick way to make a simplier and similar plot as what `matplot` does but with a ggplot version.
+#'
+#' @param x,y vectors or matrices of data for plotting. The number of rows should match. If one of them are missing, the other is taken as y and an x vector of 1:n is used. Missing values (NAs) are allowed.
+#' @param color vector of colors and they are used cyclically.
+#' @param shape vector of shapes and they are used cyclically.
+#' @param linetype vector of linetypes colors  and they are used cyclically.
+#' @param xlim,ylim ranges of x and y axes, as in [plot()].
+#' @param log Which variables to log transform ("x", "y", or "xy")
+#' s@param main,xlab,ylab Character vector giving plot title,
+#'   x axis label, and y axis label respectively.
+#' @param legend_label Character vector giving legend tables for different groups.
+#' @param legend_title Character giving legend title
+#' @param geom Character vector specifying geom(s) to draw. Defaults to
+#'  "point". Other options are "line" or c("point","line").
+#' @param asp The y/x aspect ratio
+#'
+#' @return
+#' @export
+#'
+#' @examples
+#' # Define a data set
+#' iris_sub <- subset(iris, Species == "setosa")
+#' # Use default legend and axis
+#' ggmatplot(iris_sub[,c(1,3)], iris_sub[,c(2,4)])
+#' # Modify legend label and axis
+#' ggmatplot(iris_sub[,c(1,3)], iris_sub[,c(2,4)], shape = c("s","S"), legend_label =  c("Sepal","Petal"), legend_title = "", xlab = "Length", ylab="Width")
+
+ggmatplot <- function (x, y, color = NULL, shape = NULL, linetype = NULL,
+                       xlim = c(NA, NA), ylim = c(NA, NA), log = "",
+                       main = NULL, xlab = NULL, ylab = NULL,
+                       legend_label =  NULL, legend_title = NULL,
+                       geom = "point", asp = NA){
+
+  caller_env <- parent.frame()
+
+  if (!is.character(geom)) stop("`geom` must be a character vector", call. = FALSE)
+
+  if (!missing(x)){
+    ncolx <- ncol(as.data.frame(x))
+  }
+
+  if (!missing(y)){
+    ncoly <- ncol(as.data.frame(y))
+  }
+
+  if (missing(y) & ncolx == 1) {
+    x <- data.frame(x)
+    data <- mutate(x, Observation_number = 1:n())
[…]

### 2019-07-05 README.Rmd
commit

@@
 ggmatplot(iris_sub[,c(1,3)], iris_sub[,c(2,4)], shape = c("s","S"), legend_label =  c("Sepal","Petal"), legend_title = "", xlab = "Length", ylab="Width")
+
+# Linear regression
+x <- rnorm(100,sd=2)
+y <- x*0.5+rnorm(100,sd=1)
+fit.y <- fitted(lm(y~x))
+ggmatplot(x,cbind(y,fit.y))
 ```

### 2019-07-05 R/ggmatplot.R

@@ -39,10 +39,14 @@ ggmatplot <- function (x, y, color = NULL, shape = NULL, linetype = NULL,
 
   if (!missing(x)){
     ncolx <- ncol(as.data.frame(x))
+  } else{
+    ncolx=1
   }
 
   if (!missing(y)){
     ncoly <- ncol(as.data.frame(y))
+  }else{
+    ncoly=1
   }
 
   if (missing(y) & ncolx == 1) {
@@ -61,7 +65,7 @@ ggmatplot <- function (x, y, color = NULL, shape = NULL, linetype = NULL,
     data <- rename(data, new_namey = old_namey)
     p <- qplot(x = Observation_number, y = new_namey, data = data, geom = geom) + ylab(old_namey) + xlab("Observation Number")
   } else if (missing(x) & ncoly > 1) {
-    data <- mutate(y, Observation_number = 1:n()) %>% gather(key = "Column", value = "Value", -(ncolx+1))
+    data <- mutate(y, Observation_number = 1:n()) %>% gather(key = "Column", value = "Value", -(ncoly+1))
     p <- qplot(x = Observation_number, y = Value, data = data, color = Column, shape = Column, linetype = Column, geom = geom)  + xlab("Observation Number")
   } else if (ncolx == 1 & ncoly == 1) {
     data <- data.frame(x,y)

### 2019-07-05 README.Rmd
commit

@@
 library(ggmatplot)
+## Example 1:
+X <- 1:10
+Y=X^2
+Z=X^2-2*X
+# Original way by matplot function
+matplot(X,cbind(Y,Z),pch=c(16,1),xlab="x",ylab="X^2 and X^2-2*X")
+legend(1,100,legend=c("y","Z"),pch=c(16,1)) 
+# Use ggmatplot
+ggmatplot(X,cbind(Y,Z), shape = c(16,1),legend_label =  c("X^2","X^2-2*X"),legend_title = "")
 
+## Example 2: Iris data
 iris_sub <- subset(iris, Species == "setosa")
@@
 
-# Linear regression
+# Example 3: Linear regression
 x <- rnorm(100,sd=2)

### 2019-07-05 README.Rmd
commit

--- 2019-07-03 README.Rmd
@@
 
-## Example
+Below are some basic examples which show you how to use this package.
 
-This is a basic example which shows you how to use this package:
+### Example 1
 
-```{r example}
+```{r load_package, echo=FALSE, message=FALSE, warning=FALSE}
 library(ggmatplot)
+```
 
+```{r example1}
+X <- 1:10
+Y=X^2
+Z=X^2-2*X
+
+# Original way by matplot function
+matplot(X,cbind(Y,Z),pch=c(16,1),xlab="x",ylab="X^2 and X^2-2*X")
+legend(1,100,legend=c("y","Z"),pch=c(16,1))
+
+# Use ggmatplot
+ggmatplot(X,cbind(Y,Z), shape = c(16,1),legend_label =  c("X^2","X^2-2*X"),legend_title = "")
+```
+
+#### Example 2: Iris data
+
+```{r example2}
 iris_sub <- subset(iris, Species == "setosa")
@@
 ```
+
+### Example 3: Linear regression
+
+```{r example3}
+x <- rnorm(100,sd=2)
+y <- x*0.5+rnorm(100,sd=1)
+fit.y <- fitted(lm(y~x))
+ggmatplot(x,cbind(y,fit.y))
+```

### 2019-07-05 README.Rmd
commit

@@
 # Use ggmatplot
-ggmatplot(X,cbind(Y,Z), shape = c(16,1),legend_label =  c("X^2","X^2-2*X"),legend_title = "")
+ggmatplot(X,cbind(Y,Z), shape = c(16,1),legend_label =  c("X^2","X^2-2*X"),legend_title = "",geom=c("point","line"))
 ```
@@
 # Modify legend label and axis
-ggmatplot(iris_sub[,c(1,3)], iris_sub[,c(2,4)], shape = c("s","S"), legend_label =  c("Sepal","Petal"), legend_title = "", xlab = "Length", ylab="Width")
+ggmatplot(iris_sub[,c(1,3)], iris_sub[,c(2,4)], shape = c("S","P"), legend_label =  c("Sepal","Petal"), legend_title = "", xlab = "Length", ylab="Width")
 ```

### 2019-11-06 R/ggmatplot.R
commit

@@ -8,7 +8,7 @@
 #' @param linetype vector of linetypes colors  and they are used cyclically.
 #' @param xlim,ylim ranges of x and y axes, as in [plot()].
 #' @param log Which variables to log transform ("x", "y", or "xy")
-#' s@param main,xlab,ylab Character vector giving plot title,
+#' @param main,xlab,ylab Character vector giving plot title,
 #'   x axis label, and y axis label respectively.
 #' @param legend_label Character vector giving legend tables for different groups.
 #' @param legend_title Character giving legend title
@@ -87,169 +87,50 @@ ggmatplot <- function (x, y, color = NULL, shape = NULL, linetype = NULL,
     p <- qplot(x = Value, y = new_namey, data = data, color = Column, shape = Column, linetype = Column, geom = geom) + ylab(old_namey)
   } else if (ncolx > 1 & ncoly > 1 & ncolx == ncoly) {
     colnames(x) = colnames(y) = paste0("Column ", 1:ncolx)
-    x <- gather(x, key = "Column", value = "x") %>% mutate(Observation_number = 1:n())
-    y <- gather(y, key = "Column", value = "y") %>% mutate(Observation_number = 1:n())
-    data <- full_join(x, y, by = "Observation_number") %>% rename(Group = Column.x)
-    p <- qplot(x = x, y = y, data = data, color = Group, shape = Group, linetype = Group, geom = geom)
+    x <- gather(x, key = "Group", value = "x") %>% mutate(Observation_number = 1:n())
+    y <- gather(y, key = "Group", value = "y") %>% mutate(Observation_number = 1:n())
+    data <- full_join(x, y, by = "Observation_number") %>% rename(Column = Group.x)
+    p <- qplot(x = x, y = y, data = data, color = Column, shape = Column, linetype = Column, geom = geom)
   } else{
     stop("`x`` and `y` must have only 1 or the same number of columns", call. = FALSE)
   }
 
-  if (!is.null(legend_title) & !is.null(legend_label)){
-
-    if (!is.null(color) & !is.null(shape) & !is.null(linetype))
-      p <- p + scale_color_manual(name = legend_title, labels = legend_label, values = color) +
-        scale_shape_manual(name = legend_title, labels = legend_label, values = shape) +
-        scale_linetype_manual(name = legend_title, labels = legend_label, values = linetype)
-
-    if (!is.null(color) & !is.null(shape) & is.null(linetype))
-      p <- p + scale_color_manual(name = legend_title, labels = legend_label, values = color) +
[…]

### 2019-11-06 README.Rmd
commit

--- 2019-07-05 README.Rmd
@@
 # Use ggmatplot
-ggmatplot(X,cbind(Y,Z), shape = c(16,1),legend_label =  c("X^2","X^2-2*X"),legend_title = "")
+ggmatplot(X,cbind(Y,Z), shape = c(16,1),legend_label =  c("X^2","X^2-2*X"),geom=c("point","line"))
 ```
@@
 # Modify legend label and axis
-ggmatplot(iris_sub[,c(1,3)], iris_sub[,c(2,4)], shape = c("s","S"), legend_label =  c("Sepal","Petal"), legend_title = "", xlab = "Length", ylab="Width")
+ggmatplot(iris_sub[,c(1,3)], iris_sub[,c(2,4)], color = 1 ,shape = c("S","P"), xlab = "Length", ylab="Width")
 ```

### 2019-11-14 R/ggmatplot-package.R
commit

@@ -1,7 +1,7 @@
 #' ggmatplot
 #'
-#' This package was created as a way to make a matplot with ggplot version
+#' This package was created as a way to make a matplot with ggplot version.
 #'
-#' @docType ggmatplot
+#' @docType package
 #' @name ggmatplot
 NULL

### 2019-11-14 R/ggmatplot.R

@@ -15,14 +15,11 @@
 #' @param geom Character vector specifying geom(s) to draw. Defaults to
 #'  "point". Other options are "line" or c("point","line").
 #' @param asp The y/x aspect ratio
-#'
-#' @return
 #' @export
 #'
 #' @examples
 #' # Define a data set
 #' iris_sub <- subset(iris, Species == "setosa")
-#' # Use default legend and axis
 #' ggmatplot(iris_sub[,c(1,3)], iris_sub[,c(2,4)])
 #' # Modify legend label and axis
 #' ggmatplot(iris_sub[,c(1,3)], iris_sub[,c(2,4)], shape = c("s","S"), legend_label =  c("Sepal","Petal"), legend_title = "", xlab = "Length", ylab="Width")

### 2021-08-06 based on ggpubr/examples of ggmatplot.Rmd
latest ggmatplot functions depending on package ggpubr

---
title: "Examples of ggmatplot"
author: "Xuan Liang"
output: 
    html_document:
        toc: true
        toc_float: true
        theme: journal
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = TRUE)
```

```{r setup data, message=FALSE}
library(tidyverse)
library(ggpubr)

source("ggmatlist.R") # The function is to convert a list to a matrix
source("ggmatclean.R") # The function is to transfrom a wide format data to a long format data
source("ggmatplot.R") # Plot function without ... option as an argument
source("getoption.R") # Used in ggmatplot1() function to extract what arguments can be allowed in a function of ggpubr, for example, I only tried it for plot.type "histogram". 
source("ggmatplot1.R") # Plot1 function with ... option as an argument. It seems here are some issues ralated with ... option.

# Default plot type: Scatterplot
x <- rnorm(100,sd=2)
y <- x*0.5+rnorm(100,sd=1)
fit.y <- fitted(lm(y~x))
p <- ggmatplot(x,cbind(y,fit.y)) 
p
class(p)

# Plot.type = "line" or "both"
X <- 1:10
Y <- X^2
Z <- X^2-2*X

ggmatplot(X,cbind(Y,Z), shape = c(16,1),legend_label =  c("X^2","X^2-2*X"),plot.type = "line")
ggmatplot(X,cbind(Y,Z), shape = c(16,1),legend_label =  c("X^2","X^2-2*X"),plot.type = "both")

# Plot.type = "density" or "histogram" or "boxplot"
iris_sub <- subset(iris, Species == "setosa")

ggmatplot(iris_sub[1:4], plot.type = "density")

ggmatplot(iris_sub[1:4], plot.type = "histogram")

ggmatplot(iris_sub[1:4], plot.type = "boxplot")

ggmatplot(iris_sub[1:4], plot.type = "dotplot")

ggmatplot(iris_sub[1:2], plot.type = "errorplot")

ggmatplot(iris_sub[1:4], plot.type = "violin")

ggmatplot(iris_sub[1:4], plot.type = "ecdf", xlab="Length", ylab="ecdf")

# Use a list as input
x <- list(low = rnorm(30, -5, 2),
          medium = rnorm(20),
          high = rnorm(40, 3, 0.5))

ggmatplot(x, plot.type = "both", legend_title = "Level")

ggmatplot(x, plot.type = "density", legend_title = "Level")

# Test if the ... option in ggmatplot1 works well
# Correct additional argument
ggmatplot1(x, plot.type = "density", legend_title = "Level", alpha=0.3, rug=TRUE)

# Wrong additional argument but no errors or warnings show up
ggmatplot1(x, plot.type = "density", legend_title = "Level", alpha=0.3, rugg=TRUE)
ggmatplot1(x, plot.type = "density", legend_title = "Level", alpha=0.3, bins=20)
ggmatplot1(x, plot.type = "density", legend_title = "Level", alpha=0.3, binwdith=20)

ggmatplot(x, plot.type = "histogram", legend_title = "Level")
ggmatplot1(x, plot.type = "histogram",legend_title = "Level", bins=20)
ggmatplot1(x, plot.type = "histogram",legend_title = "Level", bins=20, rug=TRUE, add="mean",alpha=0.3)

ggmatplot(x, plot.type = "boxplot", legend_title = "Level")

# Data
stocks <- data.frame(
  X = rnorm(10, 0, 1),
  Y = rnorm(10, 0, 2),
  Z = rnorm(10, 0, 4)
)
stocklong <- stocks %>% gather(stock, price)

# some nice features in ggpubr
p1 <- ggboxplot(stocklong, x = "stock", y = "price",add = "jitter",color="stock", shape = "stock")
p1

my_comparisons <- list( c("X", "Y"), c("Y", "Z"), c("X", "Z") )
p1 + stat_compare_means(comparisons = my_comparisons)+ # Add pairwise comparisons p-value
  stat_compare_means(label.y = 12)

# test the features in ggmatplot
ggmatplot(stocks, plot.type = "boxplot", legend_title = "Level")+ 
  stat_compare_means(comparisons = my_comparisons)+ # Add pairwise comparisons p-value
  stat_compare_means(label.y = 12)

```

### 2021-08-06 R/gg_gather.R

@@ -1,0 +1,3 @@
+gg_gather <- function (x,...){
+  UseMethod(x)
+}

### 2021-08-06 based on ggpubr/getoption.R

@@ -1,0 +1,15 @@
+getoption <- function(params, plot.type){
+  
+  option <- list()
+  
+  if(plot.type=="histogram")
+    allowed_options <- c("data","x","color","fill","linetype","alpha", "bins", "binwidth", "add", "add.params", "rug","add_density","position")
+  
+  for (key in names(params)) {
+    value <- params[[key]]
+    if (key %in% allowed_options) {
+      option[[key]] <- value
+    }
+  }
+  return(option)
+}

### 2021-08-06 based on ggpubr/ggmatclean.R

@@ -1,0 +1,50 @@
+ggmatclean <- function (x, y){
+
+  if(is.list(x)){
+    x <- ggmatlist(x)
+  }
+
+  x <- data.frame(x)
+  ncolx <- ncol(x)
+  
+  if (!missing(y)){
+    ncoly <- ncol(as.data.frame(y))
+    y <- data.frame(y)
+  }else{
+    ncoly<-1
+  }
+  
+   if (missing(y)){
+     data <- mutate(x, Observation_number = 1:n()) %>% gather(key = "Group", value = "yy", -(ncolx+1))
+     xname <- "Observation_number"
+     yname <- "yy"
+     group <- "Group"
+   }else if(ncolx>ncoly&ncoly==1){
+     data <- data.frame(x,y)
+     ncol <- ncol(data)
+     yname <- colnames(data)[ncol]
+     data <- mutate(data, Observation_number = 1:n()) %>% gather(key = "Group", value = "x", -c(ncol,(ncol+1)))
+     xname <- "x"
+     yname <- colnames(y)
+     group <- "Group"
+   }else if(ncoly>ncolx&ncolx==1){
+     data <- data.frame(x,y)
+     ncol <- ncol(data)
+     data <- mutate(data, Observation_number = 1:n()) %>% gather(key = "Group", value = "yy", -c(1,(ncol+1)))
+     xname <- colnames(x)
+     yname <- "yy"
+     group <- "Group"
+   }else if(ncolx==ncoly){
+     colnames(x) = colnames(y) = paste0("Column ", 1:ncolx)
+     x <- gather(x, key = "Group", value = "x") %>% mutate(Observation_number = 1:n())
+     y <- gather(y, key = "Group", value = "yy") %>% mutate(Observation_number = 1:n())
+     data <- full_join(x, y, by = "Observation_number") %>% rename(Group = Group.x) %>% select(-Group.y)
+     xname <- "x"
+     yname <- "yy"
+     group <- "Group"
+   }else{
+     stop("`x`` and `y` must have only 1 or the same number of columns", call. = FALSE)
+   }
+  
+  return(list(data=data, xname=xname, yname=yname, group=group, ncolx=ncolx, ncoly=ncoly))
+}

### 2021-08-06 based on ggpubr/ggmatlist.R

@@ -1,0 +1,5 @@
+ggmatlist <- function (x){
+  max.length <- max(sapply(x, length))
+  x <- lapply(x, function(v) { c(v, rep(NA, max.length-length(v)))})
+  x<- do.call(cbind, x)
+}

### 2021-08-06 based on ggpubr/ggmatplot.R

@@ -1,0 +1,137 @@
+ggmatplot <- function (x, y, color = NULL, shape = NULL, linetype = NULL, fill=NULL,
+                       xlim = c(NA, NA), ylim = c(NA, NA), log = "",
+                       main = NULL, xlab = NULL, ylab = NULL,
+                       legend_label =  NULL, legend_title = NULL,
+                       plot.type = "point", asp = NA){
+
+  # caller_env <- parent.frame()
+
+  if (! plot.type %in% c("point", "line", "both", "density", "histogram", "boxplot", "dotplot","errorplot",
+                         "violin","ecdf") ) 
+    stop("`plot.type` can not be this character", call. = FALSE)
+  
+  if (!missing(x)&!missing(y)){
+    data.list <- ggmatclean (x=x, y=y)
+  } else if (!missing(x)){
+    data.list <- ggmatclean (x=x)
+  } else{
+    stop("No data have been put into the ggmatplot function", call. = FALSE)
+  }
+  
+  data <- data.list$data %>% drop_na()
+  xname <- data.list$xname
+  yname <- data.list$yname
+  ncolx <- data.list$ncolx
+  ncoly <- data.list$ncoly
+  group <- data.list$group 
+  
+  dcolor <- data.list$group 
+  dfill <- data.list$group
+  dshape <- data.list$group
+  dlinetype <- data.list$group
+
+  if(missing(y)& ncolx==1){
+    dcolor = "black" 
+    dfill = "lightgray"
+    dshape = 19
+    dlinetype = 1
+  }
+  
+  
+  if(plot.type=="point"){
+    p <-ggscatter(data = data, x = xname, y = yname, color = dcolor, shape=dshape)
+  }
+  if(plot.type=="line"){
+    p <-ggline(data = data, x = xname, y = yname, color= dcolor, linetype = dlinetype, plot_type = "l")
+  }
+  if(plot.type=="both"){
+    p <-ggline(data = data, x = xname, y = yname, color= dcolor, linetype = dlinetype, shape=dshape)
+  }
+  if(plot.type=="density"){
+    p <-ggdensity(data = data, x = yname, color = dcolor, fill = dfill)+ xlab("x")
+  }
+  if(plot.type=="histogram"){
+    p <-gghistogram(data = data, x = yname, color = dcolor, fill = dfill)+ xlab("x")
+  }
[…]

### 2021-08-06 based on ggpubr/ggmatplot1.R

@@ -1,0 +1,146 @@
+ggmatplot1 <- function (x, y, color = NULL, shape = NULL, linetype = NULL, fill=NULL,
+                       xlim = c(NA, NA), ylim = c(NA, NA), log = "",
+                       main = NULL, xlab = NULL, ylab = NULL,
+                       legend_label =  NULL, legend_title = NULL,
+                       plot.type = "point", asp = NA, ...){
+  
+  # caller_env <- parent.frame()
+  
+  if (! plot.type %in% c("point", "line", "both", "density", "histogram", "boxplot", "dotplot","errorplot",
+                         "violin","ecdf") ) 
+    stop("`plot.type` can not be this character", call. = FALSE)
+  
+  if (!missing(x)&!missing(y)){
+    data.list <- ggmatclean (x=x, y=y)
+  } else if (!missing(x)){
+    data.list <- ggmatclean (x=x)
+  } else{
+    stop("No data have been put into the ggmatplot function", call. = FALSE)
+  }
+  
+  data <- data.list$data %>% drop_na()
+  xname <- data.list$xname
+  yname <- data.list$yname
+  ncolx <- data.list$ncolx
+  ncoly <- data.list$ncoly
+  group <- data.list$group 
+  
+  dcolor <- data.list$group 
+  dfill <- data.list$group
+  dshape <- data.list$group
+  dlinetype <- data.list$group
+  
+  if(missing(y)& ncolx==1){
+    dcolor = "black" 
+    dfill = "lightgray"
+    dshape = 19
+    dlinetype = 1
+  }
+  
+  params <- list(...)
+  params$data <- data
+  params$color <- dcolor
+  params$fill<- dfill
+  params$shape <- dshape
+  params$dlinetype <- dlinetype
+  
+  if(plot.type=="point"){
+    p <-ggscatter(data = data, x = xname, y = yname, color = dcolor, shape=dshape)
+  }
+  if(plot.type=="line"){
+    p <-ggline(data = data, x = xname, y = yname, color= dcolor, linetype = dlinetype, plot_type = "l")
+  }
+  if(plot.type=="both"){
+    p <-ggline(data = data, x = xname, y = yname, color= dcolor, linetype = dlinetype, shape=dshape, ...)
+  }
+  if(plot.type=="density"){
[…]

### 2021-10-08 R/ggmatplot.R
Updated function documentation

@@ -1,20 +1,17 @@
-#' Quick matplot function
+#' Quick function to plot 2 matrices against each other.
 #'
-#' `ggmatplot` is a quick way to make a simplier and similar plot as what `matplot` does but with a ggplot version.
+#' ggmatplot is a quick way to plot 2 matrices against each other using ggplot2.
 #'
-#' @param x,y vectors or matrices of data for plotting. The number of rows should match. If one of them are missing, the other is taken as y and an x vector of 1:n is used. Missing values (NAs) are allowed.
-#' @param color vector of colors and they are used cyclically.
-#' @param shape vector of shapes and they are used cyclically.
-#' @param linetype vector of linetypes colors  and they are used cyclically.
-#' @param xlim,ylim ranges of x and y axes, as in [plot()].
-#' @param log Which variables to log transform ("x", "y", or "xy")
-#' @param main,xlab,ylab Character vector giving plot title,
-#'   x axis label, and y axis label respectively.
-#' @param legend_label Character vector giving legend tables for different groups.
-#' @param legend_title Character giving legend title
-#' @param geom Character vector specifying geom(s) to draw. Defaults to
-#'  "point". Other options are "line" or c("point","line").
-#' @param asp The y/x aspect ratio
+#' @param x,y Vectors or matrices of data for plotting. The number of rows should match. If one of them are missing, the other is taken as y and an x vector of 1:n is used. Missing values (NAs) are allowed.
+#' @param color,fill Vectors of colors. Should match the number of unique groups created when x and y are pivoted into long format. If a single string, the same color will be used for all groups. Defining only one of them will update both `color` and `fill` aesthetics of the plot by default, unless they are both defined simultaneously.
+#' @param shape,linetype A vector of shapes or line types respectively. Should match the number of unique groups created when x and y are pivoted into long format. If a single string, the same shape or line type will be used for all groups.
+#' @param xlim,ylim Ranges of x and y axes. Each of them should be a two element vector specifying the lower and upper limits of the scale. If the larger value is given first, the scale will be reversed. If one of the limits is given as `NA`, the corresponding limit from the range of the data will be used.
[…]

### 2021-10-08 R/matclean.R

@@ -1,9 +1,12 @@
-#' Function to convert two matrices into a wide format dataframe
+#' Function to convert two matrices into a wide format data frame
 #'
-#' @param x,y vector or matrices of data. The number of rows should match. If one of them are missing, the other is taken as y and an x vector of 1:n is used.
+#' @param x,y Vectors or matrices of data. The number of rows of x and y should match. If one of them are missing, the other is taken as y and an x vector of 1:n is used.
 #'
-#' @return a list containing the long format dataframe, number and names of the id columns, number and names of the pivoted columns (column with group names and values)
-#' @NoRd
+#' @return A list of containing:
+#'  * data - the long format data frame.
+#'  * xname - name(s) of the ID column(s)
+#'  * yname - names of the pivoted columns(columns with group names and values)
+#' @noRd
 #'
 #' @examples
 #' # Define x and y matrices
@@ -12,6 +15,7 @@
 #' y <- iris_sub[, 3:4]
 #' # Use the defined x and y matrices as parameters
 #' matclean(x, y)
+#'
 matclean <- function(x, y) {
   x <- data.frame(x)
   ncolx <- ncol(x)

### 2021-10-08 R/pivotlonger.R

@@ -1,18 +1,19 @@
 #' Function to convert a wide dataframe into long format
 #'
-#' @param data dataframe to pivot.
-#' @param names_to a string specifying the name of the column to create from the column names of those being pivoted.
-#' @param values_to a string specifying the name of the column to create from the data stored in cell values of the columns being pivoted.
-#' @param id_cols columns that uniquely identify observations. Columns to not pivot into longer format.
+#' @param data A data frame to pivot.
+#' @param names_to A string specifying the name of the column to create from the column names of those being pivoted.
+#' @param values_to A string specifying the name of the column to create from the data stored in the columns being pivoted.
+#' @param id_cols 	[<tidy-select>](https://tidyr.tidyverse.org/reference/tidyr_tidy_select.html) Columns that uniquely identify observations and not to pivot into longer format.
 #'
-#' @return a dataframe in long format
-#' @NoRd
+#' @return A long format data frame.
+#' @noRd
 #'
 #' @examples
 #' # Define a data set
 #' iris_sub <- subset(iris, Species == "setosa")[1:4]
 #' # Use first two columns as ID columns and pivot the second two columns to long format
-#' pivotlonger(iris_sub, names_to = "Measurement", values_to = "Value", c(1, 2))
+#' pivotlonger(data = iris_sub, names_to = "Measurement", values_to = "Value", id_cols = c(1, 2))
+#'
 pivotlonger <- function(data, names_to = "", values_to = "", id_cols = NULL) {
   long_df <- data.frame(matrix(ncol = 2 + length(id_cols), nrow = 0))
   colnames(long_df) <- c(colnames(data[id_cols]), names_to, values_to)

### 2021-10-08 R/validateNumParams.R

@@ -1,13 +1,13 @@
 #' Function to validate the number of parameter values against the number of groups
 #'
-#' @param parameterValues a list of values. Should be of length 1 or correspond to the number of groups given by the 'numGroups' parameter.
-#' @param numGroups an integer denoting the number of unique values of the grouping variable
+#' @param parameterValues A list of values.
+#' @param numGroups An integer denoting the number of unique values of the grouping variable.
 #'
-#' @return a list of parameter values.
-#' @NoRd
+#' @return If valid, a list of parameter values. If invalid, an error message.
+#' @noRd
 #'
-numParameterHandler <- function(parameterValues, numGroups) {
-  # same for all groups
+validateNumParams <- function(parameterValues, numGroups) {
+  # set same value for all groups
   if (!numGroups == 1 & length(parameterValues) == 1) {
     return(rep(parameterValues, numGroups))
   }
@@ -18,7 +18,9 @@ numParameterHandler <- function(parameterValues, numGroups) {
   # values < number of unique groups
   else if (length(parameterValues) < numGroups) {
     stop(paste0("Insufficient ", substitute(parameterValues), " values. ", numGroups, " needed but only ", length(parameterValues), " provided."), call. = FALSE)
-  } else {
+  }
+  # values = number of unique groups
+  else {
     return(parameterValues)
   }
 }

### 2021-10-08 tests/testthat/test_ggmatplot_parameters.R
Updated too long test snapshot names

@@ -47,11 +47,11 @@ test_that("plot axis limits are updated based on xlim and ylim parameters", {
 
 test_that("plot axes are transformed to a log scale based on log parameter values", {
   # log(x)
-  expect_doppelganger("line plot with log transformed x axis", ggmatplot(x,y, plot.type = "line", log = "x"))
+  expect_doppelganger("line plot with log x", ggmatplot(x,y, plot.type = "line", log = "x"))
   # log(y)
-  expect_doppelganger("scatterplot with log transformed y axis", ggmatplot(x,y, plot.type = "point", log = "y"))
+  expect_doppelganger("scatterplot with log y", ggmatplot(x,y, plot.type = "point", log = "y"))
   # log(x) and log(y)
-  expect_doppelganger("point + line plot with log transformed x and y axis", ggmatplot(x,y, plot.type = "both", log = "xy"))
+  expect_doppelganger("point + line plot with log x and y", ggmatplot(x,y, plot.type = "both", log = "xy"))
 })
 
 test_that("plot title is updated based on main paramter", {

### 2021-10-08 README.Rmd
commit
Added basic pkgdown site

@@
 ---
-
-<!-- README.md is generated from README.Rmd. Please edit that file -->
 
@@
 
-<!-- badges: start -->
-<!-- badges: end -->
-
-The goal of ggmatplot is to plot the columns of one matrix against the columns of another.
+`ggmatplot` plots two matrices against each other using [`ggplot2`](https://ggplot2.tidyverse.org/)
 
@@
 
-You can install the released version of ggmatplot by the following commands:
+You can install the released version of ggmatplot from [GitHub](https://github.com/xuan-liang/ggmatplot):
 
 ```r
-library(devtools)
-install_github("xuan-liang/ggmatplot")
+# install.packages("remotes")
+remotes::install_github("xuan-liang/ggmatplot")
 ```
-
-Below are some basic examples which show you how to use this package.
-
-### Example 1
-
-```{r load_package, echo=FALSE, message=FALSE, warning=FALSE}
-library(ggmatplot)
-```
-
-```{r example1}
-X <- 1:10
-Y=X^2
-Z=X^2-2*X
-
-# Original way by matplot function
-matplot(X,cbind(Y,Z),pch=c(16,1),xlab="x",ylab="X^2 and X^2-2*X")
-legend(1,100,legend=c("y","Z"),pch=c(16,1))
-
-# Use ggmatplot
-ggmatplot(X,cbind(Y,Z), shape = c(16,1),legend_label =  c("X^2","X^2-2*X"),geom=c("point","line"))
-```
-
-#### Example 2: Iris data
-
-```{r example2}
-iris_sub <- subset(iris, Species == "setosa")
-
-# Original way by matplot function
-matplot(iris_sub[,c(1,3)], iris_sub[,c(2,4)])
-
-# Use default legend and axis
-ggmatplot(iris_sub[,c(1,3)], iris_sub[,c(2,4)])
-
-# Modify legend label and axis
-ggmatplot(iris_sub[,c(1,3)], iris_sub[,c(2,4)], color = 1 ,shape = c("S","P"), xlab = "Length", ylab="Width")
-```
-
-### Example 3: Linear regression
-
-```{r example3}
-x <- rnorm(100,sd=2)
-y <- x*0.5+rnorm(100,sd=1)
-fit.y <- fitted(lm(y~x))
-ggmatplot(x,cbind(y,fit.y))
-```

### 2021-10-10 tests/testthat/test_ggmatplot_parameters.R
Added new ggmatplot parameter tests

@@ -14,6 +14,24 @@ test_that("plot colors by color values", {
   expect_error(ggmatplot(y, plot.type = "violin", color = c("red", "blue", "yellow", "green")), "Too many color values. Only 3 needed but 4 provided.")
 })
 
+test_that("plot colors by fill values", {
+  # single fill value
+  expect_doppelganger("single fill density plot", ggmatplot(y, plot.type = "density", fill = "red"))
+  # fill values = number of groups
+  expect_doppelganger("three fill color violin plot", ggmatplot(y, plot.type = "violin", fill = c("red", "blue", "#123456")))
+  # fill values < number of groups
+  expect_error(ggmatplot(y, plot.type = "violin", fill = c("red", "blue")), "Insufficient fill values. 3 needed but only 2 provided.")
+  # fill values > number of groups
+  expect_error(ggmatplot(y, plot.type = "violin", fill = c("red", "blue", "yellow", "green")), "Too many fill values. Only 3 needed but 4 provided.")
+})
+
+test_that("plot colors by color and fill values simultaneously", {
+  # single color value, fill values = number of groups
+  expect_doppelganger("color and fill density plot", ggmatplot(y, plot.type = "density", color = "red", fill = c("red", "blue", "#123456")))
+  # single fill value, color values < number of groups
+  expect_error(ggmatplot(y, plot.type = "violin", fill = "black", color = c("red", "blue")), "Insufficient fill values. 3 needed but only 2 provided.")
+})
+
 test_that("point shapes updated based on shape values", {
   # single shape value
   expect_doppelganger("single shape scatterplot", ggmatplot(y, plot.type = "point", shape = "square"))
@@ -64,29 +82,37 @@ test_that("plot axis labels are updated based on xlab and ylab paramters", {
   expect_equal(ggmatplot(y, plot.type = "point", ylab="ylab test")$labels$y, "ylab test")
 })
 
+test_that("invalid plot types are no allowed", {
+  expect_error(ggmatplot(x,y, plot.type = "scatterplot"), "plot.type can not take this value")
+})
+
+test_that("legend labels are updated", {
+  # legend label values = number of groups
+  expect_doppelganger("plot with updated legend labels", ggmatplot(x,y, legend_label = c("lab1","lab2","lab3")))
+  # legend label values < number of groups
[…]

### 2021-10-10 R/ggmatplot.R
Updated error messages

@@ -28,7 +28,7 @@ ggmatplot <- function(x, y, color = NULL, shape = NULL, linetype = NULL, fill = 
                       legend_label = NULL, legend_title = NULL,
                       plot.type = "point", asp = NA, ...) {
   if (!plot.type %in% c("point", "line", "both", "density", "histogram", "boxplot", "dotplot", "errorplot", "violin", "ecdf")) {
-    stop("`plot.type` can not take this value", call. = FALSE)
+    stop("plot.type can not take this value", call. = FALSE)
   }
   if (!missing(x) & !missing(y)) {
     if (plot.type %in% c("density", "histogram", "boxplot", "violin")) {
@@ -38,7 +38,7 @@ ggmatplot <- function(x, y, color = NULL, shape = NULL, linetype = NULL, fill = 
   } else if (!missing(x)) {
     data.list <- matclean(x = x)
   } else {
-    stop("`x` can not be missing", call. = FALSE)
+    stop("x can not be missing", call. = FALSE)
   }
 
   params <- list(...)

### 2021-10-22 R/ggmatplot-package.R
Added new homepage structure

@@ -1,7 +1,0 @@
-#' ggmatplot
-#'
-#' This package was created as a way to make a matplot with ggplot version.
-#'
-#' @docType package
-#' @name ggmatplot
-NULL

### 2021-10-22 R/ggmatplot.R

@@ -1,4 +1,4 @@
-#' Quick function to plot 2 matrices against each other.
+#' ggmatplot
 #'
 #' `ggmatplot` is a quick way to plot wide format data. This can be done by passing the id columns and columns to be pivoted into long format, to the `ggmatplot` function as 2 matrices. The `ggmatplot` function will then plot these matrices against each other using [`ggplot2`](https://ggplot2.tidyverse.org/).
 #'
@@ -21,8 +21,7 @@
 #' iris_sub <- subset(iris, Species == "setosa")
 #' ggmatplot(iris_sub[, c(1, 3)], iris_sub[, c(2, 4)])
 #' # Modify legend label and axis
-#' ggmatplot(iris_sub[, c(1, 3)], iris_sub[, c(2, 4)], shape = c(4,6), legend_label = c("Sepal", "Petal"), legend_title = "", xlab = "Length", ylab = "Width")
-#'
+#' ggmatplot(iris_sub[, c(1, 3)], iris_sub[, c(2, 4)], shape = c(4, 6), legend_label = c("Sepal", "Petal"), legend_title = "", xlab = "Length", ylab = "Width")
 ggmatplot <- function(x, y, color = NULL, shape = NULL, linetype = NULL, fill = NULL,
                       xlim = c(NA, NA), ylim = c(NA, NA), log = "",
                       main = NULL, xlab = NULL, ylab = NULL,
@@ -57,108 +56,114 @@ ggmatplot <- function(x, y, color = NULL, shape = NULL, linetype = NULL, fill = 
     )
   )
 
-
-  if (plot_type == "point") {
-    p <- p +
-      do.call(
-        "geom_point",
-        c(
-          list(mapping = aes(
-            x = .data[[xname]],
-            y = .data[[yname]],
-            shape = Group
-          )),
-          params
+  withCallingHandlers(
+    if (plot_type == "point") {
+      p <- p +
+        do.call(
+          "geom_point",
+          c(
+            list(mapping = aes(
+              x = .data[[xname]],
+              y = .data[[yname]],
+              shape = Group
+            )),
+            params
+          )
         )
-      )
-  } else if (plot_type == "line") {
-    p <- p +
-      do.call(
-        "geom_line",
-        c(
-          list(mapping = aes(
-            x = .data[[xname]],
-            y = .data[[yname]],
-            linetype = Group
-          )),
-          params
+    } else if (plot_type == "line") {
+      p <- p +
+        do.call(
+          "geom_line",
+          c(
[…]

### 2021-10-22 tests/testthat.R
Corrected unit test errors

@@ -1,4 +1,5 @@
 library(testthat)
+library(tibble)
 library(ggmatplot)
 
 test_check("ggmatplot")

### 2021-10-22 tests/testthat/test_ggmatplot_parameters.R

@@ -29,7 +29,7 @@ test_that("plot colors by color and fill values simultaneously", {
   # single color value, fill values = number of groups
   expect_doppelganger("color and fill density plot", ggmatplot(y, plot_type = "density", color = "red", fill = c("red", "blue", "#123456")))
   # single fill value, color values < number of groups
-  expect_error(ggmatplot(y, plot_type = "violin", fill = "black", color = c("red", "blue")), "Insufficient fill values. 3 needed but only 2 provided.")
+  expect_error(ggmatplot(y, plot_type = "violin", fill = "black", color = c("red", "blue")), "Insufficient color values. 3 needed but only 2 provided.")
 })
 
 test_that("point shapes updated based on shape values", {

### 2021-10-22 R/ggmatplot.R
Reverting ggmatplot warning handler

@@ -56,114 +56,107 @@ ggmatplot <- function(x, y, color = NULL, shape = NULL, linetype = NULL, fill = 
     )
   )
 
-  withCallingHandlers(
-    if (plot_type == "point") {
-      p <- p +
-        do.call(
-          "geom_point",
-          c(
-            list(mapping = aes(
-              x = .data[[xname]],
-              y = .data[[yname]],
-              shape = Group
-            )),
-            params
-          )
+  if (plot_type == "point") {
+    p <- p +
+      do.call(
+        "geom_point",
+        c(
+          list(mapping = aes(
+            x = .data[[xname]],
+            y = .data[[yname]],
+            shape = Group
+          )),
+          params
         )
-    } else if (plot_type == "line") {
-      p <- p +
-        do.call(
-          "geom_line",
-          c(
-            list(mapping = aes(
-              x = .data[[xname]],
-              y = .data[[yname]],
-              linetype = Group
-            )),
-            params
-          )
+      )
+  } else if (plot_type == "line") {
+    p <- p +
+      do.call(
+        "geom_line",
+        c(
+          list(mapping = aes(
+            x = .data[[xname]],
+            y = .data[[yname]],
+            linetype = Group
+          )),
+          params
         )
-    } else if (plot_type == "both") {
-      p <- p +
-        do.call(
-          "geom_point",
-          c(
-            list(mapping = aes(
-              x = .data[[xname]],
-              y = .data[[yname]],
-              shape = Group
-            )),
-            params
-          )
-        ) +
-        do.call(
-          "geom_line",
-          c(
-            list(mapping = aes(
-              x = .data[[xname]],
-              y = .data[[yname]],
-              linetype = Group
-            )),
-            params
-          )
+      )
+  } else if (plot_type == "both") {
+    p <- p +
+      do.call(
+        "geom_point",
+        c(
+          list(mapping = aes(
+            x = .data[[xname]],
+            y = .data[[yname]],
+            shape = Group
+          )),
+          params
         )
-    } else if (plot_type == "density") {
-      params$alpha <- if (is.null(params$alpha)) 0.5 else params$alpha
-      p <- p +
-        do.call(
-          "geom_density",
-          c(
-            list(mapping = aes(
-              x = .data[[yname]]
-            )),
-            params
-          )
+      ) +
+      do.call(
[…]

### 2021-10-23 R/ggmatplot.R
Updated ggmatplot function argument documentation

@@ -1,27 +1,70 @@
 #' ggmatplot
 #'
-#' `ggmatplot` is a quick way to plot wide format data. This can be done by passing the id columns and columns to be pivoted into long format, to the `ggmatplot` function as 2 matrices. The `ggmatplot` function will then plot these matrices against each other using [`ggplot2`](https://ggplot2.tidyverse.org/).
 #'
-#' @param x,y Vectors or matrices of data for plotting. The number of rows should match. If one of them are missing, the other is taken as y and an x vector of 1:n is used. Missing values (NAs) are allowed.
-#' @param color,fill Vectors of colors. Should match the number of unique groups created when x and y are pivoted into long format. If a single string, the same color will be used for all groups. Defining only one of them will update both `color` and `fill` aesthetics of the plot by default, unless they are both defined simultaneously.
-#' @param shape,linetype A vector of shapes or line types respectively. Should match the number of unique groups created when x and y are pivoted into long format. If a single string, the same shape or line type will be used for all groups.
-#' @param xlim,ylim Ranges of x and y axes. Each of them should be a two element vector specifying the lower and upper limits of the scale. If the larger value is given first, the scale will be reversed. If one of the limits is given as `NA`, the corresponding limit from the range of data will be used.
-#' @param log A string defining which axes to transform into a log scale. (`x`, `y` or `xy`)
-#' @param main,xlab,ylab,legend_title Strings to update plot title, x axis label, y axis label and legend title respectively.
-#' @param legend_label A list of strings, to rename the legend labels. Should match the number of unique groups created when x and y are pivoted into long format.
-#' @param plot_type A string specifying the type of plot to draw. Possible plot types are `point`, `line`, `both`(point + line), `density`, `histogram`, `boxplot`, `dotplot`, `errorplot`, `violin`, and `ecdf`. Default plot_type is `point`.
+#' `ggmatplot` is a quick and easy way of plotting two matrices against each
+#' other. Can also be used as an alternative to pivoting data into long format
+#' before plotting, by using the index column(s) and the columns to be plotted
+#' against the index as the two matrix inputs for `ggmatplot()`.
+#'
+#' @param x,y Vectors or matrices of data for plotting.
+#'
[…]

### 2021-10-24 R/ggmatplot.R
Validated ggmatplot parameters

@@ -2,9 +2,7 @@
 #'
 #'
 #' `ggmatplot` is a quick and easy way of plotting two matrices against each
-#' other. Can also be used as an alternative to pivoting data into long format
-#' before plotting, by using the index column(s) and the columns to be plotted
-#' against the index as the two matrix inputs for `ggmatplot()`.
+#' other.
 #'
 #' @param x,y Vectors or matrices of data for plotting.
 #'
@@ -63,17 +61,21 @@
 #' ggmatplot(iris_sub[, c(1, 3)], iris_sub[, c(2, 4)])
 #' # Modify legend label and axis
 #' ggmatplot(iris_sub[, c(1, 3)], iris_sub[, c(2, 4)], shape = c(4, 6), legend_label = c("Sepal", "Petal"), legend_title = "", xlab = "Length", ylab = "Width")
-#'
-#'
-ggmatplot <- function(x, y, color = NULL, shape = NULL, linetype = NULL, fill = NULL,
-                      xlim = c(NA, NA), ylim = c(NA, NA), log = "",
-                      main = NULL, xlab = NULL, ylab = NULL,
+ggmatplot <- function(x, y, color = NULL, shape = NULL, linetype = NULL,
+                      fill = NULL, xlim = c(NA, NA), ylim = c(NA, NA),
+                      log = NULL, main = NULL, xlab = NULL, ylab = NULL,
                       legend_label = NULL, legend_title = NULL,
                       plot_type = "point", asp = NA, ...) {
-  if (!plot_type %in% c("point", "line", "both", "density", "histogram", "boxplot", "dotplot", "errorplot", "violin", "ecdf")) {
+
+  # valid plot types
+  if (!plot_type %in% c(
+    "point", "line", "both", "density", "histogram",
+    "boxplot", "dotplot", "errorplot", "violin", "ecdf"
+  )) {
     stop("plot_type can not take this value", call. = FALSE)
   }
   if (!missing(x) & !missing(y)) {
+    # only single matrix input allowed for the following plot types
     if (plot_type %in% c("density", "histogram", "boxplot", "violin")) {
       stop("This plot type only uses a single matrix input", call. = FALSE)
     }
@@ -204,16 +206,19 @@ ggmatplot <- function(x, y, color = NULL, shape = NULL, linetype = NULL, fill = 
   # number of unique groups
   numGroups <- length(unique(data$Group))
 
+  # default legend title
[…]

### 2021-10-24 R/ggmatplot.R
Updated tests for new ggmatplot parameter validation

@@ -212,13 +212,11 @@ ggmatplot <- function(x, y, color = NULL, shape = NULL, linetype = NULL,
   if (!is.null(legend_label)) {
     # values > number of unique groups
     if (length(legend_label) > numGroups) {
-      stop(paste0("Too many legend_label values. Only ", numGroups, " needed but
-                  ", length(legend_label), " provided."), call. = FALSE)
+      stop(paste0("Too many legend_label values. Only ", numGroups, " needed but " , length(legend_label), " provided."), call. = FALSE)
     }
     # values < number of unique groups
     else if (length(legend_label) < numGroups) {
-      stop(paste0("Insufficient legend_label values. ", numGroups, " needed but
-                  only ", length(legend_label), " provided."), call. = FALSE)
+      stop(paste0("Insufficient legend_label values. ", numGroups, " needed but only ", length(legend_label), " provided."), call. = FALSE)
     }
   } else {
     legend_label <- unique(data$Group)
@@ -293,7 +291,7 @@ ggmatplot <- function(x, y, color = NULL, shape = NULL, linetype = NULL,
   if (!is.null(log)) {
     # validating list of values the log parameter can take
     if (!log %in% c("x", "y", "xy")) {
-      stop("log can not take this value", call. = FALSE)
+      stop("invalid log value provided", call. = FALSE)
     } else {
       logv <- function(var) var %in% strsplit(log, "")[[1]]
       if (logv("x")) {

### 2021-10-24 tests/testthat/test_ggmatplot_parameters.R

@@ -112,12 +112,12 @@ test_that("invalid parameter values throw errors", {
   expect_error(ggmatplot(y, plot_type = "line", xlim = c(-4,a)), "ERROR")
   expect_error(ggmatplot(y, plot_type = "line", xlim = 5), "ERROR")
   # invalid log value
-  expect_error(ggmatplot(y, plot_type = "point", log = 1), "ERROR")
+  expect_error(ggmatplot(y, plot_type = "point", log = 1), "invalid log value provided")
 })
 
 test_that("unknown parameter values for plots throw errors", {
   # shapes for plot types that don't require shapes
-  expect_error(ggmatplot(y, plot_type = "density", shape = c(15, 12, 13)), "ERROR")
+  expect_warning(ggmatplot(y, plot_type = "density", shape = c(15, 12, 13)), "Ignoring unknown parameters: shape")
   # linetypes for plot types that don't require linetypes
-  expect_error(ggmatplot(x, y, plot_type = "point", linetype = c(15, 12, 13)), "ERROR")
+  expect_warning(ggmatplot(x, y, plot_type = "point", linetype = c(15, 12, 13)), "Ignoring unknown parameters: linetype")
 })

### 2021-10-24 tests/testthat/test_matclean.R

@@ -1,3 +1,5 @@
+library(tibble)
+
 test_that("converts a single matrix into long format", {
 
   x <- tibble(a = 1:3, b = 11:13, c = 31:33)

### 2021-10-24 README.Rmd
Updated readme pacakge description

@@
 
-`ggmatplot` plots two matrices against each other using [`ggplot2`](https://ggplot2.tidyverse.org/)
+`ggmatplot` is a quick and easy way of plotting the columns of two matrices against each other.
+It is built upon [ggplot2](https://ggplot2.tidyverse.org/), and is aimed at providing a more direct and simple approach to creating plots using matrices.
+
+`ggmatplot` can also be considered as an alternative to pivoting data into long format for visualizing using `ggplot2`.
 
@@
 
-You can install the released version of ggmatplot from [GitHub](https://github.com/xuan-liang/ggmatplot):
+You can install the released version of `ggmatplot` from [GitHub](https://github.com/xuan-liang/ggmatplot):
 
@@
 ```
+
+## Examples
+
+```{r message=FALSE}
+library(ggmatplot)
+```
+
+```{r}
+USPersonalExpenditure
+
+ggmatplot(
+  rownames(USPersonalExpenditure),
+  USPersonalExpenditure[, 1:5],
+  plot_type = "point",
+  xlab = "Category",
+  ylab = "Expenditure (in Billions of Dollars)",
+  legend_title = "Year",
+  legend_label = c(1940, 1945, 1950, 1955, 1960)
+) +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```

### 2021-10-24 README.Rmd
updated pkgdown site home page

--- 2021-10-08 README.Rmd
@@
   fig.path = "man/figures/README-",
-  out.width = "100%"
+  out.width = "80%",
+  message = FALSE
 )
@@
 
-`ggmatplot` plots two matrices against each other using [`ggplot2`](https://ggplot2.tidyverse.org/)
+`ggmatplot` is a quick and easy way of plotting the columns of two matrices 
+against each other. It is built upon [ggplot2](https://ggplot2.tidyverse.org/), 
+and is aimed at providing a more direct and simple approach to creating plots 
+using matrices.
+
+`ggmatplot` can also be considered as an alternative to pivoting data into long 
+format and visualizing using `ggplot2`.
 
@@
 
-You can install the released version of ggmatplot from [GitHub](https://github.com/xuan-liang/ggmatplot):
+The latest version can be installed from [GitHub](https://github.com/xuan-liang/ggmatplot):
 
@@
 ```
+
+## Examples
+
+```{r}
+library(ggmatplot)
+library(ggplot2)
+```
+
+```{r point}
+# matrix x
+x <- rnorm(100,sd=2)
+
+# matrix y
+y = x*0.5+rnorm(100,sd=1)
+fit.y = fitted(lm(y~x))
+z <- cbind(y, fit.y)
+
+ggmatplot(x, z)
+```
+
+```{r line}
+# matrix x
+x <- 1:10
+
+# matrix y
+y <- cbind(
+  square = x^2,
+  cube = x^3
+)
+
+ggmatplot(x, y,
+          plot_type = "line",
+          size = 1,
+          color = c("blue", "purple"),
+          ylim = c(0,750)) +
+  theme_minimal()
+```
+
+```{r both}
+USPersonalExpenditure
+
+# matrix x
+x <- rownames(USPersonalExpenditure)
+
+# matrix y
+y <- USPersonalExpenditure[, 1:5]
+
+ggmatplot(x, y,
+          plot_type = "both",
+          xlab = "Category",
+          ylab = "Expenditure (in Billions of Dollars)",
+          legend_title = "Year",
+          legend_label = c(1940, 1945, 1950, 1955, 1960)
+          ) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+
+```{r density}
+iris_sub <- subset(iris, Species == "setosa")
+
+# matrix X
+x <- (iris_sub[,1:2])
+head(x,5)
+
+ggmatplot(x, plot_type = "density") + 
+  # scaling the density estimate to a maximum of 1
+  aes(y = ..scaled..) +
+  theme_bw()
+```
+
+```{r boxplot}
+iris_sub <- subset(iris, Species == "setosa")
+
+# matrix X
+x <- (iris_sub[,1:4])
+head(x,5)
+
+ggmatplot(x, plot_type = "boxplot",
+          alpha = 0, # removing fill values
+          xlab = "", ylab = "")
+```
+
+```{r violin}
+iris_sub <- subset(iris, Species == "setosa")
+
+# matrix X
+x <- (iris_sub[,1:2])
+head(x,5)
+
+ggmatplot(x, plot_type = "violin",
+          color = c("#00AFBB", "#E7B800"),
+          xlab = "", ylab = "")
+```
+
+```{r histogram}
+iris_sub <- subset(iris, Species == "setosa")
+
+# matrix X
+x <- (iris_sub[,1:4])
+head(x,5)
+
+ggmatplot(x, plot_type = "histogram",
+          xlab = "Group",
+          color = "black",
+          fill = c("#F8766D", "#7CAE00", "#00BFC4", "#C77CFF"))
+```

### 2021-10-24 R/ggmatplot.R

@@ -1,8 +1,8 @@
 #' ggmatplot
 #'
 #'
-#' `ggmatplot` is a quick and easy way of plotting two matrices against each
-#' other.
+#' `ggmatplot` is a quick and easy way of plotting the columns of two matrices
+#' against each other.
 #'
 #' @param x,y Vectors or matrices of data for plotting.
 #'

### 2021-10-24 R/ggmatplot.R
Updated invalid paramter error message

@@ -261,7 +261,7 @@ ggmatplot <- function(x, y, color = NULL, shape = NULL, linetype = NULL,
     if (plot_type %in% c("point", "both")) {
       shape <- validateNumParams(shape, numGroups)
     } else {
-      warning("Ignoring unknown parameters: shape", call. = FALSE)
+      warning(paste0("shape is an invalid parameter for plot type: ", plot_type), call. = FALSE)
     }
     p <- p + scale_shape_manual(
       name = legend_title, labels = legend_label,
@@ -276,7 +276,7 @@ ggmatplot <- function(x, y, color = NULL, shape = NULL, linetype = NULL,
     if (!plot_type %in% c("point")) {
       linetype <- validateNumParams(linetype, numGroups)
     } else {
-      warning("Ignoring unknown parameters: linetype", call. = FALSE)
+      warning(paste0("linetype is an invalid parameter for plot type: ", plot_type), call. = FALSE)
     }
     p <- p + scale_linetype_manual(
       name = legend_title, labels = legend_label,

### 2021-10-24 tests/testthat/test_ggmatplot_parameters.R

@@ -117,7 +117,7 @@ test_that("invalid parameter values throw errors", {
 
 test_that("unknown parameter values for plots throw errors", {
   # shapes for plot types that don't require shapes
-  expect_warning(ggmatplot(y, plot_type = "density", shape = c(15, 12, 13)), "Ignoring unknown parameters: shape")
+  expect_warning(ggmatplot(y, plot_type = "density", shape = c(15, 12, 13)), "shape is an invalid parameter for plot type: density")
   # linetypes for plot types that don't require linetypes
-  expect_warning(ggmatplot(x, y, plot_type = "point", linetype = c(15, 12, 13)), "Ignoring unknown parameters: linetype")
+  expect_warning(ggmatplot(x, y, plot_type = "point", linetype = c(15, 12, 13)), "linetype is an invalid parameter for plot type: point")
 })

### 2021-10-29 R/ggmatplot.R
Updated ggmatplot lims parameter validation and tests

@@ -308,8 +308,18 @@ ggmatplot <- function(x, y, color = NULL, shape = NULL, linetype = NULL,
   if (!is.null(xlab)) p <- p + xlab(xlab)
   if (!is.null(ylab)) p <- p + ylab(ylab)
 
-  if (!missing(xlim)) p <- p + xlim(xlim)
-  if (!missing(ylim)) p <- p + ylim(ylim)
+  if (!missing(xlim)) {
+    if(length(xlim) == 2) {
+      p <- p + xlim(xlim)
+    } else {
+      stop("xlim must be a two element vector", call. = FALSE)
+    }}
+  if (!missing(ylim)) {
+    if(length(ylim) == 2) {
+      p <- p + ylim(ylim)
+    } else {
+      stop("ylim must be a two element vector", call. = FALSE)
+    }}
 
   if (!is.na(asp)) p <- p + theme(aspect.ratio = asp)
 

### 2021-10-29 tests/testthat/test_ggmatplot_parameters.R

@@ -8,10 +8,6 @@ test_that("plot colors by color values", {
   expect_doppelganger("single color scatterplot", ggmatplot(x, y, plot_type = "point", color = "red"))
   # color values = number of groups
   expect_doppelganger("three color violin plot", ggmatplot(y, plot_type = "violin", color = c("red", "blue", "#123456")))
-  # color values < number of groups
-  expect_error(ggmatplot(y, plot_type = "violin", color = c("red", "blue")), "Insufficient color values. 3 needed but only 2 provided.")
-  # color values > number of groups
-  expect_error(ggmatplot(y, plot_type = "violin", color = c("red", "blue", "yellow", "green")), "Too many color values. Only 3 needed but 4 provided.")
 })
 
 test_that("plot colors by fill values", {
@@ -19,17 +15,11 @@ test_that("plot colors by fill values", {
   expect_doppelganger("single fill density plot", ggmatplot(y, plot_type = "density", fill = "red"))
   # fill values = number of groups
   expect_doppelganger("three fill color violin plot", ggmatplot(y, plot_type = "violin", fill = c("red", "blue", "#123456")))
-  # fill values < number of groups
-  expect_error(ggmatplot(y, plot_type = "violin", fill = c("red", "blue")), "Insufficient fill values. 3 needed but only 2 provided.")
-  # fill values > number of groups
-  expect_error(ggmatplot(y, plot_type = "violin", fill = c("red", "blue", "yellow", "green")), "Too many fill values. Only 3 needed but 4 provided.")
 })
 
 test_that("plot colors by color and fill values simultaneously", {
   # single color value, fill values = number of groups
   expect_doppelganger("color and fill density plot", ggmatplot(y, plot_type = "density", color = "red", fill = c("red", "blue", "#123456")))
-  # single fill value, color values < number of groups
-  expect_error(ggmatplot(y, plot_type = "violin", fill = "black", color = c("red", "blue")), "Insufficient color values. 3 needed but only 2 provided.")
 })
 
 test_that("point shapes updated based on shape values", {
@@ -37,10 +27,6 @@ test_that("point shapes updated based on shape values", {
   expect_doppelganger("single shape scatterplot", ggmatplot(y, plot_type = "point", shape = "square"))
   # shape values = number of groups
[…]

### 2021-10-29 README.Rmd
Updated package overview on readme

@@
 
-`ggmatplot` is a quick and easy way of plotting the columns of two matrices 
-against each other. It is built upon [ggplot2](https://ggplot2.tidyverse.org/), 
-and is aimed at providing a more direct and simple approach to creating plots 
-using matrices.
+`ggmatplot` is a quick and easy way of plotting the columns of two matrices or 
+data frames against each other using [`ggplot2`](https://ggplot2.tidyverse.org/).
 
-`ggmatplot` can also be considered as an alternative to pivoting data into long 
-format and visualizing using `ggplot2`.
+## Overview
+
+[`ggplot2`](https://ggplot2.tidyverse.org/) requires wide format data to be 
+wrangled into long format for plotting, which can be quite cumbersome when 
+creating simple plots. Therefore, the motivation for `ggmatplot` was to provide 
+a solution that allows [`ggplot2`](https://ggplot2.tidyverse.org/) to handle 
+wide format data. Although `ggmatplot` doesn't provide the same flexibility 
+as [`ggplot2`](https://ggplot2.tidyverse.org/), it can be used as a 
+workaround for having to wrangle wide format data into long format for plotting 
+with [`ggplot2`](https://ggplot2.tidyverse.org/).
+
+`ggmatplot` is built upon [`ggplot2`](https://ggplot2.tidyverse.org/), and its 
+functionality is inspired by 
+[`matplot`](https://www.rdocumentation.org/packages/graphics/versions/3.6.2/topics/matplot).
+Therefore, `ggmatplot` can be considered as a `ggplot` version of 
+[`matplot`](https://www.rdocumentation.org/packages/graphics/versions/3.6.2/topics/matplot).
+
+Similar to [`matplot`](https://www.rdocumentation.org/packages/graphics/versions/3.6.2/topics/matplot),
+`ggmatplot` plots a vector against the columns of a matrix, or the columns of 
+two matrices against each other, or a vector/matrix on its own. However, unlike
+[`matplot`](https://www.rdocumentation.org/packages/graphics/versions/3.6.2/topics/matplot),
+`ggmatplot` returns a `ggplot` object. Therefore, 
+[ggplot add ons](https://ggplot2.tidyverse.org/reference/index.html) such as 
+scales, faceting specifications, coordinate systems and themes can also be added
+on to the resultant `ggplot` object.
 

### 2021-10-30 README.Rmd
Added dependency on ggplot2 instead of using as an import

@@
 as [`ggplot2`](https://ggplot2.tidyverse.org/), it can be used as a 
-workaround for having to wrangle wide format data into long format for plotting 
-with [`ggplot2`](https://ggplot2.tidyverse.org/).
+workaround for having to wrangle wide format data into long format and creating
+simple plots using [`ggplot2`](https://ggplot2.tidyverse.org/).
 
@@
 library(ggmatplot)
-library(ggplot2)
 ```

### 2021-10-30 README.Rmd
Updated readme examples

@@
   out.width = "80%",
-  message = FALSE
+  message = FALSE,
+  warning = FALSE
 )
@@
 
+This example plots a vector x against each column of matrix z using the default `plot_type` of `ggmatplot()` - 'point'
+
 ```{r point}
-# matrix x
-x <- rnorm(100,sd=2)
+# vector x
+x <- c(rnorm(100, sd = 2))
 
-# matrix y
-y = x*0.5+rnorm(100,sd=1)
-fit.y = fitted(lm(y~x))
-z <- cbind(y, fit.y)
+# matrix z
+y <- x * 0.5 + rnorm(100, sd = 1)
+fit.y <- fitted(lm(y ~ x))
+z <- cbind(actual = y,
+           fitted = fit.y)
 
@@
 ```
+
+If two matrices with equal number of columns are used, the corresponding columns of the matrices will be plotted against each other.
+i.e.
+column 1 of matrix x will be plotted against column 1 of matrix y, column 2 of matrix x will be plotted against column 2 of matrix y, etc.
+
+This example uses the iris dataset, with matrices x and y as shown below.
+The Sepal.Length is plotted against Sepal.Width and the Petal.Length is plotted against Petal.Width.
+Therefore the groups 'Column 1' and 'Column 2' can be interpreted as 'Sepal' and 'Petal' respectively.
+
+```{r point-columns}
+x <- (iris[, c(1,3)])
+head(x,5)
+
+y <- (iris[, c(2,4)])
+head(y,5)
+
+ggmatplot(x,y)
+```
+
+This example creates a line plot of vector x against the columns of matrix y.
+The plot is therefore grouped by the columns of vector y.
+Custom colors are assigned to each group using the `color` parameter, and the limits of the y axis are updated using the `ylim` parameter.
+Further, a ggplot theme is added on to the resultant ggplot object.
 
@@
   plot_type = "line",
-          size = 1,
   color = c("blue", "purple"),
-          ylim = c(0,750)) +
+  ylim = c(0, 750)
+) +
   theme_minimal()
 ```
+
+This is a plot of the US personal expenditure over 5 categories and 5 years, and is a simple example of how wide format data can used with `ggmatplot()`.
+Note how the expenditure categories to be used on the x axis is used as vector x, and the expenditure values is used in wide format as matrix y - with its columns corresponding to the grouping structure.
+
+The plot uses the `plot_type` 'both', which is a combination of 'point' and 'line' plots. Its customized using `ggmatplot()` parameters and a `ggplot` theme as well.
 
@@
 
-# matrix x
+# vector x
 x <- rownames(USPersonalExpenditure)
 
-# matrix y
-y <- USPersonalExpenditure[, 1:5]
-
-ggmatplot(x, y,
+ggmatplot(x, USPersonalExpenditure,
   plot_type = "both",
@@
 
+Density plots only accept a single matrix or data frame and will group the plot based on its columns.
+The following density plot uses a two column matrix, and groups the plot by the two columns.
+While the default density estimate is represented in the measurement units of the data, an aesthetic mapping is added on to the ggplot object to scale the density estimate to a maximum of 1.
+
 ```{r density}
-iris_sub <- subset(iris, Species == "setosa")
-
 # matrix X
-x <- (iris_sub[,1:2])
-head(x,5)
+x <- (iris[, 1:2])
 
 ggmatplot(x, plot_type = "density") +
-  # scaling the density estimate to a maximum of 1
   aes(y = ..scaled..) +
@@
 
+Boxplots too accept only a single matrix or data frame, and uses its columns as individual groups.
+While `ggmatplot` plots are filled by default, the fill color can be made transparent by using `alpha = 0`.
+
+It is also worth noticing that `alpha` isn't a parameter defined in `ggmatplot()`, but can be used.
+This is because `ggmatplot` is built upon `ggplot2`, and each `plot_type` corresponds to a [`geom`](https://ggplot2.tidyverse.org/reference/index.html#section-geoms).
+Therefore, all valid parameters with the underlying [`ggplot2 geom`](https://ggplot2.tidyverse.org/reference/index.html#section-geoms) can be used with `ggmatplot()`
+
 ```{r boxplot}
-iris_sub <- subset(iris, Species == "setosa")
+# matrix X
+x <- (iris[, 1:4])
 
-# matrix X
-x <- (iris_sub[,1:4])
-head(x,5)
-
-ggmatplot(x, plot_type = "boxplot",
+ggmatplot(x,
+  plot_type = "boxplot",
   alpha = 0, # removing fill values
-          xlab = "", ylab = "")
+  xlab = "", ylab = ""
+)
 ```
 
+Violin plots too accepts a single matrix or data frame input, and behaves similar to density plots and boxplots.
+
+This plot updates the colors of the two groups using the `color` parameter, and it can be seen that the fill of the violin plots has been updated too.
+This is because updating either the `color` or `fill` parameter will automatically update the other, unless they are both defined simultaneously.
+
 ```{r violin}
-iris_sub <- subset(iris, Species == "setosa")
+# matrix X
+x <- (iris[, 1:2])
 
-# matrix X
-x <- (iris_sub[,1:2])
-head(x,5)
-
-ggmatplot(x, plot_type = "violin",
+ggmatplot(x,
+  plot_type = "violin",
   color = c("#00AFBB", "#E7B800"),
-          xlab = "", ylab = "")
+  xlab = "", ylab = ""
+)
 ```
 
+Similar to density, violin and box plots, histograms too accept a single matrix or data frame input and groups the plot using its columns.
+The histogram in the following example uses a matrix of 4 columns, and therefore groups the plots based on these 4 columns.
+The plot is also faceted by group.
+
+The `color` and `fill` parameters have been defined simultaneously on this plot.
+However, only a single `color` value is defined whereas the number of `fill` colors correspond to the number of groups.
+If a single value is defined it will be used over all groups, like the black line color is used across all groups in this example.
+
 ```{r histogram}
-iris_sub <- subset(iris, Species == "setosa")
+# matrix X
+x <- (iris[, 1:4])
 
-# matrix X
-x <- (iris_sub[,1:4])
-head(x,5)
-
-ggmatplot(x, plot_type = "histogram",
+ggmatplot(x,
+  plot_type = "histogram",
   xlab = "Group",
   color = "black",
-          fill = c("#F8766D", "#7CAE00", "#00BFC4", "#C77CFF"))
+  fill = c("#F8766D", "#7CAE00", "#00BFC4", "#C77CFF")
+) +
+  facet_wrap(~Group, scales = "free")
 ```

### 2021-11-01 README.Rmd
Updated example descriptions

@@
 
-This example plots a vector x against each column of matrix z using the default `plot_type` of `ggmatplot()` - 'point'
+The first example plots a vector x against each column of matrix z using the default `plot_type` of `ggmatplot()` - 'point'
 
@@
 
-This example uses the iris dataset, with matrices x and y as shown below.
-The Sepal.Length is plotted against Sepal.Width and the Petal.Length is plotted against Petal.Width.
+The next example uses the iris dataset, with matrices x and y as shown below.
+The `Sepal.Length` is plotted against `Sepal.Width` and the `Petal.Length` is plotted against `Petal.Width`.
 Therefore the groups 'Column 1' and 'Column 2' can be interpreted as 'Sepal' and 'Petal' respectively.
@@
 
-This example creates a line plot of vector x against the columns of matrix y.
-The plot is therefore grouped by the columns of vector y.
-Custom colors are assigned to each group using the `color` parameter, and the limits of the y axis are updated using the `ylim` parameter.
-Further, a ggplot theme is added on to the resultant ggplot object.
+The next example creates a line plot of vector x against the columns of matrix y.
+Although the lines would be represented using different colors by default, the `color` parameter allows custom colors to be assigned to them.
+The following plot assigns custom colors to the lines, and the limits of the y axis are updated using the `ylim` parameter. Further, a ggplot theme is added on to the resultant ggplot object.
 
@@
 
-This is a plot of the US personal expenditure over 5 categories and 5 years, and is a simple example of how wide format data can used with `ggmatplot()`.
+Next is plot of the US personal expenditure over 5 categories and 5 years, and is a simple example of how wide format data can used with `ggmatplot()`.
 Note how the expenditure categories to be used on the x axis is used as vector x, and the expenditure values is used in wide format as matrix y - with its columns corresponding to the grouping structure.

### 2021-11-04 README.Rmd
mild wording changes

@@
 wrangled into long format for plotting, which can be quite cumbersome when 
-creating simple plots. Therefore, the motivation for `ggmatplot` was to provide 
+creating simple plots. Therefore, the motivation for `ggmatplot` is to provide 
 a solution that allows [`ggplot2`](https://ggplot2.tidyverse.org/) to handle 
@@
 
-The first example plots a vector x against each column of matrix z using the default `plot_type` of `ggmatplot()` - 'point'
+The first example plots a vector against each column of matrix using the default `plot_type = "point"` of `ggmatplot()`.
+We consider a simple case that we have a covariate vector `x` and a matrix `z` with the response `y` and the fitted value `fit.y` as the two columns.
 
@@
 i.e.
-column 1 of matrix x will be plotted against column 1 of matrix y, column 2 of matrix x will be plotted against column 2 of matrix y, etc.
+column 1 of matrix `x` will be plotted against column 1 of matrix `y`, column 2 of matrix `x` will be plotted against column 2 of matrix `y`, etc.
 
-The next example uses the iris dataset, with matrices x and y as shown below.
+The next example uses the iris dataset, with matrices `x` and `y` as shown below.
 The `Sepal.Length` is plotted against `Sepal.Width` and the `Petal.Length` is plotted against `Petal.Width`.
 Therefore the groups 'Column 1' and 'Column 2' can be interpreted as 'Sepal' and 'Petal' respectively.
+Later we will provide examples on how to customize the legend name.
 

### 2021-11-04 README.Rmd
changes in readme

@@
 Therefore the groups 'Column 1' and 'Column 2' can be interpreted as 'Sepal' and 'Petal' respectively.
-Later we will provide examples on how to customize the legend name.
+To make the plot more meaningful, we can add the legend label and axis names by `legend_label`, `xlab` and `ylab`.
 
@@
 ggmatplot(x,y)
+ggmatplot(x,y, 
+          xlab = "Length",
+          ylab = "Width",
+          legend_label = c("Sepal","Petal"))
 ```
 
-The next example creates a line plot of vector x against the columns of matrix y.
+The next example creates a line plot of vector `x` against the columns of matrix `y` by using `plot_type = "line"`.
 Although the lines would be represented using different colors by default, the `color` parameter allows custom colors to be assigned to them.
@@
 
-Next is plot of the US personal expenditure over 5 categories and 5 years, and is a simple example of how wide format data can used with `ggmatplot()`.
+Next is plot of the US personal expenditure over 5 categories and 5 years, and is a simple example of how wide format data can be used with `ggmatplot()`.
 Note how the expenditure categories to be used on the x axis is used as vector x, and the expenditure values is used in wide format as matrix y - with its columns corresponding to the grouping structure.
 
-The plot uses the `plot_type` 'both', which is a combination of 'point' and 'line' plots. Its customized using `ggmatplot()` parameters and a `ggplot` theme as well.
+The plot specifies the plot type as `plot_type = "both"`, which is a combination of 'point' and 'line' plots. It customized using `ggmatplot()` parameters and a `ggplot` theme as well.
 
@@
 ```{r density}
-# matrix X
+# matrix x
 x <- (iris[, 1:2])
@@
 ggmatplot(x, plot_type = "density") +
-  aes(y = ..scaled..) +
+  aes(y = stat(scaled)) +
   theme_bw()
@@
 
-Boxplots too accept only a single matrix or data frame, and uses its columns as individual groups.
+Boxplots accept only a single matrix or data frame as well, and uses its columns as individual groups.
 While `ggmatplot` plots are filled by default, the fill color can be made transparent by using `alpha = 0`.
@@
 ```{r boxplot}
-# matrix X
+# matrix x
 x <- (iris[, 1:4])
@@
 ```{r violin}
-# matrix X
+# matrix x
 x <- (iris[, 1:2])
@@
 ) +
-  facet_wrap(~Group, scales = "free")
+  facet_wrap( ~ Group, scales = "free")
 ```

### 2021-11-07 R/ggmatplot.R
Added ecdf and dotplot

@@ -76,7 +76,7 @@ ggmatplot <- function(x, y, color = NULL, shape = NULL, linetype = NULL,
   }
   if (!missing(x) & !missing(y)) {
     # only single matrix input allowed for the following plot types
-    if (plot_type %in% c("density", "histogram", "boxplot", "violin")) {
+    if (plot_type %in% c("density", "histogram", "boxplot", "violin", "dotplot", "ecdf")) {
       stop("This plot type only uses a single matrix input", call. = FALSE)
     }
     data.list <- matclean(x = x, y = y)
@@ -201,7 +201,34 @@ ggmatplot <- function(x, y, color = NULL, shape = NULL, linetype = NULL,
           params
         )
       )
+  } else if (plot_type == "dotplot") {
+    params$alpha <- if (is.null(params$alpha)) 0.5 else params$alpha
+    p <- p +
+      do.call(
+        "geom_dotplot",
+        c(
+          list(mapping = aes(
+            x = Group,
+            y = .data[[yname]]
+          ),
+          binaxis = "y",
+          stackdir = "center"),
+          params
+        )
+      )
+  } else if (plot_type == "ecdf") {
+    p <- p +
+      do.call(
+        "stat_ecdf",
+        c(
+          list(mapping = aes(
+            x = .data[[yname]]
+          )),
+          params
+        )
+      )
   }
+
 
   # number of unique groups
   numGroups <- length(unique(data$Group))
@@ -273,7 +300,7 @@ ggmatplot <- function(x, y, color = NULL, shape = NULL, linetype = NULL,
 
   if (!is.null(linetype)) {
     # linetype parameter is invalid for the following plot types
-    if (!plot_type %in% c("point")) {
+    if (!plot_type %in% c("point", "dotplot")) {
       linetype <- validateNumParams(linetype, numGroups)
     } else {
       warning(paste0("linetype is an invalid parameter for plot type: ", plot_type), call. = FALSE)

### 2021-11-07 README.Rmd
changes in readme
Updated readme with ecdf and dotplot examples

@@
 Therefore the groups 'Column 1' and 'Column 2' can be interpreted as 'Sepal' and 'Petal' respectively.
-To make the plot more meaningful, we can add the legend label and axis names by `legend_label`, `xlab` and `ylab`.
+To make the plot more meaningful, we can further add the legend label and axis names by `legend_label`, `xlab` and `ylab`.
 
@@
 
-Similar to density, violin and box plots, histograms too accept a single matrix or data frame input and groups the plot using its columns.
+Dotplots too accept a single matrix input and plots the distribution of each of its columns.
+
+The next example uses the `plot_type = "dotplot"` to visualise the distribution of the data, and also customizes the position of the legend using a `ggplot` theme as well.
+
+```{r dotplot}
+# matrix x
+x <- (iris[, 1:2])
+
+ggmatplot(x,
+  plot_type = "dotplot",
+  color = c("#00AFBB", "#E7B800"),
+  xlab = "", ylab = ""
+) +
+  theme(legend.position = "bottom")
+```
+
+Similar to density, violin, dotplots, and box plots, histograms too accept a single matrix or data frame input and groups the plot using its columns.
 The histogram in the following example uses a matrix of 4 columns, and therefore groups the plots based on these 4 columns.
@@
 ```{r histogram}
-# matrix X
+# matrix x
 x <- (iris[, 1:4])
@@
 ```
+
+The next example is of the `plot_type = ecdf`, and also uses a single matrix input to plot out the empirical cumulative distributions of the columns of the matrix individually.
+
+```{r ecdf}
+# matrix x
+x <- (iris[, 1:4])
+
+ggmatplot(x,
+  plot_type = "ecdf",
+  xlab = "Group",
+  size = 1
+) +
+  theme_minimal()
+```

### 2021-11-07 tests/testthat/test_ggmatplot_baseplots.R
Updated ecdf and dotplot tests

@@ -4,43 +4,46 @@ x <- iris[, 1]
 y <- iris[, 3:4]
 
 test_that("scatterplot draws correctly", {
-  p <- ggmatplot(x, y, plot_type = "point")
-  expect_doppelganger("basic scatterplot", p)
+  expect_doppelganger("basic scatterplot", ggmatplot(x, y, plot_type = "point"))
 })
 
 test_that("line plot draws correctly", {
-  p <- ggmatplot(x, y, plot_type = "line")
-  expect_doppelganger("basic line plot", p)
+  expect_doppelganger("basic line plot", ggmatplot(x, y, plot_type = "line"))
 })
 
 test_that("point + line plot draws correctly", {
-  p <- ggmatplot(x, y, plot_type = "both")
-  expect_doppelganger("basic point + line plot", p)
+  expect_doppelganger("basic point + line plot", ggmatplot(x, y, plot_type = "both"))
 })
 
 test_that("density plot draws correctly", {
-  p <- ggmatplot(y, plot_type = "density")
-  expect_doppelganger("basic density plot", p)
+  expect_doppelganger("basic density plot", ggmatplot(y, plot_type = "density"))
 })
 
 test_that("histogram draws correctly", {
-  p <- ggmatplot(y, plot_type = "histogram")
-  expect_doppelganger("basic histogram", p)
+  expect_doppelganger("basic histogram", ggmatplot(y, plot_type = "histogram"))
 })
 
 test_that("violin plot draws correctly", {
-  p <- ggmatplot(y, plot_type = "violin")
-  expect_doppelganger("basic violin plot", p)
+  expect_doppelganger("basic violin plot", ggmatplot(y, plot_type = "violin"))
 })
 
 test_that("boxplot draws correctly", {
-  p <- ggmatplot(y, plot_type = "boxplot")
-  expect_doppelganger("basic boxplot", p)
+  expect_doppelganger("basic boxplot", ggmatplot(y, plot_type = "boxplot"))
 })
 
-test_that("density, histogram, violin and boxplot do not accept multiple matrices", {
+test_that("dotplot draws correctly", {
+  expect_doppelganger("basic dotplot", ggmatplot(y, plot_type = "dotplot"))
+})
+
+test_that("ecdf draws correctly", {
+  expect_doppelganger("basic ecdf", ggmatplot(y, plot_type = "ecdf"))
+})
+
+test_that("density, histogram, ecdf, dotplot, violin and boxplot do not accept multiple matrices", {
[…]

### 2021-11-07 README.Rmd
formatting code

@@
 fit.y <- fitted(lm(y ~ x))
-z <- cbind(actual = y,
-           fitted = fit.y)
+z <- cbind(
+  actual = y,
+  fitted = fit.y
+)
 
@@
 ```{r point-columns}
-x <- (iris[, c(1,3)])
-head(x,5)
+x <- (iris[, c(1, 3)])
+head(x, 5)
 
-y <- (iris[, c(2,4)])
-head(y,5)
+y <- (iris[, c(2, 4)])
+head(y, 5)
 
-ggmatplot(x,y)
-ggmatplot(x,y, 
+ggmatplot(x, y)
+ggmatplot(x, y,
   xlab = "Length",
   ylab = "Width",
-          legend_label = c("Sepal","Petal"))
+  legend_label = c("Sepal", "Petal")
+)
 ```
@@
 ) +
-  facet_wrap( ~ Group, scales = "free")
+  facet_wrap(~Group, scales = "free")
 ```

### 2021-11-07 R/ggmatplot.R
Updates to DESCRIPTION

@@ -61,11 +61,11 @@
 #' ggmatplot(iris_sub[, c(1, 3)], iris_sub[, c(2, 4)])
 #' # Modify legend label and axis
 #' ggmatplot(iris_sub[, c(1, 3)], iris_sub[, c(2, 4)], shape = c(4, 6), legend_label = c("Sepal", "Petal"), legend_title = "", xlab = "Length", ylab = "Width")
-ggmatplot <- function(x, y, color = NULL, shape = NULL, linetype = NULL,
+ggmatplot <- function(x, y, plot_type = "point", color = NULL, shape = NULL, linetype = NULL,
                       fill = NULL, xlim = c(NA, NA), ylim = c(NA, NA),
                       log = NULL, main = NULL, xlab = NULL, ylab = NULL,
                       legend_label = NULL, legend_title = NULL,
-                      plot_type = "point", asp = NA, ...) {
+                       asp = NA, ...) {
 
   # valid plot types
   if (!plot_type %in% c(

### 2021-11-16 R/errorplotstats.R
Added errorplot

@@ -1,0 +1,63 @@
+#' Function to compute descriptive statistics to be used for errorplots
+#'
+#' @param data Resultant data frame from matclean()
+#' @param desc_stat Statistics to calculate. Either of `mean_se`, `mean_sd`, `mean_range`, `median_iqr`, `median_range`
+#'
+#' @return A data frame with descriptive statistics for each group.
+
+#' @noRd
+#'
+#' @examples
+#' # Define a data set
+#' iris_sub <- subset(iris, Species == "setosa")[1:4]
+#' # Use ggmatplot::matclean to pivot the dataframe into long format
+#' data <- matclean(iris_sub)$data
+#' errorplotstats(data, desc_stat = "mean_se")
+errorplotstats <- function(data, desc_stat = "") {
+  if (desc_stat == "mean_se") {
+    errorplot_data <- do.call(rbind, by(data, list(data$Group), function(x) {
+      c(
+        y = mean(x$x, na.rm = TRUE),
+        ymin = mean(x$x, na.rm = TRUE) - sd(x$x, na.rm = TRUE) / sqrt(nrow(x)),
+        ymax = mean(x$x, na.rm = TRUE) + sd(x$x, na.rm = TRUE) / sqrt(nrow(x))
+      )
+    }))
+  } else if (desc_stat == "mean_sd") {
+    errorplot_data <- do.call(rbind, by(data, list(data$data$Group), function(x) {
+      c(
+        y = mean(x$x, na.rm = TRUE),
+        ymin = mean(x$x, na.rm = TRUE) - sd(x$x, na.rm = TRUE),
+        ymax = mean(x$x, na.rm = TRUE) + sd(x$x, na.rm = TRUE)
+      )
+    }))
+  } else if (desc_stat == "mean_range") {
+    errorplot_data <- do.call(rbind, by(data, list(data$Group), function(x) {
+      c(
+        y = mean(x$x, na.rm = TRUE),
+        ymin = min(x$x, na.rm = TRUE),
+        ymax = max(x$x, na.rm = TRUE)
+      )
+    }))
+  } else if (desc_stat == "median_iqr") {
+    errorplot_data <- do.call(rbind, by(data, list(data$Group), function(x) {
+      c(
+        y = median(x$x, na.rm = TRUE),
+        ymin = quantile(x$x, probs = 0.25, na.rm = TRUE, names = FALSE),
[…]

### 2021-11-16 R/ggmatplot.R

@@ -61,11 +61,11 @@
 #' ggmatplot(iris_sub[, c(1, 3)], iris_sub[, c(2, 4)])
 #' # Modify legend label and axis
 #' ggmatplot(iris_sub[, c(1, 3)], iris_sub[, c(2, 4)], shape = c(4, 6), legend_label = c("Sepal", "Petal"), legend_title = "", xlab = "Length", ylab = "Width")
-ggmatplot <- function(x, y, plot_type = "point", color = NULL, shape = NULL, linetype = NULL,
-                      fill = NULL, xlim = c(NA, NA), ylim = c(NA, NA),
-                      log = NULL, main = NULL, xlab = NULL, ylab = NULL,
-                      legend_label = NULL, legend_title = NULL,
-                       asp = NA, ...) {
+ggmatplot <- function(x, y, plot_type = "point", color = NULL, fill = NULL,
+                      shape = NULL, linetype = NULL, xlim = c(NA, NA),
+                      ylim = c(NA, NA), log = NULL, main = NULL,
+                      xlab = NULL, ylab = NULL, legend_label = NULL,
+                      legend_title = NULL, desc_stat = "mean_se", asp = NA, ...) {
 
   # valid plot types
   if (!plot_type %in% c(
@@ -88,7 +88,21 @@ ggmatplot <- function(x, y, plot_type = "point", color = NULL, shape = NULL, lin
 
   params <- list(...)
 
-  data <- data.list$data
+  # valid plot types for desc_stat parameter
+  if (desc_stat != "mean_se" & plot_type != "errorplot") {
+    warning(paste0("desc_stat is an invalid parameter for plot type: ", plot_type), call. = FALSE)
+  }
+
+  if (plot_type == "errorplot") {
+    if (desc_stat %in% c("mean_se", "mean_sd", "mean_range", "median_iqr", "median_range")) {
+      data <- errorplotstats(data.list$data, desc_stat)
+    } else {
+      stop("desc_stat can not take this value", call. = FALSE)
+    }
+  } else {
+    data <- data.list$data
+  }
+
   xname <- data.list$xname
   yname <- data.list$yname
 
@@ -207,12 +221,14 @@ ggmatplot <- function(x, y, plot_type = "point", color = NULL, shape = NULL, lin
       do.call(
         "geom_dotplot",
         c(
-          list(mapping = aes(
-            x = Group,
-            y = .data[[yname]]
+          list(
+            mapping = aes(
+              x = Group,
+              y = .data[[yname]]
+            ),
[…]

### 2021-11-16 based on ggpubr/getoption.R
Deleting ggpubr examples

@@ -1,15 +1,0 @@
-getoption <- function(params, plot.type){
-  
-  option <- list()
-  
-  if(plot.type=="histogram")
-    allowed_options <- c("data","x","color","fill","linetype","alpha", "bins", "binwidth", "add", "add.params", "rug","add_density","position")
-  
-  for (key in names(params)) {
-    value <- params[[key]]
-    if (key %in% allowed_options) {
-      option[[key]] <- value
-    }
-  }
-  return(option)
-}

### 2021-11-16 based on ggpubr/ggmatclean.R

@@ -1,50 +1,0 @@
-ggmatclean <- function (x, y){
-
-  if(is.list(x)){
-    x <- ggmatlist(x)
-  }
-
-  x <- data.frame(x)
-  ncolx <- ncol(x)
-  
-  if (!missing(y)){
-    ncoly <- ncol(as.data.frame(y))
-    y <- data.frame(y)
-  }else{
-    ncoly<-1
-  }
-  
-   if (missing(y)){
-     data <- mutate(x, Observation_number = 1:n()) %>% gather(key = "Group", value = "yy", -(ncolx+1))
-     xname <- "Observation_number"
-     yname <- "yy"
-     group <- "Group"
-   }else if(ncolx>ncoly&ncoly==1){
-     data <- data.frame(x,y)
-     ncol <- ncol(data)
-     yname <- colnames(data)[ncol]
-     data <- mutate(data, Observation_number = 1:n()) %>% gather(key = "Group", value = "x", -c(ncol,(ncol+1)))
-     xname <- "x"
-     yname <- colnames(y)
-     group <- "Group"
-   }else if(ncoly>ncolx&ncolx==1){
-     data <- data.frame(x,y)
-     ncol <- ncol(data)
-     data <- mutate(data, Observation_number = 1:n()) %>% gather(key = "Group", value = "yy", -c(1,(ncol+1)))
-     xname <- colnames(x)
-     yname <- "yy"
-     group <- "Group"
-   }else if(ncolx==ncoly){
-     colnames(x) = colnames(y) = paste0("Column ", 1:ncolx)
-     x <- gather(x, key = "Group", value = "x") %>% mutate(Observation_number = 1:n())
-     y <- gather(y, key = "Group", value = "yy") %>% mutate(Observation_number = 1:n())
-     data <- full_join(x, y, by = "Observation_number") %>% rename(Group = Group.x) %>% select(-Group.y)
-     xname <- "x"
-     yname <- "yy"
-     group <- "Group"
-   }else{
-     stop("`x`` and `y` must have only 1 or the same number of columns", call. = FALSE)
-   }
-  
-  return(list(data=data, xname=xname, yname=yname, group=group, ncolx=ncolx, ncoly=ncoly))
-}

### 2021-11-16 based on ggpubr/ggmatlist.R

@@ -1,5 +1,0 @@
-ggmatlist <- function (x){
-  max.length <- max(sapply(x, length))
-  x <- lapply(x, function(v) { c(v, rep(NA, max.length-length(v)))})
-  x<- do.call(cbind, x)
-}

### 2021-11-16 based on ggpubr/ggmatplot.R

@@ -1,137 +1,0 @@
-ggmatplot <- function (x, y, color = NULL, shape = NULL, linetype = NULL, fill=NULL,
-                       xlim = c(NA, NA), ylim = c(NA, NA), log = "",
-                       main = NULL, xlab = NULL, ylab = NULL,
-                       legend_label =  NULL, legend_title = NULL,
-                       plot.type = "point", asp = NA){
-
-  # caller_env <- parent.frame()
-
-  if (! plot.type %in% c("point", "line", "both", "density", "histogram", "boxplot", "dotplot","errorplot",
-                         "violin","ecdf") ) 
-    stop("`plot.type` can not be this character", call. = FALSE)
-  
-  if (!missing(x)&!missing(y)){
-    data.list <- ggmatclean (x=x, y=y)
-  } else if (!missing(x)){
-    data.list <- ggmatclean (x=x)
-  } else{
-    stop("No data have been put into the ggmatplot function", call. = FALSE)
-  }
-  
-  data <- data.list$data %>% drop_na()
-  xname <- data.list$xname
-  yname <- data.list$yname
-  ncolx <- data.list$ncolx
-  ncoly <- data.list$ncoly
-  group <- data.list$group 
-  
-  dcolor <- data.list$group 
-  dfill <- data.list$group
-  dshape <- data.list$group
-  dlinetype <- data.list$group
-
-  if(missing(y)& ncolx==1){
-    dcolor = "black" 
-    dfill = "lightgray"
-    dshape = 19
-    dlinetype = 1
-  }
-  
-  
-  if(plot.type=="point"){
-    p <-ggscatter(data = data, x = xname, y = yname, color = dcolor, shape=dshape)
-  }
-  if(plot.type=="line"){
-    p <-ggline(data = data, x = xname, y = yname, color= dcolor, linetype = dlinetype, plot_type = "l")
-  }
-  if(plot.type=="both"){
-    p <-ggline(data = data, x = xname, y = yname, color= dcolor, linetype = dlinetype, shape=dshape)
-  }
-  if(plot.type=="density"){
-    p <-ggdensity(data = data, x = yname, color = dcolor, fill = dfill)+ xlab("x")
-  }
-  if(plot.type=="histogram"){
-    p <-gghistogram(data = data, x = yname, color = dcolor, fill = dfill)+ xlab("x")
-  }
[…]

### 2021-11-16 based on ggpubr/ggmatplot1.R

@@ -1,146 +1,0 @@
-ggmatplot1 <- function (x, y, color = NULL, shape = NULL, linetype = NULL, fill=NULL,
-                       xlim = c(NA, NA), ylim = c(NA, NA), log = "",
-                       main = NULL, xlab = NULL, ylab = NULL,
-                       legend_label =  NULL, legend_title = NULL,
-                       plot.type = "point", asp = NA, ...){
-  
-  # caller_env <- parent.frame()
-  
-  if (! plot.type %in% c("point", "line", "both", "density", "histogram", "boxplot", "dotplot","errorplot",
-                         "violin","ecdf") ) 
-    stop("`plot.type` can not be this character", call. = FALSE)
-  
-  if (!missing(x)&!missing(y)){
-    data.list <- ggmatclean (x=x, y=y)
-  } else if (!missing(x)){
-    data.list <- ggmatclean (x=x)
-  } else{
-    stop("No data have been put into the ggmatplot function", call. = FALSE)
-  }
-  
-  data <- data.list$data %>% drop_na()
-  xname <- data.list$xname
-  yname <- data.list$yname
-  ncolx <- data.list$ncolx
-  ncoly <- data.list$ncoly
-  group <- data.list$group 
-  
-  dcolor <- data.list$group 
-  dfill <- data.list$group
-  dshape <- data.list$group
-  dlinetype <- data.list$group
-  
-  if(missing(y)& ncolx==1){
-    dcolor = "black" 
-    dfill = "lightgray"
-    dshape = 19
-    dlinetype = 1
-  }
-  
-  params <- list(...)
-  params$data <- data
-  params$color <- dcolor
-  params$fill<- dfill
-  params$shape <- dshape
-  params$dlinetype <- dlinetype
-  
-  if(plot.type=="point"){
-    p <-ggscatter(data = data, x = xname, y = yname, color = dcolor, shape=dshape)
-  }
-  if(plot.type=="line"){
-    p <-ggline(data = data, x = xname, y = yname, color= dcolor, linetype = dlinetype, plot_type = "l")
-  }
-  if(plot.type=="both"){
-    p <-ggline(data = data, x = xname, y = yname, color= dcolor, linetype = dlinetype, shape=dshape, ...)
-  }
-  if(plot.type=="density"){
[…]

### 2021-11-16 R/ggmatplot.R
Updated ggmatplot argument documentation

@@ -1,15 +1,21 @@
 #' ggmatplot
 #'
 #'
-#' `ggmatplot` is a quick and easy way of plotting the columns of two matrices
-#' against each other.
+#' `ggmatplot` is a quick and easy way of plotting the columns of two matrices or
+#'  data frames against each other using [`ggplot2`](https://ggplot2.tidyverse.org/).
 #'
-#' @param x,y Vectors or matrices of data for plotting.
+#' @param x,y Vectors or matrices of data.
 #'
 #' * The number of rows of `x` and `y` should be the same.
+#' * Either `x` or `y` should be a vector, unless the number of columns of `x` and
+#' `y` are the same.
 #' * Missing values (NAs) are allowed.
 #' * If either `x` or `y` is missing, the other is used as `y` and a vector of
 #' `1:n` is used as `x`.
+#'
+#' @param plot_type A string specifying the type of plot. Possible plot types are
+#' `point`, `line`, `both`(point + line), `density`, `histogram`, `boxplot`,
+#' `dotplot`, `errorplot`, `violin`, and `ecdf`. Default plot_type is `point`.
 #'
 #' @param color,fill Vectors of colors. Defining only one of them will update
 #' both `color` and `fill` aesthetics of the plot by default, unless they are
@@ -19,6 +25,11 @@
 #' matrices `x` or `y`, and will correspond to each of those columns.
 #' * If only a single color is given, the same color will be used for all columns.
 #'
+#' @param main,xlab,ylab,legend_title Strings to update plot title, x axis label,
+#'  y axis label and legend title respectively.
+#'
+#' @param legend_label A vector of strings, to rename the legend labels.
+#'
 #' @param shape,linetype A vector of shapes or line types respectively.
 #'
 #' * The number of shapes/line types should match the higher number of columns of
@@ -26,25 +37,21 @@
 #' * If only a single shape/line type is given, the same shape/line type will be
 #' used for all columns.
 #'
-#' @param xlim,ylim Ranges of x and y axes. Each of them should be a two element
-#' vector specifying the lower and upper limits of the scale. If the larger
-#' value is given first, the scale will be reversed. If one of the limits is
-#' given as `NA`, the corresponding limit from the range of data will be used.
+#' @param xlim,ylim Ranges of x and y axes.
+#'
+#' * Each of them should be a two element vector specifying the lower and upper
[…]

### 2021-11-18 R/ggmatplot.R
Added list of underlying ggplot layers on to reference

@@ -40,9 +40,10 @@
 #' @param xlim,ylim Ranges of x and y axes.
 #'
 #' * Each of them should be a two element vector specifying the lower and upper
-#' limits of the scale. If the larger value is given first, the scale will be
-#' reversed. If one of the limits is given as `NA`, the corresponding limit from
-#'  the range of data will be used.
+#' limits of the scale.
+#' * If the larger value is given first, the scale will be reversed. If one of
+#' the limits is given as `NA`, the corresponding limit from the range of data
+#' will be used.
 #'
 #' @param log A string defining which axes to transform into a log scale.
 #' (`x`, `y` or `xy`)
@@ -53,13 +54,30 @@
 #' `errorplot`. Possible values are `mean_se`, `mean_sd`, `mean_range`,
 #'  `median_iqr` and `median_range`. Default desc_stat is `mean_se`.
 #'
-#' @param ... Other arguments passed on to the plot. They can be used to set an
-#' aesthetic to a fixed value, like `alpha = 0.4` or `size = 2`. They may also
-#' be parameters passed on to the corresponding [geoms](https://ggplot2.tidyverse.org/reference/).
+#' @param ... Other arguments passed on to the plot. Possible arguments are those
+#' that can be passed on to the [underlying ggplot layer](#plot-types).
 #'
 #' @import ggplot2
 #' @export
 #' @md
+#'
+#' @section Plot Types:
+#'
+#' `ggmatplot`plots are built upon `ggplot2 layers`. The following is a list of
+#' `ggmatplot` plot types, along with their underlying
+#' [`ggplot geoms`](https://ggplot2.tidyverse.org/reference/index.html#section-geoms)
+#' or [`stats`](https://ggplot2.tidyverse.org/reference/index.html#section-stats).
+#'
+#' * \strong{point} \code{\link[ggplot2]{geom_point}}
+#' * \strong{line} \code{\link[ggplot2]{geom_line}}
+#' * \strong{both} \code{\link[ggplot2]{geom_point}} + \code{\link[ggplot2]{geom_line}}
+#' * \strong{density} \code{\link[ggplot2]{geom_density}}
+#' * \strong{histogram} \code{\link[ggplot2]{geom_histogram}}
+#' * \strong{boxplot} \code{\link[ggplot2]{geom_boxplot}}
+#' * \strong{dotplot} \code{\link[ggplot2]{geom_dotplot}}
+#' * \strong{errorplot} \code{\link[ggplot2]{geom_pointrange}}
[…]

### 2021-11-18 R/errorplotstats.R
Created internal descriptive stats calculation functions

@@ -18,16 +18,16 @@ errorplotstats <- function(data, desc_stat = "") {
     errorplot_data <- do.call(rbind, by(data, list(data$Group), function(x) {
       c(
         y = mean(x$x, na.rm = TRUE),
-        ymin = mean(x$x, na.rm = TRUE) - sd(x$x, na.rm = TRUE) / sqrt(nrow(x)),
-        ymax = mean(x$x, na.rm = TRUE) + sd(x$x, na.rm = TRUE) / sqrt(nrow(x))
+        ymin = mean(x$x, na.rm = TRUE) - (standard_dev(x$x[!is.na(x$x)]) / sqrt(nrow(x))),
+        ymax = mean(x$x, na.rm = TRUE) + (standard_dev(x$x[!is.na(x$x)]) / sqrt(nrow(x)))
       )
     }))
   } else if (desc_stat == "mean_sd") {
-    errorplot_data <- do.call(rbind, by(data, list(data$data$Group), function(x) {
+    errorplot_data <- do.call(rbind, by(data, list(data$Group), function(x) {
       c(
         y = mean(x$x, na.rm = TRUE),
-        ymin = mean(x$x, na.rm = TRUE) - sd(x$x, na.rm = TRUE),
-        ymax = mean(x$x, na.rm = TRUE) + sd(x$x, na.rm = TRUE)
+        ymin = mean(x$x, na.rm = TRUE) - standard_dev(x$x[!is.na(x$x)]),
+        ymax = mean(x$x, na.rm = TRUE) + standard_dev(x$x[!is.na(x$x)])
       )
     }))
   } else if (desc_stat == "mean_range") {
@@ -41,15 +41,15 @@ errorplotstats <- function(data, desc_stat = "") {
   } else if (desc_stat == "median_iqr") {
     errorplot_data <- do.call(rbind, by(data, list(data$Group), function(x) {
       c(
-        y = median(x$x, na.rm = TRUE),
-        ymin = quantile(x$x, probs = 0.25, na.rm = TRUE, names = FALSE),
-        ymax = quantile(x$x, probs = 0.75, na.rm = TRUE, names = FALSE)
+        y = get_median(x$x[!is.na(x$x)]),
+        ymin = get_quantile(x$x[!is.na(x$x)], quantile = 0.25),
+        ymax = get_quantile(x$x[!is.na(x$x)], quantile = 0.75)
       )
     }))
   } else if (desc_stat == "median_range") {
[…]

### 2021-11-18 R/ggmatplot.R

@@ -85,12 +85,16 @@
 #' iris_sub <- subset(iris, Species == "setosa")
 #' ggmatplot(iris_sub[, c(1, 3)], iris_sub[, c(2, 4)])
 #' # Modify legend label and axis
-#' ggmatplot(iris_sub[, c(1, 3)], iris_sub[, c(2, 4)], shape = c(4, 6), legend_label = c("Sepal", "Petal"), legend_title = "", xlab = "Length", ylab = "Width")
+#' ggmatplot(iris_sub[, c(1, 3)], iris_sub[, c(2, 4)], shape = c(4, 6),
+#' legend_label = c("Sepal", "Petal"), legend_title = "", xlab = "Length", ylab = "Width")
 ggmatplot <- function(x, y, plot_type = "point", color = NULL, fill = NULL,
                       shape = NULL, linetype = NULL, xlim = c(NA, NA),
                       ylim = c(NA, NA), log = NULL, main = NULL,
                       xlab = NULL, ylab = NULL, legend_label = NULL,
                       legend_title = NULL, desc_stat = "mean_se", asp = NA, ...) {
+
+  # binding global variables to objects
+  Group <- ymin <- ymax <- NULL
 
   # valid plot types
   if (!plot_type %in% c(

### 2021-11-18 README.Rmd
Updated documentation with errorplot example

@@
 Although the lines would be represented using different colors by default, the `color` parameter allows custom colors to be assigned to them.
+
 The following plot assigns custom colors to the lines, and the limits of the y axis are updated using the `ylim` parameter. Further, a ggplot theme is added on to the resultant ggplot object.
@@
 It is also worth noticing that `alpha` isn't a parameter defined in `ggmatplot()`, but can be used.
-This is because `ggmatplot` is built upon `ggplot2`, and each `plot_type` corresponds to a [`geom`](https://ggplot2.tidyverse.org/reference/index.html#section-geoms).
-Therefore, all valid parameters with the underlying [`ggplot2 geom`](https://ggplot2.tidyverse.org/reference/index.html#section-geoms) can be used with `ggmatplot()`
+This is because `ggmatplot` is built upon `ggplot2`, and each `plot_type` corresponds to a [`geom`](https://ggplot2.tidyverse.org/reference/index.html#section-geoms) as listed [here](./reference/ggmatplot.html#plot-types).
+Therefore, all valid parameters with the underlying [`ggplot2 geom`](https://ggplot2.tidyverse.org/reference/index.html#section-geoms) can be used with `ggmatplot()`.
 
@@
 ```
+
+Error plots also accept only a single matrix input, and plots out error bars for each column of the matrix.
+The `desc_stat` parameter of `ggmatplot()` can be used to define what the mid point and error bars of the plot should represent.
+
+The next example, plots out an `errorplot` using the medians and interquartile ranges of each variable.
+
+```{r errorplot}
+# matrix x
+x <- (iris[, 1:4])
+
+ggmatplot(x,
+  plot_type = "errorplot",
+  desc_stat = "median_iqr",
+  xlab = "Group",
+  size = 1
+) +
+  theme_minimal()
+```

### 2021-11-18 R/ggmatplot.R

@@ -55,7 +55,7 @@
 #'  `median_iqr` and `median_range`. Default desc_stat is `mean_se`.
 #'
 #' @param ... Other arguments passed on to the plot. Possible arguments are those
-#' that can be passed on to the [underlying ggplot layer](#plot-types).
+#' that can be passed on to the [underlying ggplot layers](#plot-types).
 #'
 #' @import ggplot2
 #' @export

### 2021-11-18 tests/testthat/test_ggmatplot_baseplots.R
Added errorplot tests

@@ -39,11 +39,16 @@ test_that("ecdf draws correctly", {
   expect_doppelganger("basic ecdf", ggmatplot(y, plot_type = "ecdf"))
 })
 
-test_that("density, histogram, ecdf, dotplot, violin and boxplot do not accept multiple matrices", {
+test_that("errorplot draws correctly", {
+  expect_doppelganger("basic errorplot", ggmatplot(y, plot_type = "errorplot"))
+})
+
+test_that("density, histogram, ecdf, dotplot, violin, errorplot and boxplot do not accept multiple matrices", {
   expect_error(ggmatplot(x, y, plot_type = "density"), "This plot type only uses a single matrix input")
   expect_error(ggmatplot(x, y, plot_type = "histogram"), "This plot type only uses a single matrix input")
   expect_error(ggmatplot(x, y, plot_type = "ecdf"), "This plot type only uses a single matrix input")
   expect_error(ggmatplot(x, y, plot_type = "dotplot"), "This plot type only uses a single matrix input")
   expect_error(ggmatplot(x, y, plot_type = "violin"), "This plot type only uses a single matrix input")
+  expect_error(ggmatplot(x, y, plot_type = "errorplot"), "This plot type only uses a single matrix input")
   expect_error(ggmatplot(x, y, plot_type = "boxplot"), "This plot type only uses a single matrix input")
 })

### 2021-11-18 tests/testthat/test_ggmatplot_parameters.R

@@ -68,7 +68,7 @@ test_that("plot axis labels are updated based on xlab and ylab paramters", {
   expect_equal(ggmatplot(y, plot_type = "point", ylab="ylab test")$labels$y, "ylab test")
 })
 
-test_that("invalid plot types are no allowed", {
+test_that("invalid plot types are not allowed", {
   expect_error(ggmatplot(x,y, plot_type = "scatterplot"), "plot_type can not take this value")
 })
 
@@ -84,6 +84,11 @@ test_that("legend title is updated", {
 test_that("plot is resized by aspect ratio(asp)", {
   expect_doppelganger("plot resized by aspect ratio", ggmatplot(x,y, asp = 0.5))
 })
+
+test_that("errorplot updates using the defined desc_stat", {
+  expect_doppelganger("errorplot with medians and range errorbars", ggmatplot(y, plot_type = "errorplot", desc_stat = "median_range"))
+})
+
 
 test_that("invalid parameter values throw errors", {
   # color values < number of groups
@@ -109,6 +114,8 @@ test_that("invalid parameter values throw errors", {
   expect_error(ggmatplot(y, plot_type = "line", xlim = c(1,2,5)), "xlim must be a two element vector")
   # invalid log value
   expect_error(ggmatplot(y, plot_type = "point", log = 1), "invalid log value provided")
+  # invalid des_stat value
+  expect_error(ggmatplot(y, plot_type = "errorplot", desc_stat = "mean_sed"), "desc_stat can not take this value")
 })
 
 test_that("unknown parameter values for plots throw errors", {
@@ -116,4 +123,6 @@ test_that("unknown parameter values for plots throw errors", {
   expect_warning(ggmatplot(y, plot_type = "density", shape = c(15, 12, 13)), "shape is an invalid parameter for plot type: density")
   # linetypes for plot types that don't require linetypes
   expect_warning(ggmatplot(x, y, plot_type = "point", linetype = c(15, 12, 13)), "linetype is an invalid parameter for plot type: point")
+  # desc_stat for plot types that don't require descriptive statistics
+  expect_warning(ggmatplot(x, plot_type = "density", desc_stat = "median_iqr"), "desc_stat is an invalid parameter for plot type: density")
 })

### 2021-11-21 tests/testthat.R
Added errorplot stats tests

@@ -1,5 +1,6 @@
 library(testthat)
-library(tibble)
 library(ggmatplot)
+library(tibble)
+library(stats)
 
 test_check("ggmatplot")

### 2021-11-21 R/errorplotstats.R
Added dependency on stats package

@@ -4,7 +4,11 @@
 #' @param desc_stat Statistics to calculate. Either of `mean_se`, `mean_sd`, `mean_range`, `median_iqr`, `median_range`
 #'
 #' @return A data frame with descriptive statistics for each group.
-
+#'
+#' @importFrom stats sd
+#' @importFrom stats median
+#' @importFrom stats quantile
+#'
 #' @noRd
 #'
 #' @examples

### 2021-11-21 tests/testthat/test_errorplotstats.R

@@ -1,5 +1,7 @@
+library(stats)
+
 iris_sub <- subset(iris, Species == "setosa")[1:4]
-data <- ggmatplot::matclean(iris_sub)$data
+data <- matclean(iris_sub)$data
 
 
 errorstats_mean_se <- errorplotstats(data, desc_stat = "mean_se")
@@ -8,8 +10,8 @@ errorstats_median_iqr <- errorplotstats(data, desc_stat = "median_iqr")
 errorstats_median_range <- errorplotstats(data, desc_stat = "median_range")
 
 test_that("format of resultant data frame is as expected", {
-  expect_named(errorstats, c("Group", "y", "ymin", "ymax"))
-  expect_setequal(errorstats$Group, colnames(iris_sub))
+  expect_named(errorstats_mean_se, c("Group", "y", "ymin", "ymax"))
+  expect_setequal(errorstats_mean_se$Group, colnames(iris_sub))
 })
 
 test_that("column means are calculated as expected", {

### 2021-11-21 R/errorplotstats.R
Truncated code at 80 characters per line

@@ -1,7 +1,8 @@
 #' Function to compute descriptive statistics to be used for errorplots
 #'
 #' @param data Resultant data frame from matclean()
-#' @param desc_stat Statistics to calculate. Either of `mean_se`, `mean_sd`, `mean_range`, `median_iqr`, `median_range`
+#' @param desc_stat Statistics to calculate. Either of `mean_se`, `mean_sd`,
+#' `mean_range`, `median_iqr`, `median_range`
 #'
 #' @return A data frame with descriptive statistics for each group.
 #'
@@ -22,8 +23,10 @@ errorplotstats <- function(data, desc_stat = "") {
     errorplot_data <- do.call(rbind, by(data, list(data$Group), function(x) {
       c(
         y = mean(x$x, na.rm = TRUE),
-        ymin = mean(x$x, na.rm = TRUE) - (sd(x$x, na.rm = TRUE) / sqrt(nrow(x))),
-        ymax = mean(x$x, na.rm = TRUE) + (sd(x$x, na.rm = TRUE) / sqrt(nrow(x)))
+        ymin = mean(x$x, na.rm = TRUE) -
+          (sd(x$x, na.rm = TRUE) / sqrt(nrow(x))),
+        ymax = mean(x$x, na.rm = TRUE) +
+          (sd(x$x, na.rm = TRUE) / sqrt(nrow(x)))
       )
     }))
   } else if (desc_stat == "mean_sd") {

### 2021-11-21 R/ggmatplot.R

@@ -1,20 +1,21 @@
 #' ggmatplot
 #'
 #'
-#' `ggmatplot` is a quick and easy way of plotting the columns of two matrices or
-#'  data frames against each other using [`ggplot2`](https://ggplot2.tidyverse.org/).
+#' `ggmatplot` is a quick and easy way of plotting the columns of two matrices
+#' or data frames against each other using
+#' [`ggplot2`](https://ggplot2.tidyverse.org/).
 #'
 #' @param x,y Vectors or matrices of data.
 #'
 #' * The number of rows of `x` and `y` should be the same.
-#' * Either `x` or `y` should be a vector, unless the number of columns of `x` and
-#' `y` are the same.
+#' * Either `x` or `y` should be a vector, unless the number of columns of `x`
+#' and `y` are the same.
 #' * Missing values (NAs) are allowed.
 #' * If either `x` or `y` is missing, the other is used as `y` and a vector of
 #' `1:n` is used as `x`.
 #'
-#' @param plot_type A string specifying the type of plot. Possible plot types are
-#' `point`, `line`, `both`(point + line), `density`, `histogram`, `boxplot`,
+#' @param plot_type A string specifying the type of plot. Possible plot types
+#' are `point`, `line`, `both`(point + line), `density`, `histogram`, `boxplot`,
 #' `dotplot`, `errorplot`, `violin`, and `ecdf`. Default plot_type is `point`.
 #'
 #' @param color,fill Vectors of colors. Defining only one of them will update
@@ -23,19 +24,20 @@
 #'
 #' * The number of colors should match the higher number of columns of
 #' matrices `x` or `y`, and will correspond to each of those columns.
-#' * If only a single color is given, the same color will be used for all columns.
+#' * If only a single color is given, the same color will be used for all
+#' columns.
 #'
-#' @param main,xlab,ylab,legend_title Strings to update plot title, x axis label,
-#'  y axis label and legend title respectively.
+#' @param main,xlab,ylab,legend_title Strings to update plot title, x axis
+#' label, y axis label and legend title respectively.
 #'
 #' @param legend_label A vector of strings, to rename the legend labels.
 #'
 #' @param shape,linetype A vector of shapes or line types respectively.
 #'
-#' * The number of shapes/line types should match the higher number of columns of
[…]

### 2021-11-21 R/matclean.R

@@ -1,6 +1,8 @@
 #' Function to convert two matrices into a wide format data frame
 #'
-#' @param x,y Vectors or matrices of data. The number of rows of x and y should match. If one of them are missing, the other is taken as y and an x vector of 1:n is used.
+#' @param x,y Vectors or matrices of data. The number of rows of x and y should
+#' match. If one of them are missing, the other is taken as y and an x vector
+#' of 1:n is used.
 #'
 #' @return A list of containing:
 #'  * data - the long format data frame.
@@ -15,7 +17,6 @@
 #' y <- iris_sub[, 3:4]
 #' # Use the defined x and y matrices as parameters
 #' matclean(x, y)
-#'
 matclean <- function(x, y) {
   x <- data.frame(x)
   ncolx <- ncol(x)
@@ -39,14 +40,20 @@ matclean <- function(x, y) {
     ncol <- ncol(data)
     yname <- colnames(data)[ncol]
     data$Observation_number <- 1:nrow(data)
-    data <- pivotlonger(data, names_to = "Group", values_to = xname, c(ncol, (ncol + 1)))
+    data <- pivotlonger(data,
+      names_to = "Group", values_to = xname,
+      c(ncol, (ncol + 1))
+    )
   } else if (ncoly > ncolx & ncolx == 1) {
     xname <- colnames(x)
     yname <- "y"
     data <- data.frame(x, y)
     ncol <- ncol(data)
     data$Observation_number <- 1:nrow(data)
-    data <- pivotlonger(data, names_to = "Group", values_to = yname, c(1, (ncol + 1)))
+    data <- pivotlonger(data,
+      names_to = "Group", values_to = yname,
+      c(1, (ncol + 1))
+    )
   } else if (ncolx == ncoly) {
     xname <- "x"
     yname <- "y"
@@ -61,7 +68,8 @@ matclean <- function(x, y) {
     names(data)[names(data) == paste0("Group", ".x")] <- "Group"
     data <- subset(data, select = -c(get(paste0("Group", ".y"))))
   } else {
-    stop("Either x or y must have only 1 column, or both x and y must have the same number of columns", call. = FALSE)
+    stop("Either x or y must have only 1 column, or both x and y must have the
+         same number of columns", call. = FALSE)
   }
 
   return(list(data = data, xname = xname, yname = yname))

### 2021-11-21 R/pivotlonger.R

@@ -11,8 +11,10 @@
 #' @examples
 #' # Define a data set
 #' iris_sub <- subset(iris, Species == "setosa")[1:4]
-#' # Use first two columns as ID columns and pivot the second two columns to long format
-#' pivotlonger(data = iris_sub, names_to = "Measurement", values_to = "Value", id_cols = c(1, 2))
+#' # Use first two columns as ID columns and pivot the second two columns to
+#' # long format
+#' pivotlonger(data = iris_sub, names_to = "Measurement", values_to = "Value",
+#' id_cols = c(1, 2))
 #'
 pivotlonger <- function(data, names_to = "", values_to = "", id_cols = NULL) {
   long_df <- data.frame(matrix(ncol = 2 + length(id_cols), nrow = 0))

### 2021-11-21 R/validateNumParams.R

@@ -1,7 +1,9 @@
-#' Function to validate the number of parameter values against the number of groups
+#' Function to validate the number of parameter values against the number of
+#' groups
 #'
 #' @param parameterValues A list of values.
-#' @param numGroups An integer denoting the number of unique values of the grouping variable.
+#' @param numGroups An integer denoting the number of unique values of the
+#' grouping variable.
 #'
 #' @return If valid, a list of parameter values. If invalid, an error message.
 #' @noRd
@@ -13,11 +15,17 @@ validateNumParams <- function(parameterValues, numGroups) {
   }
   # values > number of unique groups
   else if (length(parameterValues) > numGroups) {
-    stop(paste0("Too many ", substitute(parameterValues), " values. Only ", numGroups, " needed but ", length(parameterValues), " provided."), call. = FALSE)
+    stop(paste0(
+      "Too many ", substitute(parameterValues), " values. Only ",
+      numGroups, " needed but ", length(parameterValues), " provided."
+    ), call. = FALSE)
   }
   # values < number of unique groups
   else if (length(parameterValues) < numGroups) {
-    stop(paste0("Insufficient ", substitute(parameterValues), " values. ", numGroups, " needed but only ", length(parameterValues), " provided."), call. = FALSE)
+    stop(paste0(
+      "Insufficient ", substitute(parameterValues), " values. ",
+      numGroups, " needed but only ", length(parameterValues), " provided."
+    ), call. = FALSE)
   }
   # values = number of unique groups
   else {

### 2021-11-21 tests/testthat/test_errorplotstats.R

@@ -10,32 +10,68 @@ errorstats_median_iqr <- errorplotstats(data, desc_stat = "median_iqr")
 errorstats_median_range <- errorplotstats(data, desc_stat = "median_range")
 
 test_that("format of resultant data frame is as expected", {
-  expect_named(errorstats_mean_se, c("Group", "y", "ymin", "ymax"))
-  expect_setequal(errorstats_mean_se$Group, colnames(iris_sub))
+  expect_named(
+    errorstats_mean_se,
+    c("Group", "y", "ymin", "ymax")
+  )
+  expect_setequal(
+    errorstats_mean_se$Group,
+    colnames(iris_sub)
+  )
 })
 
 test_that("column means are calculated as expected", {
-  suppressWarnings(expect_setequal(errorstats_mean_se$y, colMeans(iris_sub)))
+  suppressWarnings(expect_setequal(
+    errorstats_mean_se$y,
+    colMeans(iris_sub)
+  ))
 })
 
 test_that("column medians are calculated as expected", {
-  suppressWarnings(expect_setequal(errorstats_median_iqr$y, sapply(iris_sub, median)))
+  suppressWarnings(expect_setequal(
+    errorstats_median_iqr$y,
+    sapply(iris_sub, median)
+  ))
 })
 
 test_that("column errors bars are calculated as expected", {
   # when error bars use standard errors
-  suppressWarnings(expect_setequal(errorstats_mean_se$ymin, (sapply(iris_sub, mean)-(sapply(iris_sub, sd)/sqrt(nrow(iris_sub))))))
-  suppressWarnings(expect_setequal(errorstats_mean_se$ymax, (sapply(iris_sub, mean)+(sapply(iris_sub, sd)/sqrt(nrow(iris_sub))))))
+  suppressWarnings(expect_setequal(
+    errorstats_mean_se$ymin,
+    (sapply(iris_sub, mean) - (sapply(iris_sub, sd) / sqrt(nrow(iris_sub))))
+  ))
+  suppressWarnings(expect_setequal(
+    errorstats_mean_se$ymax,
+    (sapply(iris_sub, mean) + (sapply(iris_sub, sd) / sqrt(nrow(iris_sub))))
+  ))
 
   # when error bars use standard deviations
-  suppressWarnings(expect_setequal(errorstats_mean_sd$ymin, (sapply(iris_sub, mean)-(sapply(iris_sub, sd)))))
[…]

### 2021-11-21 tests/testthat/test_ggmatplot_baseplots.R

@@ -4,51 +4,103 @@ x <- iris[, 1]
 y <- iris[, 3:4]
 
 test_that("scatterplot draws correctly", {
-  expect_doppelganger("basic scatterplot", ggmatplot(x, y, plot_type = "point"))
+  expect_doppelganger(
+    "basic scatterplot",
+    ggmatplot(x, y, plot_type = "point")
+  )
 })
 
 test_that("line plot draws correctly", {
-  expect_doppelganger("basic line plot", ggmatplot(x, y, plot_type = "line"))
+  expect_doppelganger(
+    "basic line plot",
+    ggmatplot(x, y, plot_type = "line")
+  )
 })
 
 test_that("point + line plot draws correctly", {
-  expect_doppelganger("basic point + line plot", ggmatplot(x, y, plot_type = "both"))
+  expect_doppelganger(
+    "basic point + line plot",
+    ggmatplot(x, y, plot_type = "both")
+  )
 })
 
 test_that("density plot draws correctly", {
-  expect_doppelganger("basic density plot", ggmatplot(y, plot_type = "density"))
+  expect_doppelganger(
+    "basic density plot",
+    ggmatplot(y, plot_type = "density")
+  )
 })
 
 test_that("histogram draws correctly", {
-  expect_doppelganger("basic histogram", ggmatplot(y, plot_type = "histogram"))
+  expect_doppelganger(
+    "basic histogram",
+    ggmatplot(y, plot_type = "histogram")
+  )
 })
 
 test_that("violin plot draws correctly", {
-  expect_doppelganger("basic violin plot", ggmatplot(y, plot_type = "violin"))
+  expect_doppelganger(
+    "basic violin plot",
+    ggmatplot(y, plot_type = "violin")
+  )
 })
 
 test_that("boxplot draws correctly", {
-  expect_doppelganger("basic boxplot", ggmatplot(y, plot_type = "boxplot"))
+  expect_doppelganger(
+    "basic boxplot",
+    ggmatplot(y, plot_type = "boxplot")
+  )
 })
 
 test_that("dotplot draws correctly", {
-  expect_doppelganger("basic dotplot", ggmatplot(y, plot_type = "dotplot"))
+  expect_doppelganger(
+    "basic dotplot",
+    ggmatplot(y, plot_type = "dotplot")
+  )
 })
 
 test_that("ecdf draws correctly", {
[…]

### 2021-11-21 tests/testthat/test_ggmatplot_parameters.R

@@ -5,124 +5,266 @@ y <- iris[, 2:4]
 
 test_that("plot colors by color values", {
   # single color value
-  expect_doppelganger("single color scatterplot", ggmatplot(x, y, plot_type = "point", color = "red"))
+  expect_doppelganger(
+    "single color scatterplot",
+    ggmatplot(x, y, plot_type = "point", color = "red")
+  )
   # color values = number of groups
-  expect_doppelganger("three color violin plot", ggmatplot(y, plot_type = "violin", color = c("red", "blue", "#123456")))
+  expect_doppelganger(
+    "three color violin plot",
+    ggmatplot(y, plot_type = "violin", color = c("red", "blue", "#123456"))
+  )
 })
 
 test_that("plot colors by fill values", {
   # single fill value
-  expect_doppelganger("single fill density plot", ggmatplot(y, plot_type = "density", fill = "red"))
+  expect_doppelganger(
+    "single fill density plot",
+    ggmatplot(y, plot_type = "density", fill = "red")
+  )
   # fill values = number of groups
-  expect_doppelganger("three fill color violin plot", ggmatplot(y, plot_type = "violin", fill = c("red", "blue", "#123456")))
+  expect_doppelganger(
+    "three fill color violin plot",
+    ggmatplot(y, plot_type = "violin", fill = c("red", "blue", "#123456"))
+  )
 })
 
 test_that("plot colors by color and fill values simultaneously", {
   # single color value, fill values = number of groups
-  expect_doppelganger("color and fill density plot", ggmatplot(y, plot_type = "density", color = "red", fill = c("red", "blue", "#123456")))
+  expect_doppelganger(
+    "color and fill density plot",
+    ggmatplot(y,
+      plot_type = "density", color = "red",
+      fill = c("red", "blue", "#123456")
+    )
+  )
 })
 
 test_that("point shapes updated based on shape values", {
   # single shape value
-  expect_doppelganger("single shape scatterplot", ggmatplot(y, plot_type = "point", shape = "square"))
+  expect_doppelganger(
+    "single shape scatterplot",
+    ggmatplot(y, plot_type = "point", shape = "square")
+  )
   # shape values = number of groups
-  expect_doppelganger("three shape scatterplot", ggmatplot(x, y, plot_type = "point", shape = c(15, 12, 3)))
[…]

### 2021-11-21 tests/testthat/test_matclean.R

@@ -1,7 +1,6 @@
 library(tibble)
 
 test_that("converts a single matrix into long format", {
-
   x <- tibble(a = 1:3, b = 11:13, c = 31:33)
   outputList <- matclean(x)
 
@@ -12,15 +11,17 @@ test_that("converts a single matrix into long format", {
 
   expect_equal(outputList$data$Group, rep(names(x), nrow(x)))
   expect_equal(outputList$data$x, c(1, 11, 31, 2, 12, 32, 3, 13, 33))
-  expect_equal(outputList$data$Observation_number, unlist(lapply(1:nrow(x), function(var) rep(var, ncol(x)))))
-
+  expect_equal(
+    outputList$data$Observation_number,
+    unlist(lapply(1:nrow(x), function(var) rep(var, ncol(x))))
+  )
 })
 
-test_that("converts two matrices(x and y) into long format (ncolx>ncoly & ncoly=1)", {
-
+test_that("converts two matrices(x and y) into long format
+          (ncolx>ncoly & ncoly=1)", {
   x <- tibble(a = 1:3, b = 11:13, c = 31:33)
   y <- tibble(d = 51:53)
-  outputList <- matclean(x,y)
+  outputList <- matclean(x, y)
 
   expect_equal(outputList$xname, "x")
   expect_equal(outputList$yname, names(y))
@@ -28,15 +29,21 @@ test_that("converts two matrices(x and y) into long format (ncolx>ncoly & ncoly=
 
   expect_equal(outputList$data$Group, rep(names(x), nrow(x)))
   expect_equal(outputList$data$x, c(1, 11, 31, 2, 12, 32, 3, 13, 33))
-  expect_equal(outputList$data$Observation_number, unlist(lapply(1:nrow(x), function(var) rep(var, ncol(x)))))
-  expect_equal(outputList$data$d, unlist(lapply(y$d, function(var) rep(var, ncol(x)))))
+  expect_equal(
+    outputList$data$Observation_number,
+    unlist(lapply(1:nrow(x), function(var) rep(var, ncol(x))))
+  )
+  expect_equal(
+    outputList$data$d,
+    unlist(lapply(y$d, function(var) rep(var, ncol(x))))
+  )
 })
 
-test_that("converts two matrices(x and y) into long format (ncolx<ncoly & ncolx=1)", {
-
+test_that("converts two matrices(x and y) into long format
+          (ncolx<ncoly & ncolx=1)", {
[…]

### 2021-11-21 tests/testthat/test_pivotlonger.R

@@ -3,16 +3,19 @@ test_that("can pivot columns to long format based on a single id column", {
   pv <- pivotlonger(df, names_to = "Group", values_to = "Value", id_cols = 1)
 
   expect_named(pv, c("x", "Group", "Value"))
-  expect_equal(pv$Group, rep(names(df[,2:3]), nrow(df)))
+  expect_equal(pv$Group, rep(names(df[, 2:3]), nrow(df)))
   expect_equal(pv$Value, c(11, 31, 12, 32, 13, 33))
 })
 
 test_that("can pivot columns to long format based on a multiple id columns", {
   df <- tibble(x = 1:3, y = 11:13, z = 31:33, a = 41:43, b = 51:53)
-  pv <- pivotlonger(df, names_to = "Group", values_to = "Value", id_cols = c(1,2,3))
+  pv <- pivotlonger(df,
+    names_to = "Group", values_to = "Value",
+    id_cols = c(1, 2, 3)
+  )
 
   expect_named(pv, c("x", "y", "z", "Group", "Value"))
-  expect_equal(pv$Group, rep(names(df[,4:5]), nrow(df)))
+  expect_equal(pv$Group, rep(names(df[, 4:5]), nrow(df)))
   expect_equal(pv$Value, c(41, 51, 42, 52, 43, 53))
 })
 
@@ -22,17 +25,20 @@ test_that("NA values are preserved in both id columns and pivoted columns", {
 
   expect_named(pv, c("x", "Group", "Value"))
   expect_equal(pv$x, c(1, 1, 2, 2, NA, NA))
-  expect_equal(pv$Group, rep(names(df[,2:3]), nrow(df)))
+  expect_equal(pv$Group, rep(names(df[, 2:3]), nrow(df)))
   expect_equal(pv$Value, c(NA, 3, 2, 4, 3, 5))
 })
 
 test_that("id columns can have different data types", {
-  df <- tibble(x = 1:3, y = c("a","b","c"), z = 31:33)
-  pv <- pivotlonger(df, names_to = "Group", values_to = "Value", id_cols = c(1,2))
+  df <- tibble(x = 1:3, y = c("a", "b", "c"), z = 31:33)
+  pv <- pivotlonger(df,
+    names_to = "Group", values_to = "Value",
+    id_cols = c(1, 2)
+  )
 
   expect_named(pv, c("x", "y", "Group", "Value"))
[…]

### 2021-11-21 R/matclean.R
Corrected matclean.R error text

@@ -68,8 +68,9 @@ matclean <- function(x, y) {
     names(data)[names(data) == paste0("Group", ".x")] <- "Group"
     data <- subset(data, select = -c(get(paste0("Group", ".y"))))
   } else {
-    stop("Either x or y must have only 1 column, or both x and y must have the
-         same number of columns", call. = FALSE)
+    stop(
+    "Either x or y must have only 1 column, or both x and y must have the same number of columns",
+    call. = FALSE)
   }
 
   return(list(data = data, xname = xname, yname = yname))

### 2021-11-21 R/errorplotstats.R
Added codemeta.json

@@ -6,9 +6,7 @@
 #'
 #' @return A data frame with descriptive statistics for each group.
 #'
-#' @importFrom stats sd
-#' @importFrom stats median
-#' @importFrom stats quantile
+#' @importFrom stats sd median quantile
 #'
 #' @noRd
 #'

### 2021-11-21 R/ggmatplot.R

@@ -66,9 +66,9 @@
 #' @section Plot Types:
 #'
 #' `ggmatplot`plots are built upon `ggplot2 layers`. The following is a list of
-#' `ggmatplot` plot types, along with their underlying [`ggplot geoms`]
-#' (https://ggplot2.tidyverse.org/reference/index.html#section-geoms) or
-#'  [`stats`](https://ggplot2.tidyverse.org/reference/index.html#section-stats).
+#' `ggmatplot` plot types, along with their underlying
+#' [`ggplot geoms`](https://ggplot2.tidyverse.org/reference/index.html#section-geoms)
+#' or [`stats`](https://ggplot2.tidyverse.org/reference/index.html#section-stats).
 #'
 #' * \strong{point} \code{\link[ggplot2]{geom_point}}
 #' * \strong{line} \code{\link[ggplot2]{geom_line}}

### 2021-11-23 README.Rmd
mild wording changes

@@
 Next is plot of the US personal expenditure over 5 categories and 5 years, and is a simple example of how wide format data can be used with `ggmatplot()`.
-Note how the expenditure categories to be used on the x axis is used as vector x, and the expenditure values is used in wide format as matrix y - with its columns corresponding to the grouping structure.
+Note how the expenditure categories to be used on the x axis is used as vector `x`, and the expenditure values is used in wide format as matrix `y` - with its columns corresponding to the grouping structure.
 

### 2021-11-24 README.Rmd
spelling check

@@
 
-The next example uses the `plot_type = "dotplot"` to visualise the distribution of the data, and also customizes the position of the legend using a `ggplot` theme as well.
+The next example uses the `plot_type = "dotplot"` to visualize the distribution of the data, and also customizes the position of the legend using a `ggplot` theme as well.
 

### 2021-11-24 tests/testthat/test_ggmatplot_baseplots.R
skipping vdiffr tests for cran release

@@ -4,6 +4,8 @@ x <- iris[, 1]
 y <- iris[, 3:4]
 
 test_that("scatterplot draws correctly", {
+  skip_on_cran()
+  skip_on_os("windows")
   expect_doppelganger(
     "basic scatterplot",
     ggmatplot(x, y, plot_type = "point")
@@ -11,6 +13,8 @@ test_that("scatterplot draws correctly", {
 })
 
 test_that("line plot draws correctly", {
+  skip_on_cran()
+  skip_on_os("windows")
   expect_doppelganger(
     "basic line plot",
     ggmatplot(x, y, plot_type = "line")
@@ -18,6 +22,8 @@ test_that("line plot draws correctly", {
 })
 
 test_that("point + line plot draws correctly", {
+  skip_on_cran()
+  skip_on_os("windows")
   expect_doppelganger(
     "basic point + line plot",
     ggmatplot(x, y, plot_type = "both")
@@ -25,6 +31,8 @@ test_that("point + line plot draws correctly", {
 })
 
 test_that("density plot draws correctly", {
+  skip_on_cran()
+  skip_on_os("windows")
   expect_doppelganger(
     "basic density plot",
     ggmatplot(y, plot_type = "density")
@@ -32,6 +40,8 @@ test_that("density plot draws correctly", {
 })
 
 test_that("histogram draws correctly", {
+  skip_on_cran()
+  skip_on_os("windows")
   expect_doppelganger(
     "basic histogram",
     ggmatplot(y, plot_type = "histogram")
@@ -39,6 +49,8 @@ test_that("histogram draws correctly", {
 })
 
 test_that("violin plot draws correctly", {
+  skip_on_cran()
+  skip_on_os("windows")
   expect_doppelganger(
     "basic violin plot",
     ggmatplot(y, plot_type = "violin")
@@ -46,6 +58,8 @@ test_that("violin plot draws correctly", {
 })
 
 test_that("boxplot draws correctly", {
+  skip_on_cran()
+  skip_on_os("windows")
   expect_doppelganger(
     "basic boxplot",
     ggmatplot(y, plot_type = "boxplot")
@@ -53,6 +67,8 @@ test_that("boxplot draws correctly", {
 })
 
 test_that("dotplot draws correctly", {
+  skip_on_cran()
+  skip_on_os("windows")
   expect_doppelganger(
     "basic dotplot",
[…]

### 2021-11-24 tests/testthat/test_ggmatplot_parameters.R
skip vdiffr tests for cran release

@@ -4,6 +4,8 @@ x <- iris[, 1]
 y <- iris[, 2:4]
 
 test_that("plot colors by color values", {
+  skip_on_cran()
+  skip_on_os("windows")
   # single color value
   expect_doppelganger(
     "single color scatterplot",
@@ -17,6 +19,8 @@ test_that("plot colors by color values", {
 })
 
 test_that("plot colors by fill values", {
+  skip_on_cran()
+  skip_on_os("windows")
   # single fill value
   expect_doppelganger(
     "single fill density plot",
@@ -30,6 +34,8 @@ test_that("plot colors by fill values", {
 })
 
 test_that("plot colors by color and fill values simultaneously", {
+  skip_on_cran()
+  skip_on_os("windows")
   # single color value, fill values = number of groups
   expect_doppelganger(
     "color and fill density plot",
@@ -41,6 +47,8 @@ test_that("plot colors by color and fill values simultaneously", {
 })
 
 test_that("point shapes updated based on shape values", {
+  skip_on_cran()
+  skip_on_os("windows")
   # single shape value
   expect_doppelganger(
     "single shape scatterplot",
@@ -54,6 +62,8 @@ test_that("point shapes updated based on shape values", {
 })
 
 test_that("line types updated based on linetype values", {
+  skip_on_cran()
+  skip_on_os("windows")
   # single linetype value
   expect_doppelganger(
     "single linetype line plot",
@@ -77,6 +87,8 @@ test_that("line types updated based on linetype values", {
 })
 
 test_that("plot axis limits are updated based on xlim and ylim parameters", {
+  skip_on_cran()
+  skip_on_os("windows")
   # x axis limits only
   suppressWarnings(expect_doppelganger(
     "density plot with x axis limits",
@@ -96,6 +108,8 @@ test_that("plot axis limits are updated based on xlim and ylim parameters", {
 
 test_that("plot axes are transformed to a log scale based on log parameter
           values", {
+  skip_on_cran()
+  skip_on_os("windows")
   # log(x)
   expect_doppelganger(
     "line plot with log x",
@@ -114,6 +128,8 @@ test_that("plot axes are transformed to a log scale based on log parameter
 })
 
 test_that("plot title is updated based on main paramter", {
+  skip_on_cran()
+  skip_on_os("windows")
   expect_equal(
[…]

### 2021-11-29 README.Rmd
update README to fix url issue

@@
 It is also worth noticing that `alpha` isn't a parameter defined in `ggmatplot()`, but can be used.
-This is because `ggmatplot` is built upon `ggplot2`, and each `plot_type` corresponds to a [`geom`](https://ggplot2.tidyverse.org/reference/index.html#section-geoms) as listed [here](./reference/ggmatplot.html#plot-types).
+This is because `ggmatplot` is built upon `ggplot2`, and each `plot_type` corresponds to a [`geom`](https://ggplot2.tidyverse.org/reference/index.html#section-geoms) as listed [here](https://xuan-liang.github.io/ggmatplot/reference/ggmatplot.html#plot-types).
 Therefore, all valid parameters with the underlying [`ggplot2 geom`](https://ggplot2.tidyverse.org/reference/index.html#section-geoms) can be used with `ggmatplot()`.

### 2021-11-29 R/ggmatplot.R
Documented ggmatplot function return value

@@ -59,6 +59,9 @@
 #' @param ... Other arguments passed on to the plot. Possible arguments are
 #' those that can be passed on to the [underlying ggplot layers](#plot-types).
 #'
+#' @return A ggplot object. The columns of the input matrices will be plotted
+#' against each other using the defined plot type.
+#'
 #' @import ggplot2
 #' @export
 #' @md

### 2021-11-29 R/matclean.R

@@ -4,7 +4,7 @@
 #' match. If one of them are missing, the other is taken as y and an x vector
 #' of 1:n is used.
 #'
-#' @return A list of containing:
+#' @return A list containing:
 #'  * data - the long format data frame.
 #'  * xname - name(s) of the ID column(s)
 #'  * yname - names of the pivoted columns(columns with group names and values)

### 2021-12-10 README.Rmd
Update install instructions in README
Test pdf issues on README

@@
 
-The latest version can be installed from [GitHub](https://github.com/xuan-liang/ggmatplot):
+The latest version can be installed from [CRAN](https://CRAN.R-project.org/package=ggmatplot):
 
-```r
+```{r eval=FALSE}
+install.packages("ggmatplot")
+```
+
+Or you can install the development version from [GitHub](https://github.com/xuan-liang/ggmatplot):
+
+```{r eval=FALSE}
 # install.packages("remotes")
@@
 
-The next example, plots out an `errorplot` using the medians and interquartile ranges of each variable.
+The next example construct out an `errorplot` using the medians and interquartile ranges of each variable.
 

### 2021-12-10 tests/testthat/test_errorplotstats.R
updated errorplotstats tests

@@ -22,15 +22,15 @@ test_that("format of resultant data frame is as expected", {
 
 test_that("column means are calculated as expected", {
   suppressWarnings(expect_setequal(
-    errorstats_mean_se$y,
-    colMeans(iris_sub)
+    round(errorstats_mean_se$y,3),
+    round(sapply(iris_sub, mean),3)
   ))
 })
 
 test_that("column medians are calculated as expected", {
   suppressWarnings(expect_setequal(
-    errorstats_median_iqr$y,
-    sapply(iris_sub, median)
+    round(errorstats_median_iqr$y,3),
+    round(sapply(iris_sub, median),3)
   ))
 })
 

### 2021-12-11 tests/testthat/test_errorplotstats.R
Updated errorplotstats tests to remove names from named vectors

@@ -23,14 +23,14 @@ test_that("format of resultant data frame is as expected", {
 test_that("column means are calculated as expected", {
   suppressWarnings(expect_setequal(
     round(errorstats_mean_se$y,3),
-    round(sapply(iris_sub, mean),3)
+    round(as.numeric(sapply(iris_sub, mean)),3)
   ))
 })
 
 test_that("column medians are calculated as expected", {
   suppressWarnings(expect_setequal(
     round(errorstats_median_iqr$y,3),
-    round(sapply(iris_sub, median),3)
+    round(as.numeric(sapply(iris_sub, median)),3)
   ))
 })
 

### 2022-02-28 Paper/paper.Rmd
rewrote the summary of the paper, partly reworked the statement of need, chaned the author order so it is alphabetically ordered (except first author)

---
title: '`ggmatplot`: An R package for data visualization on wide-format data'
tags:
- R
- grammar of graphics
- ggplot2
- data visualisation
authors:
  - name: Xuan Liang
    orcid: 0000-0002-0792-3751
    affiliation: 1
  - name: Francis Hui
    orcid: 0000-0003-0765-3533
    affiliation: 1
  - name: Dilinie Seimon
    orcid: 0000-0003-4431-6077
    affiliation: 2
  - name: Emi Tanaka
    orcid: 0000-0002-1455-259X
    affiliation: 2
affiliations:
  - name: Australian National University
    index: 1
  - name: Monash University
    index: 2
citation_author: Liang et. al.
date: "`r Sys.Date()`"
year: 2022
bibliography: paper.bib
output: rticles::joss_article
csl: apa.csl
journal: JOSS
---

# Summary

The layered grammar of graphics [@Wickham2010-kt], implemented as the `ggplot2` package [@Wickham2016] in the statistical language R [@rstats], is a powerful tool to create versatile statistical graphics. This graphical system, however, requires input data to be organised in a manner that a data column is mapped to an aesthetic element (e.g. x-coordinate, y-coordinate, color, size), which create friction in constructing plots with an aesthetic element that span multiple columns in the data by requiring users to re-organise the data.

The `ggmatplot`, built upon `ggplot2`, is an R-package that allows quick plotting across the columns of matrics or data with the result returned as a `ggplot` object. The package is inspired by the function `matplot()` in the core R `graphics` system, thus `ggmatplot` can be considered as a `ggplot` version of `matplot` with the benefits of customising the plots as any other `ggplot` objects via `ggplot2` functions.

# Statement of need

Data can be tidied in a rectangular form where each row represents an observational unit, each column represents a variable, and each cell represents a value [@Wickham2014-gy]. What constitutes a variable, hence a column, in a tidy data are dependent upon interpretation.

---

The increased use of a tidy approach in structuring datasets results in wrangling datasets into wide format: with variables as columns and observations as rows(https://vita.had.co.nz/papers/tidy-data.html). While this introduces a more meaningful mapping of the data into its structure, the grammar of graphics requires wide format data to be wrangled back into long format for visualizing. This can be quite cumbersome when creating simple plots, and converting a tidy data frame back to a long format with multiple variables stacked upon each other can also be considered counter-intuitive.

Therefore, the motivation for `ggmatplot` is to provide a solution that allows `ggplot2` to handle wide format data. Although `ggmatplot` doesn't provide the same flexibility as `ggplot2`, it can be used as a workaround for having to wrangle wide format data into long format and creating simple plots using `ggplot2`.

# Examples

# Acknowledgements

# References

### 2022-02-28 Paper/paper.Rmd
middle initials for FH, corrected ANU affiliation

@@
     affiliation: 1
-  - name: Francis Hui
+  - name: Francis K. C. Hui
     orcid: 0000-0003-0765-3533
@@
 affiliations:
-  - name: Australian National University
+  - name: The Australian National University
     index: 1

### 2022-03-02 Paper/paper.Rmd
rewrote statement of need and added example data

@@
 
+```{r setup, include = FALSE}
+library(tidyverse)
+library(knitr)
+library(kableExtra)
+knitr::opts_chunk$set(message = FALSE,
+                      warning = FALSE)
+
+# data
+
+wide_df <- tribble(~restaurant, ~person, ~food_rating, ~service_rating, ~ambience_rating, ~overall_rating,
+        "R1", "P1", 4, 3, 4, 4,
+        "R1", "P2", 4, 5, 4, 4,
+        "R1", "P3", 3, 4, 5, 3, 
+        "R2", "P1", 2, 4, 4, 3, 
+        "R2", "P2", 3, 4, 4, 3)
+
+long_df <- pivot_longer(wide_df, contains("rating"), 
+                        names_to = "rating_type",
+                        names_pattern = "(.+)_rating",
+                        values_to = "rating")
+
+abund_df <- tribble(~Site, ~`Soil dry mass`, ~`Moss`, ~Alopcune, ~Arctlute, ~Pardpull, ~Trocterr, ~Zoraspin,
+                    1, 2.3321, 3.0445, 10, 0, 45, 57, 4,
+                    2, 3.0493, 1.0986, 2, 0, 37, 65, 9,
+                    3, 2.5572, 2.3979, 20, 0, 45, 66, 1)
+
+table_styling <- function(df, ...) {
+  kable(df, ..., booktabs = TRUE, linesep = "") %>% 
+    kable_classic()
+}
+
+data("spider", package = "mvabund")
+```
+
 # Summary
 
-The layered grammar of graphics [@Wickham2010-kt], implemented as the `ggplot2` package [@Wickham2016] in the statistical language R [@rstats], is a powerful tool to create versatile statistical graphics.
+The layered grammar of graphics [@Wickham2010-kt], implemented as the `ggplot2` package [@Wickham2016] in the statistical language R [@rstats], is a powerful and popular tool to create versatile statistical graphics.
 This graphical system, however, requires input data to be organised in a manner that a data column is mapped to an aesthetic element (e.g.
-x-coordinate, y-coordinate, color, size), which create friction in constructing plots with an aesthetic element that span multiple columns in the data by requiring users to re-organise the data.
+x-coordinate, y-coordinate, color, size), which create friction in constructing plots with an aesthetic element that span multiple columns in the original data by requiring users to re-organise the data.
 
-The `ggmatplot`, built upon `ggplot2`, is an R-package that allows quick plotting across the columns of matrics or data with the result returned as a `ggplot` object.
+The `ggmatplot`, built upon `ggplot2`, is an R-package that allows quick plotting across the columns of matrices or data with the result returned as a `ggplot` object.
 The package is inspired by the function `matplot()` in the core R `graphics` system, thus `ggmatplot` can be considered as a `ggplot` version of `matplot` with the benefits of customising the plots as any other `ggplot` objects via `ggplot2` functions.
@@
 
-Data can be tidied in a rectangular form where each row represents an observational unit, each column represents a variable, and each cell represents a value [@Wickham2014-gy].
-What constitutes a variable, hence a column, in a tidy data are dependent upon interpretation.
-
----
-
-The increased use of a tidy approach in structuring datasets results in wrangling datasets into wide format: with variables as columns and observations as rows(https://vita.had.co.nz/papers/tidy-data.html).
-While this introduces a more meaningful mapping of the data into its structure, the grammar of graphics requires wide format data to be wrangled back into long format for visualizing.
-This can be quite cumbersome when creating simple plots, and converting a tidy data frame back to a long format with multiple variables stacked upon each other can also be considered counter-intuitive.
-
-Therefore, the motivation for `ggmatplot` is to provide a solution that allows `ggplot2` to handle wide format data.
-Although `ggmatplot` doesn't provide the same flexibility as `ggplot2`, it can be used as a workaround for having to wrangle wide format data into long format and creating simple plots using `ggplot2`.
+Input data to construct plots with `ggplot2` require data to be organised in a manner that maps data columns to aesthetic elements.
+This required form is, however, not consistent with some common data formats, specifically data where the column names contain information to be mapped into the plot or the variables to be mapped into the same aesthetic elements.
+Consequently, plotting with `ggplot2` interrupts the workflow of a user that is trying to quickly explore these types of data.
+The `ggmatplot` R-package provides a solution to this common friction to produce plots with `ggplot2`.
 
 # Examples
+
+Data can be tidied in a rectangular form where each row represents an observational unit, each column represents a variable, and each cell represents a value [@Wickham2014-gy].
+In some cases, what constitutes a variable (or observational unit), hence a column (or row), in a tidy data can be dependent upon interpretation or downstream interest (e.g.
+Tables \ref{tab:tab1} and \ref{tab:tab2} can be both considered as tidy data), but a clear violation of tidy data principles is when the column names contain data values, e.g.
+Table \ref{tab:tab3} contain the name of the species across a number of column names.
+
+```{r tab1, echo = FALSE}
+table_styling(wide_df, col.names = c("Restaurant", "Person", "Food",
+                                     "Service", "Ambience", "Overall"),
+              caption = 'Restaurant rating data in "tidy" form') %>% 
+  add_header_above(c(" ", " ", "Rating" = 4))
+```
+
+```{r tab2, echo = FALSE}
+table_styling(long_df, col.names = c("Restauant", "Person", "Rating type", "Rating"),
+              caption = 'Restaurant rating data in "molten" form')
+```
+
+```{r tab3, echo = FALSE}
+table_styling(abund_df,
+              caption = 'Spider abundance data with environmental covariates.') %>% 
+  add_header_above(c(" ", "Environment covariates" = 2, "Species abundance" = 5))
+```
 

### 2022-03-02 Paper/paper.Rmd
reworked example

@@
 
-wide_df <- tribble(~restaurant, ~person, ~food_rating, ~service_rating, ~ambience_rating, ~overall_rating,
-        "R1", "P1", 4, 3, 4, 4,
-        "R1", "P2", 4, 5, 4, 4,
-        "R1", "P3", 3, 4, 5, 3, 
-        "R2", "P1", 2, 4, 4, 3, 
-        "R2", "P2", 3, 4, 4, 3)
+wide_df <- tribble(~restaurant, ~food_rating, ~service_rating, ~ambience_rating, ~overall_rating,
+        "R1", 4, 3, 4, 4,
+        "R2", 4, 5, 4, 4,
+        "R3", 3, 4, 5, 3, 
+        "R4", 2, 4, 4, 3, 
+        "R5", 3, 4, 4, 3)
 
@@
 ```{r tab1, echo = FALSE}
-table_styling(wide_df, col.names = c("Restaurant", "Person", "Food",
+table_styling(wide_df, col.names = c("Restaurant", "Food",
                                      "Service", "Ambience", "Overall"),
               caption = 'Restaurant rating data in "tidy" form') %>% 
-  add_header_above(c(" ", " ", "Rating" = 4))
+  add_header_above(c(" ", "Rating" = 4))
 ```
@@
 ```{r tab2, echo = FALSE}
-table_styling(long_df, col.names = c("Restauant", "Person", "Rating type", "Rating"),
+table_styling(long_df, col.names = c("Restauant", "Rating type", "Rating"),
               caption = 'Restaurant rating data in "molten" form')
@@
 
+```{r plot1}
+library(ggmatplot)
+ggmatplot(x = select(wide_df, contains("rating")),
+          plot_type = "both",
+          xlab = "Restaurant")
+```
+
 # Acknowledgements

### 2022-03-03 Paper/paper.Rmd
statement of need updated

@@
 Input data to construct plots with `ggplot2` require data to be organised in a manner that maps data columns to aesthetic elements.
-This required form is, however, not consistent with some common data formats, specifically data where the column names contain information to be mapped into the plot or the variables to be mapped into the same aesthetic elements.
+This required form is, however, not consistent with some prevelant data formats, specifically multivariate data where the rows correspond to observational units and continguous columns correspond to the responses of the same measurement units.
+Examples of common utilization of this form include gene expression data in bioinformatics, where rows correspond to samples, column names correspond to genes, and cell values correspond to the expression level; and multi-abundance data in ecology, where rows correspond to site, column names correspond to species name, and cell values correspond to count.
 Consequently, plotting with `ggplot2` interrupts the workflow of a user that is trying to quickly explore these types of data.
-The `ggmatplot` R-package provides a solution to this common friction to produce plots with `ggplot2`.
+The `ggmatplot` R-package provides a solution to this common friction in producing plots with `ggplot2`.
 

### 2022-03-04 Paper/paper.Rmd
expanded on examples

--- 2022-03-02 Paper/paper.Rmd
@@
 affiliations:
-  - name: The Australian National University
+  - name: Research School of Finance, Actuarial Studies and Statistics, The Australian National University
     index: 1
-  - name: Monash University
+  - name: Department of Econometrics and Business Statistics, Monash University
     index: 2
@@
 bibliography: paper.bib
-output: rticles::joss_article
+output: 
+  rticles::joss_article:
+    keep_tex: true
 csl: apa.csl
@@
 Input data to construct plots with `ggplot2` require data to be organised in a manner that maps data columns to aesthetic elements.
-This required form is, however, not consistent with some common data formats, specifically data where the column names contain information to be mapped into the plot or the variables to be mapped into the same aesthetic elements.
-Consequently, plotting with `ggplot2` interrupts the workflow of a user that is trying to quickly explore these types of data.
-The `ggmatplot` R-package provides a solution to this common friction to produce plots with `ggplot2`.
-
-# Examples
-
-Data can be tidied in a rectangular form where each row represents an observational unit, each column represents a variable, and each cell represents a value [@Wickham2014-gy].
+This generally works well where data is tidied in a rectangular form, referred to as "tidy data" [@Wickham2014-gy], where each row represents an observational unit, each column represents a variable, and each cell represents a value.
 In some cases, what constitutes a variable (or observational unit), hence a column (or row), in a tidy data can be dependent upon interpretation or downstream interest (e.g.
@@
                                      "Service", "Ambience", "Overall"),
-              caption = 'Restaurant rating data in "tidy" form') %>% 
+              caption = 'Restaurant rating data in "tidy" form.
+The first column shows the restaurant ID, and the next four columns show the average ratings (out of 5) for food, service, ambience and overall, respectively.') %>%
   add_header_above(c(" ", "Rating" = 4))
@@
 table_styling(long_df, col.names = c("Restauant", "Rating type", "Rating"),
-              caption = 'Restaurant rating data in "molten" form')
+              caption = 'Another form for the restaurant rating data in Table \\ref{tab:tab1}. In @Wickham2014-gy, this format is called the "molten" data.')
 ```
@@
 table_styling(abund_df,
-              caption = 'Spider abundance data with environmental covariates.') %>% 
+              caption = 'Spider abundance data with environmental covariates.
+The rows correspond to the site, the first two columns are environmental covariates that measure the soil dry mass and cover moss, and the following five columns shows the abundance of the species.') %>%
   add_header_above(c(" ", "Environment covariates" = 2, "Species abundance" = 5))
@@
 
-```{r plot1}
+The organisation of the data is largely dependent on the downstream analysis and there is no one correct way to do this.
+Some forms of multivariate data, e.g.
+Table \ref{tab:tab3}, are prevalent in the field because it aligns as an input data for a modelling software and/or the format is more convenient for input or view of the data in spreadsheet format.
+However, this format is not consistent with the required format for `ggplot2`, and consequently, plotting with `ggplot2` interrupts the workflow of a user that is trying to quickly visualise these types of data.
+The `ggmatplot` R-package provides a solution to this common friction in producing plots with `ggplot2`.
+
+# Examples
+
+In this section we demonstrate the use of the `ggmatplot` package and contrast the specification with `ggplot2` with data wrangling using `dplyr` and `tidyr` [@Wickham2019] using the example data in Tables \ref{tab:tab1} and \ref{tab:tab3}, which are stored in the objects `wide_df` and `abun_df`, respectively.
+
+## Example 1
+
+```{r matplot1, fig.height = 2}
 library(ggmatplot)
-ggmatplot(x = select(wide_df, contains("rating")),
-          plot_type = "both",
-          xlab = "Restaurant")
+ggmatplot(x = wide_df[, -1], plot_type = "both",
+          xlab = "Restaurant",  ylab = "Rating", legend_title = "Type")
 ```
+
+```{r ggplot1, eval = FALSE}
+library(ggplot2)
+library(tidyr) # or library(tidyverse)
+wide_df %>% 
+  pivot_longer(contains("rating"), 
+               names_to = "rating_type",
+               values_to = "rating") %>% 
+  ggplot(aes(restaurant, rating, color = rating_type)) + 
+  geom_point() +
+  geom_line(aes(group = rating_type,
+                linetype = rating_type))
+```
+
+## Example 2
+
+```{r matplot2, fig.height = 2}
+ggmatplot(x = wide_df[, 2:4], y = wide_df[, 5], plot_type = "both")
+```
+
+# Discussion
+
+The `ggmatplot` R-package provides a solution to a common friction to quickly plotting multivariate data where the primary interest is mapping the column names as an aesthetic element.
+The solution provided however is a recipe-driven approach where the user can only produce plot types as many there are included in the `plot_type` option.
+Future development of the package could benefit from using a grammar approach, like in @Wilkinson2005-oz and @Wickham2010-kt, where plot types can be extensible.
 
@@
 
+FKCH was supported by ARC DECRA XXX.
+
 # References

### 2022-03-04 Paper/paper.tex
expanded on examples

\documentclass[10pt,a4paper,onecolumn]{article}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{tikz}
\usepackage{hyperref}
\hypersetup{colorlinks,breaklinks,
            urlcolor=[rgb]{0.0, 0.5, 1.0},
            linkcolor=[rgb]{0.0, 0.5, 1.0}}
\usepackage{caption}
\usepackage{tcolorbox}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{seqsplit}
\usepackage{fixltx2e} % provides \textsubscript
\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{paper.bib}

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.0cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Default font ------------------------------------------------------------
% \renewcommand\familydefault{\sfdefault}

% --- Style -------------------------------------------------------------------
\renewcommand{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-0.75cm}\includegraphics[width=5.5cm]{/Library/Frameworks/R.framework/Versions/4.1/Resources/library/rticles/rmarkdown/templates/joss/resources/JOSS-logo.png}}
\fancyhead[C]{}
\fancyhead[R]{}
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\footnotesize{\sffamily Liang et.
al., (2022). \texttt{ggmatplot}: An R package for data visualization on
wide-format
data. \textit{Journal of Open Source Software}, (), . \href{https://doi.org/}{https://doi.org/}}}

\fancyfoot[R]{\sffamily \thepage}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Macros ---------

\definecolor{linky}{rgb}{0.0, 0.5, 1.0}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

\newcommand{\ExternalLink}{%
   \tikz[x=1.2ex, y=1.2ex, baseline=-0.05ex]{%
       \begin{scope}[x=1ex, y=1ex]
           \clip (-0.1,-0.1)
               --++ (-0, 1.2)
               --++ (0.6, 0)
               --++ (0, -0.6)
               --++ (0.6, 0)
               --++ (0, -1);
           \path[draw,
               line width = 0.5,
               rounded corners=0.5]
               (0,0) rectangle (1,1);
       \end{scope}
       \path[draw, line width = 0.5] (0.5, 0.5)
           -- (1, 1);
       \path[draw, line width = 0.5] (0.6, 1)
           -- (1, 1) -- (1, 0.6);
       }
   }

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}

\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}

\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}

\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={ggmatplot: An R package for data visualization on wide-format data},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% Pandoc citation processing
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for calculating minipage widths
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\title{\texttt{ggmatplot}: An R package for data visualization on
wide-format data}

        \author[1]{Xuan Liang}
          \author[1]{Francis K. C. Hui}
          \author[2]{Dilinie Seimon}
          \author[2]{Emi Tanaka}

      \affil[1]{Research School of Finance, Actuarial Studies and
Statistics, The Australian National University}
      \affil[2]{Department of Econometrics and Business Statistics,
Monash University}
  \date{\vspace{-5ex}}

\begin{document}
\maketitle

\marginpar{
  %\hrule
  \sffamily\small

  {\bfseries DOI:} \href{https://doi.org/}{\color{linky}{}}

  \vspace{2mm}

  {\bfseries Software}
  \begin{itemize}
    \setlength\itemsep{0em}
    \item \href{}{\color{linky}{Review}} \ExternalLink
    \item \href{}{\color{linky}{Repository}} \ExternalLink
    \item \href{}{\color{linky}{Archive}} \ExternalLink
  \end{itemize}

  \vspace{2mm}

  {\bfseries Submitted:} \\
  {\bfseries Published:}

  \vspace{2mm}
  {\bfseries License}\\
  Authors of papers retain copyright and release the work under a Creative Commons Attribution 4.0 International License (\href{http://creativecommons.org/licenses/by/4.0/}{\color{linky}{CC-BY}}).
}

\hypertarget{summary}{%
\section{Summary}\label{summary}}

The layered grammar of graphics (H. Wickham, 2010), implemented as the
\texttt{ggplot2} package (Hadley Wickham, 2016) in the statistical
language R (R Core Team, 2021), is a powerful and popular tool to create
versatile statistical graphics. This graphical system, however, requires
input data to be organised in a manner that a data column is mapped to
an aesthetic element (e.g.~x-coordinate, y-coordinate, color, size),
which create friction in constructing plots with an aesthetic element
that span multiple columns in the original data by requiring users to
re-organise the data.

The \texttt{ggmatplot}, built upon \texttt{ggplot2}, is an R-package
that allows quick plotting across the columns of matrices or data with
the result returned as a \texttt{ggplot} object. The package is inspired
by the function \texttt{matplot()} in the core R \texttt{graphics}
system, thus \texttt{ggmatplot} can be considered as a \texttt{ggplot}
version of \texttt{matplot} with the benefits of customising the plots
as any other \texttt{ggplot} objects via \texttt{ggplot2} functions.

\hypertarget{statement-of-need}{%
\section{Statement of need}\label{statement-of-need}}

Input data to construct plots with \texttt{ggplot2} require data to be
organised in a manner that maps data columns to aesthetic elements. This
generally works well where data is tidied in a rectangular form,
referred to as ``tidy data'' (Hadley Wickham, 2014), where each row
represents an observational unit, each column represents a variable, and
each cell represents a value. In some cases, what constitutes a variable
(or observational unit), hence a column (or row), in a tidy data can be
dependent upon interpretation or downstream interest (e.g.~Tables
\ref{tab:tab1} and \ref{tab:tab2} can be both considered as tidy data),
but a clear violation of tidy data principles is when the column names
contain data values, e.g.~Table \ref{tab:tab3} contain the name of the
species across a number of column names.

\begin{table}

\caption{\label{tab:tab1}Restaurant rating data in "tidy" form. The first column shows the restaurant ID, and the next four columns show the average ratings (out of 5) for food, service, ambience and overall, respectively.}
\centering
\begin{tabular}[t]{lrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Rating} \\
\cmidrule(l{3pt}r{3pt}){2-5}
Restaurant & Food & Service & Ambience & Overall\\
\midrule
R1 & 4 & 3 & 4 & 4\\
R2 & 4 & 5 & 4 & 4\\
R3 & 3 & 4 & 5 & 3\\
R4 & 2 & 4 & 4 & 3\\
R5 & 3 & 4 & 4 & 3\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}

\caption{\label{tab:tab2}Another form for the restaurant rating data in Table \ref{tab:tab1}. In @Wickham2014-gy, this format is called the "molten" data.}
\centering
\begin{tabular}[t]{llr}
\toprule
Restauant & Rating type & Rating\\
\midrule
R1 & food & 4\\
R1 & service & 3\\
R1 & ambience & 4\\
R1 & overall & 4\\
R2 & food & 4\\
R2 & service & 5\\
R2 & ambience & 4\\
R2 & overall & 4\\
R3 & food & 3\\
R3 & service & 4\\
R3 & ambience & 5\\
R3 & overall & 3\\
R4 & food & 2\\
R4 & service & 4\\
R4 & ambience & 4\\
R4 & overall & 3\\
R5 & food & 3\\
R5 & service & 4\\
R5 & ambience & 4\\
R5 & overall & 3\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}

\caption{\label{tab:tab3}Spider abundance data with environmental covariates. The rows correspond to the site, the first two columns are environmental covariates that measure the soil dry mass and cover moss, and the following five columns shows the abundance of the species.}
\centering
\begin{tabular}[t]{rrrrrrrr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Environment covariates} & \multicolumn{5}{c}{Species abundance} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-8}
Site & Soil dry mass & Moss & Alopcune & Arctlute & Pardpull & Trocterr & Zoraspin\\
\midrule
1 & 2.3321 & 3.0445 & 10 & 0 & 45 & 57 & 4\\
2 & 3.0493 & 1.0986 & 2 & 0 & 37 & 65 & 9\\
3 & 2.5572 & 2.3979 & 20 & 0 & 45 & 66 & 1\\
\bottomrule
\end{tabular}
\end{table}

The organisation of the data is largely dependent on the downstream
analysis and there is no one correct way to do this. Some forms of
multivariate data, e.g.~Table \ref{tab:tab3}, are prevalent in the field
because it aligns as an input data for a modelling software and/or the
format is more convenient for input or view of the data in spreadsheet
format. However, this format is not consistent with the required format
for \texttt{ggplot2}, and consequently, plotting with \texttt{ggplot2}
interrupts the workflow of a user that is trying to quickly visualise
these types of data. The \texttt{ggmatplot} R-package provides a
solution to this common friction in producing plots with
\texttt{ggplot2}.

\hypertarget{examples}{%
\section{Examples}\label{examples}}

In this section we demonstrate the use of the \texttt{ggmatplot} package
and contrast the specification with \texttt{ggplot2} with data wrangling
using \texttt{dplyr} and \texttt{tidyr} (Hadley Wickham et al., 2019)
using the example data in Tables \ref{tab:tab1} and \ref{tab:tab3},
which are stored in the objects \texttt{wide\_df} and \texttt{abun\_df},
respectively.

\hypertarget{example-1}{%
\subsection{Example 1}\label{example-1}}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(ggmatplot)}
\FunctionTok{ggmatplot}\NormalTok{(}\AttributeTok{x =}\NormalTok{ wide\_df[, }\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{], }\AttributeTok{plot\_type =} \StringTok{"both"}\NormalTok{,}
          \AttributeTok{xlab =} \StringTok{"Restaurant"}\NormalTok{,  }\AttributeTok{ylab =} \StringTok{"Rating"}\NormalTok{, }\AttributeTok{legend\_title =} \StringTok{"Type"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{paper_files/figure-latex/matplot1-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(ggplot2)}
\FunctionTok{library}\NormalTok{(tidyr) }\CommentTok{\# or library(tidyverse)}
\NormalTok{wide\_df }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{pivot\_longer}\NormalTok{(}\FunctionTok{contains}\NormalTok{(}\StringTok{"rating"}\NormalTok{), }
               \AttributeTok{names\_to =} \StringTok{"rating\_type"}\NormalTok{,}
               \AttributeTok{values\_to =} \StringTok{"rating"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(restaurant, rating, }\AttributeTok{color =}\NormalTok{ rating\_type)) }\SpecialCharTok{+} 
  \FunctionTok{geom\_point}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{geom\_line}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{group =}\NormalTok{ rating\_type,}
                \AttributeTok{linetype =}\NormalTok{ rating\_type))}
\end{Highlighting}
\end{Shaded}

\hypertarget{example-2}{%
\subsection{Example 2}\label{example-2}}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggmatplot}\NormalTok{(}\AttributeTok{x =}\NormalTok{ wide\_df[, }\DecValTok{2}\SpecialCharTok{:}\DecValTok{4}\NormalTok{], }\AttributeTok{y =}\NormalTok{ wide\_df[, }\DecValTok{5}\NormalTok{], }\AttributeTok{plot\_type =} \StringTok{"both"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{paper_files/figure-latex/matplot2-1.pdf}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

The \texttt{ggmatplot} R-package provides a solution to a common
friction to quickly plotting multivariate data where the primary
interest is mapping the column names as an aesthetic element. The
solution provided however is a recipe-driven approach where the user can
only produce plot types as many there are included in the
\texttt{plot\_type} option. Future development of the package could
benefit from using a grammar approach, like in Wilkinson (2005) and H.
Wickham (2010), where plot types can be extensible.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

FKCH was supported by ARC DECRA XXX.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-rstats}{}}%
R Core Team. (2021). \emph{R: A language and environment for statistical
computing}. Vienna, Austria: R Foundation for Statistical Computing.
Retrieved from \url{https://www.R-project.org/}

\leavevmode\vadjust pre{\hypertarget{ref-Wickham2010-kt}{}}%
Wickham, H. (2010). A layered grammar of graphics. \emph{Journal of
computational and graphical statistics: a joint publication of American
Statistical Association, Institute of Mathematical Statistics, Interface
Foundation of North America}.

\leavevmode\vadjust pre{\hypertarget{ref-Wickham2014-gy}{}}%
Wickham, Hadley. (2014). Tidy data. \emph{Journal of Statistical
Software}, \emph{59}(10), 1--23.
doi:\href{https://doi.org/10.18637/jss.v059.i10}{10.18637/jss.v059.i10}

\leavevmode\vadjust pre{\hypertarget{ref-Wickham2016}{}}%
Wickham, Hadley. (2016). \emph{ggplot2: Elegant graphics for data
analysis}. Springer-Verlag New York. Retrieved from
\url{https://ggplot2.tidyverse.org}

\leavevmode\vadjust pre{\hypertarget{ref-Wickham2019}{}}%
Wickham, Hadley, Averick, M., Bryan, J., Chang, W., McGowan, L. D.,
François, R., Grolemund, G., et al. (2019). Welcome to the {tidyverse}.
\emph{Journal of Open Source Software}, \emph{4}(43), 1686.
doi:\href{https://doi.org/10.21105/joss.01686}{10.21105/joss.01686}

\leavevmode\vadjust pre{\hypertarget{ref-Wilkinson2005-oz}{}}%
Wilkinson, L. (2005). \emph{The grammar of graphics}. Springer.

\end{CSLReferences}

\end{document}

### 2022-05-05 Paper/paper.Rmd
paper first draft done

@@
 journal: JOSS
+nocite: '@mosaicData'
 ---
@@
 wide_df <- tribble(~restaurant, ~food_rating, ~service_rating, ~ambience_rating, ~overall_rating,
-        "R1", 4, 3, 4, 4,
-        "R2", 4, 5, 4, 4,
-        "R3", 3, 4, 5, 3, 
-        "R4", 2, 4, 4, 3, 
-        "R5", 3, 4, 4, 3)
+        "R1", 4.3, 3.4, 4.3, 4.9,
+        "R2", 4.3, 5.0, 4.5, 4.4,
+        "R3", 3.2, 4.4, 5.0, 3.0, 
+        "R4", 2.3, 4.6, 4.4, 3.8, 
+        "R5", 3.9, 4.8, 4.2, 3.3)
 
@@
 
-abund_df <- tribble(~Site, ~`Soil dry mass`, ~`Moss`, ~Alopcune, ~Arctlute, ~Pardpull, ~Trocterr, ~Zoraspin,
-                    1, 2.3321, 3.0445, 10, 0, 45, 57, 4,
-                    2, 3.0493, 1.0986, 2, 0, 37, 65, 9,
-                    3, 2.5572, 2.3979, 20, 0, 45, 66, 1)
-
+# abund_df <- tribble(~Site, ~`Soil dry mass`, ~`Moss`, ~Alopcune, ~Arctlute, ~Pardpull, ~Trocterr, ~Zoraspin,
+#                     1, 2.3321, 3.0445, 10, 0, 45, 57, 4,
+#                     2, 3.0493, 1.0986, 2, 0, 37, 65, 9,
+#                     3, 2.5572, 2.3979, 20, 0, 45, 66, 1)
+# 
 table_styling <- function(df, ...) {
@@
 }
+# 
+# data("spider", package = "mvabund")
 
-data("spider", package = "mvabund")
+data(SnowGR, package = "mosaicData")
+
 ```
@@
 The first column shows the restaurant ID, and the next four columns show the average ratings (out of 5) for food, service, ambience and overall, respectively.') %>%
-  add_header_above(c(" ", "Rating" = 4))
+  add_header_above(c(" ", "Average rating" = 4))
 ```
@@
 ```{r tab2, echo = FALSE}
-table_styling(long_df, col.names = c("Restauant", "Rating type", "Rating"),
-              caption = 'Another form for the restaurant rating data in Table \\ref{tab:tab1}. In @Wickham2014-gy, this format is called the "molten" data.')
+table_styling(long_df, col.names = c("Restauant", "Rating type", "Average rating"),
+              caption = 'Another form for the restaurant rating data in Table \\ref{tab:tab1}. In Wickham (2014), this format is called the "molten" data.')
 ```
@@
 ```{r tab3, echo = FALSE}
-table_styling(abund_df,
-              caption = 'Spider abundance data with environmental covariates.
-The rows correspond to the site, the first two columns are environmental covariates that measure the soil dry mass and cover moss, and the following five columns shows the abundance of the species.') %>%
-  add_header_above(c(" ", "Environment covariates" = 2, "Species abundance" = 5))
+head(SnowGR[, 1:11]) %>% 
+  mutate(Jan = as.character(Jan)) %>% 
+  replace_na(list(Jan = ".")) %>% 
+  table_styling(caption = 'The first 6 rows and 11 columns of the snowfall data for Grand Rapids, Michigan in the `mosaicData` R pacakge (Prium, Kaplan \\& Horton, 2021).',
+                align = "r")
 ```
@@
 
-In this section we demonstrate the use of the `ggmatplot` package and contrast the specification with `ggplot2` with data wrangling using `dplyr` and `tidyr` [@Wickham2019] using the example data in Tables \ref{tab:tab1} and \ref{tab:tab3}, which are stored in the objects `wide_df` and `abun_df`, respectively.
+In this section we demonstrate the use of the `ggmatplot` package and contrast the specification with `ggplot2` after data wrangling using `dplyr` and `tidyr` [@Wickham2019] using the example data in Tables \ref{tab:tab1} and \ref{tab:tab3}, which are stored in the objects `wide_df` and `SnowGR`, respectively.
 
@@
 
-```{r matplot1, fig.height = 2}
+The code below constructs a line plot (superimposed with a point) of the various types (food, service and ambience) of rating, contained in columns 2 to 4 of `wide_df`, against the overall restaurant rating in column 5 of `wide_df` as shown in Figure 1.
+
+```{r matplot2, fig.height = 2, fig.cap = "Line plot of the food, service and ambience rating against the overall restaurant rating."}
 library(ggmatplot)
-ggmatplot(x = wide_df[, -1], plot_type = "both",
-          xlab = "Restaurant",  ylab = "Rating", legend_title = "Type")
+ggmatplot(x = wide_df[, 2:4], y = wide_df[, 5], plot_type = "both",
+          xlab = "Rating",  ylab = "Overall rating", legend_title = "Type") 
 ```
 
-```{r ggplot1, eval = FALSE}
+In contrast to the above, the data must be wrangled to a long form first before plotting in the code below to obtain the same figure as Figure 1.
+This adds a small, but noticable, friction to the workflow for the user that is looking to quickly explore their data.
+
+```{r ggplot2, fig.height = 2, eval = FALSE}
 library(ggplot2)
@@
 wide_df %>% 
-  pivot_longer(contains("rating"), 
+  select(contains("rating")) %>% 
+  pivot_longer(-overall_rating, 
                names_to = "rating_type",
                values_to = "rating") %>% 
-  ggplot(aes(restaurant, rating, color = rating_type)) + 
-  geom_point() +
-  geom_line(aes(group = rating_type,
-                linetype = rating_type))
+  ggplot(aes(rating, overall_rating, color = rating_type)) + 
+  geom_point(aes(shape = rating_type)) +
+  geom_line(aes(group = rating_type, linetype = rating_type)) +
+  labs(x = "Restaurant", y = "Rating", 
+       color = "Type", linetype = "Type", shape = "Type")
 ```
@@
 
-```{r matplot2, fig.height = 2}
-ggmatplot(x = wide_df[, 2:4], y = wide_df[, 5], plot_type = "both")
+The example code draws the boxplot of each column of amount of snowfall across months in the `SnowGR` data as shown in Figure 2.
+As the resulting object is a `ggplot` object, the user can leverage the `ggplot` functions to modify the output (e.g.
+removal of the legend).
+
+```{r matplot3, fig.height = 2, fig.cap="The distribution of the amount of snowfall at Grand Rapids, Michigan, across months from 1893-2011."}
+library(ggmatplot)
+ggmatplot(x = SnowGR[, 3:14], plot_type = "boxplot",
+          xlab = "Month",  ylab = "Snowfall") +
+  guides(color = "none", fill = "none")
+```
+
+The equivalent code for the above to produce Figure 2 without using `ggmatplot` is given below.
+There is again a slight friction in putting the data in the right format before plotting.
+The original wide data format like those shown in Table \ref{tab:tab3} is common so an analyst who has to repeat these tasks can benefit from a quick approach.
+
+```{r ggplot3, fig.height = 2, eval = FALSE}
+library(ggplot2)
+library(tidyr) # or library(tidyverse)
+SnowGR %>% 
+  pivot_longer(Jul:Jun, 
+               names_to = "Month",
+               values_to = "Snowfall") %>% 
+  ggplot(aes(Month, Snowfall)) + 
+  geom_boxplot(aes(color = Month, fill = Month), alpha = 0.5) +
+  guides(color = "none", fill = "none")
 ```

### 2022-05-05 Paper/paper.tex
paper first draft done

@@
 
-The layered grammar of graphics (H. Wickham, 2010), implemented as the
-\texttt{ggplot2} package (Hadley Wickham, 2016) in the statistical
-language R (R Core Team, 2021), is a powerful and popular tool to create
-versatile statistical graphics. This graphical system, however, requires
-input data to be organised in a manner that a data column is mapped to
-an aesthetic element (e.g.~x-coordinate, y-coordinate, color, size),
-which create friction in constructing plots with an aesthetic element
-that span multiple columns in the original data by requiring users to
+The layered grammar of graphics (Wickham, 2010), implemented as the
+\texttt{ggplot2} package (Wickham, 2016) in the statistical language R
+(R Core Team, 2021), is a powerful and popular tool to create versatile
+statistical graphics. This graphical system, however, requires input
+data to be organised in a manner that a data column is mapped to an
+aesthetic element (e.g.~x-coordinate, y-coordinate, color, size), which
+create friction in constructing plots with an aesthetic element that
+span multiple columns in the original data by requiring users to
 re-organise the data.
@@
 generally works well where data is tidied in a rectangular form,
-referred to as ``tidy data'' (Hadley Wickham, 2014), where each row
-represents an observational unit, each column represents a variable, and
-each cell represents a value. In some cases, what constitutes a variable
-(or observational unit), hence a column (or row), in a tidy data can be
+referred to as ``tidy data'' (Wickham, 2014), where each row represents
+an observational unit, each column represents a variable, and each cell
+represents a value. In some cases, what constitutes a variable (or
+observational unit), hence a column (or row), in a tidy data can be
 dependent upon interpretation or downstream interest (e.g.~Tables
@@
 \toprule
-\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Rating} \\
+\multicolumn{1}{c}{ } & \multicolumn{4}{c}{Average rating} \\
 \cmidrule(l{3pt}r{3pt}){2-5}
@@
 \midrule
-R1 & 4 & 3 & 4 & 4\\
-R2 & 4 & 5 & 4 & 4\\
-R3 & 3 & 4 & 5 & 3\\
-R4 & 2 & 4 & 4 & 3\\
-R5 & 3 & 4 & 4 & 3\\
+R1 & 4.3 & 3.4 & 4.3 & 4.9\\
+R2 & 4.3 & 5.0 & 4.5 & 4.4\\
+R3 & 3.2 & 4.4 & 5.0 & 3.0\\
+R4 & 2.3 & 4.6 & 4.4 & 3.8\\
+R5 & 3.9 & 4.8 & 4.2 & 3.3\\
 \bottomrule
@@
 
-\caption{\label{tab:tab2}Another form for the restaurant rating data in Table \ref{tab:tab1}. In @Wickham2014-gy, this format is called the "molten" data.}
+\caption{\label{tab:tab2}Another form for the restaurant rating data in Table \ref{tab:tab1}. In Wickham (2014), this format is called the "molten" data.}
 \centering
@@
 \toprule
-Restauant & Rating type & Rating\\
+Restauant & Rating type & Average rating\\
 \midrule
-R1 & food & 4\\
-R1 & service & 3\\
-R1 & ambience & 4\\
-R1 & overall & 4\\
-R2 & food & 4\\
-R2 & service & 5\\
-R2 & ambience & 4\\
-R2 & overall & 4\\
-R3 & food & 3\\
-R3 & service & 4\\
-R3 & ambience & 5\\
-R3 & overall & 3\\
-R4 & food & 2\\
-R4 & service & 4\\
-R4 & ambience & 4\\
-R4 & overall & 3\\
-R5 & food & 3\\
-R5 & service & 4\\
-R5 & ambience & 4\\
-R5 & overall & 3\\
+R1 & food & 4.3\\
+R1 & service & 3.4\\
+R1 & ambience & 4.3\\
+R1 & overall & 4.9\\
+R2 & food & 4.3\\
+R2 & service & 5.0\\
+R2 & ambience & 4.5\\
+R2 & overall & 4.4\\
+R3 & food & 3.2\\
+R3 & service & 4.4\\
+R3 & ambience & 5.0\\
+R3 & overall & 3.0\\
+R4 & food & 2.3\\
+R4 & service & 4.6\\
+R4 & ambience & 4.4\\
+R4 & overall & 3.8\\
+R5 & food & 3.9\\
+R5 & service & 4.8\\
+R5 & ambience & 4.2\\
+R5 & overall & 3.3\\
 \bottomrule
@@
 
-\caption{\label{tab:tab3}Spider abundance data with environmental covariates.
-The rows correspond to the site, the first two columns are environmental covariates that measure the soil dry mass and cover moss, and the following five columns shows the abundance of the species.}
+\caption{\label{tab:tab3}The first 6 rows and 11 columns of the snowfall data for Grand Rapids, Michigan in the `mosaicData` R pacakge (Prium, Kaplan \& Horton, 2021).}
 \centering
-\begin{tabular}[t]{rrrrrrrr}
+\begin{tabular}[t]{rrrrrrrrrrr}
 \toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Environment covariates} & \multicolumn{5}{c}{Species abundance} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-8}
-Site & Soil dry mass & Moss & Alopcune & Arctlute & Pardpull & Trocterr & Zoraspin\\
+SeasonStart & SeasonEnd & Jul & Aug & Sep & Oct & Nov & Dec & Jan & Feb & Mar\\
 \midrule
-1 & 2.3321 & 3.0445 & 10 & 0 & 45 & 57 & 4\\
-2 & 3.0493 & 1.0986 & 2 & 0 & 37 & 65 & 9\\
-3 & 2.5572 & 2.3979 & 20 & 0 & 45 & 66 & 1\\
+1893 & 1894 & 0 & 0 & 0 & 0.0 & 8.0 & 24.9 & 12.5 & 6.8 & 4.8\\
+1894 & 1895 & 0 & 0 & 0 & 0.0 & 7.5 & 5.3 & 21.5 & 8.0 & 22.5\\
+1895 & 1896 & 0 & 0 & 0 & 0.4 & 23.2 & 15.0 & . & 8.5 & 2.0\\
+1896 & 1897 & 0 & 0 & 0 & 0.2 & 8.0 & 8.0 & 4.9 & 11.2 & 12.0\\
+1897 & 1898 & 0 & 0 & 0 & 0.0 & 1.4 & 8.0 & 15.5 & 29.5 & 0.0\\
+1898 & 1899 & 0 & 0 & 0 & 0.0 & 18.5 & 18.0 & 20 & 3.4 & 16.0\\
 \bottomrule
@@
 In this section we demonstrate the use of the \texttt{ggmatplot} package
-and contrast the specification with \texttt{ggplot2} with data wrangling
-using \texttt{dplyr} and \texttt{tidyr} (Hadley Wickham et al., 2019)
+and contrast the specification with \texttt{ggplot2} after data
+wrangling using \texttt{dplyr} and \texttt{tidyr} (Wickham et al., 2019)
 using the example data in Tables \ref{tab:tab1} and \ref{tab:tab3},
-which are stored in the objects \texttt{wide\_df} and \texttt{abun\_df},
+which are stored in the objects \texttt{wide\_df} and \texttt{SnowGR},
 respectively.
@@
 
+The code below constructs a line plot (superimposed with a point) of the
+various types (food, service and ambience) of rating, contained in
+columns 2 to 4 of \texttt{wide\_df}, against the overall restaurant
+rating in column 5 of \texttt{wide\_df} as shown in Figure 1.
+
 \begin{Shaded}
@@
 \FunctionTok{library}\NormalTok{(ggmatplot)}
-\FunctionTok{ggmatplot}\NormalTok{(}\AttributeTok{x =}\NormalTok{ wide\_df[, }\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{], }\AttributeTok{plot\_type =} \StringTok{"both"}\NormalTok{,}
-          \AttributeTok{xlab =} \StringTok{"Restaurant"}\NormalTok{,  }\AttributeTok{ylab =} \StringTok{"Rating"}\NormalTok{, }\AttributeTok{legend\_title =} \StringTok{"Type"}\NormalTok{)}
+\FunctionTok{ggmatplot}\NormalTok{(}\AttributeTok{x =}\NormalTok{ wide\_df[, }\DecValTok{2}\SpecialCharTok{:}\DecValTok{4}\NormalTok{], }\AttributeTok{y =}\NormalTok{ wide\_df[, }\DecValTok{5}\NormalTok{], }\AttributeTok{plot\_type =} \StringTok{"both"}\NormalTok{,}
+          \AttributeTok{xlab =} \StringTok{"Rating"}\NormalTok{,  }\AttributeTok{ylab =} \StringTok{"Overall rating"}\NormalTok{, }\AttributeTok{legend\_title =} \StringTok{"Type"}\NormalTok{) }
 \end{Highlighting}
@@
 
-\includegraphics{paper_files/figure-latex/matplot1-1.pdf}
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/matplot2-1.pdf}
+\caption{Line plot of the food, service and ambience rating against the
+overall restaurant rating.}
+\end{figure}
+
+In contrast to the above, the data must be wrangled to a long form first
+before plotting in the code below to obtain the same figure as Figure 1.
+This adds a small, but noticable, friction to the workflow for the user
+that is looking to quickly explore their data.
 
@@
 \NormalTok{wide\_df }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{pivot\_longer}\NormalTok{(}\FunctionTok{contains}\NormalTok{(}\StringTok{"rating"}\NormalTok{), }
+  \FunctionTok{select}\NormalTok{(}\FunctionTok{contains}\NormalTok{(}\StringTok{"rating"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{pivot\_longer}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{overall\_rating, }
                \AttributeTok{names\_to =} \StringTok{"rating\_type"}\NormalTok{,}
                \AttributeTok{values\_to =} \StringTok{"rating"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(restaurant, rating, }\AttributeTok{color =}\NormalTok{ rating\_type)) }\SpecialCharTok{+} 
-  \FunctionTok{geom\_point}\NormalTok{() }\SpecialCharTok{+}
-  \FunctionTok{geom\_line}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{group =}\NormalTok{ rating\_type,}
-                \AttributeTok{linetype =}\NormalTok{ rating\_type))}
+  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(rating, overall\_rating, }\AttributeTok{color =}\NormalTok{ rating\_type)) }\SpecialCharTok{+} 
+  \FunctionTok{geom\_point}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{shape =}\NormalTok{ rating\_type)) }\SpecialCharTok{+}
+  \FunctionTok{geom\_line}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{group =}\NormalTok{ rating\_type, }\AttributeTok{linetype =}\NormalTok{ rating\_type)) }\SpecialCharTok{+}
+  \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Restaurant"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Rating"}\NormalTok{, }
+       \AttributeTok{color =} \StringTok{"Type"}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"Type"}\NormalTok{, }\AttributeTok{shape =} \StringTok{"Type"}\NormalTok{)}
 \end{Highlighting}
@@
 
+The example code draws the boxplot of each column of amount of snowfall
+across months in the \texttt{SnowGR} data as shown in Figure 2. As the
+resulting object is a \texttt{ggplot} object, the user can leverage the
+\texttt{ggplot} functions to modify the output (e.g.~removal of the
+legend).
+
 \begin{Shaded}
 \begin{Highlighting}[]
-\FunctionTok{ggmatplot}\NormalTok{(}\AttributeTok{x =}\NormalTok{ wide\_df[, }\DecValTok{2}\SpecialCharTok{:}\DecValTok{4}\NormalTok{], }\AttributeTok{y =}\NormalTok{ wide\_df[, }\DecValTok{5}\NormalTok{], }\AttributeTok{plot\_type =} \StringTok{"both"}\NormalTok{)}
+\FunctionTok{library}\NormalTok{(ggmatplot)}
+\FunctionTok{ggmatplot}\NormalTok{(}\AttributeTok{x =}\NormalTok{ SnowGR[, }\DecValTok{3}\SpecialCharTok{:}\DecValTok{14}\NormalTok{], }\AttributeTok{plot\_type =} \StringTok{"boxplot"}\NormalTok{,}
+          \AttributeTok{xlab =} \StringTok{"Month"}\NormalTok{,  }\AttributeTok{ylab =} \StringTok{"Snowfall"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{guides}\NormalTok{(}\AttributeTok{color =} \StringTok{"none"}\NormalTok{, }\AttributeTok{fill =} \StringTok{"none"}\NormalTok{)}
 \end{Highlighting}
@@
 
-\includegraphics{paper_files/figure-latex/matplot2-1.pdf}
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/matplot3-1.pdf}
+\caption{The distribution of the amount of snowfall at Grand Rapids,
+Michigan, across months from 1893-2011.}
+\end{figure}
+
+The equivalent code for the above to produce Figure 2 without using
+\texttt{ggmatplot} is given below. There is again a slight friction in
+putting the data in the right format before plotting. The original wide
+data format like those shown in Table \ref{tab:tab3} is common so an
+analyst who has to repeat these tasks can benefit from a quick approach.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{library}\NormalTok{(ggplot2)}
+\FunctionTok{library}\NormalTok{(tidyr) }\CommentTok{\# or library(tidyverse)}
+\NormalTok{SnowGR }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{pivot\_longer}\NormalTok{(Jul}\SpecialCharTok{:}\NormalTok{Jun, }
+               \AttributeTok{names\_to =} \StringTok{"Month"}\NormalTok{,}
+               \AttributeTok{values\_to =} \StringTok{"Snowfall"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(Month, Snowfall)) }\SpecialCharTok{+} 
+  \FunctionTok{geom\_boxplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{color =}\NormalTok{ Month, }\AttributeTok{fill =}\NormalTok{ Month), }\AttributeTok{alpha =} \FloatTok{0.5}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{guides}\NormalTok{(}\AttributeTok{color =} \StringTok{"none"}\NormalTok{, }\AttributeTok{fill =} \StringTok{"none"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 \texttt{plot\_type} option. Future development of the package could
-benefit from using a grammar approach, like in Wilkinson (2005) and H.
+benefit from using a grammar approach, like in Wilkinson (2005) and
 Wickham (2010), where plot types can be extensible.
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-mosaicData}{}}%
+Pruim, R., Kaplan, D., \& Horton, N. (2021). \emph{mosaicData: Project
+MOSAIC data sets}. Retrieved from
+\url{https://CRAN.R-project.org/package=mosaicData}
+
 \leavevmode\vadjust pre{\hypertarget{ref-rstats}{}}%
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Wickham2014-gy}{}}%
-Wickham, Hadley. (2014). Tidy data. \emph{Journal of Statistical
-Software}, \emph{59}(10), 1--23.
+Wickham, H. (2014). Tidy data. \emph{Journal of Statistical Software},
+\emph{59}(10), 1--23.
 doi:\href{https://doi.org/10.18637/jss.v059.i10}{10.18637/jss.v059.i10}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Wickham2016}{}}%
-Wickham, Hadley. (2016). \emph{ggplot2: Elegant graphics for data
-analysis}. Springer-Verlag New York. Retrieved from
+Wickham, H. (2016). \emph{ggplot2: Elegant graphics for data analysis}.
+Springer-Verlag New York. Retrieved from
 \url{https://ggplot2.tidyverse.org}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Wickham2019}{}}%
-Wickham, Hadley, Averick, M., Bryan, J., Chang, W., McGowan, L. D.,
+Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D.,
 François, R., Grolemund, G., et al. (2019). Welcome to the {tidyverse}.

### 2022-05-05 Paper/paper.Rmd
Adding edits and comments to JOSS ms

@@
 The layered grammar of graphics [@Wickham2010-kt], implemented as the `ggplot2` package [@Wickham2016] in the statistical language R [@rstats], is a powerful and popular tool to create versatile statistical graphics.
-This graphical system, however, requires input data to be organised in a manner that a data column is mapped to an aesthetic element (e.g.
-x-coordinate, y-coordinate, color, size), which create friction in constructing plots with an aesthetic element that span multiple columns in the original data by requiring users to re-organise the data.
+However, this graphical system requires input data to be organised in a manner that a data column is mapped to an aesthetic element (e.g.
+x-coordinate, y-coordinate, color, size), which creates friction in constructing plots with an aesthetic element that span multiple columns in the original data by requiring users to re-organise the data.
+<!-- Should we use the words long and wide format data frames here, or is that wrong?
+I only ask because I (maybe others?) think about ggplot requiring long format data, but we want to wide formatting?
+-->
 
 The `ggmatplot`, built upon `ggplot2`, is an R-package that allows quick plotting across the columns of matrices or data with the result returned as a `ggplot` object.
-The package is inspired by the function `matplot()` in the core R `graphics` system, thus `ggmatplot` can be considered as a `ggplot` version of `matplot` with the benefits of customising the plots as any other `ggplot` objects via `ggplot2` functions.
+The package is inspired by the function `matplot()` in the core R `graphics` system -- as such, `ggmatplot` may be considered as a `ggplot` version of `matplot` with the benefits of customising the plots as any other `ggplot` objects via `ggplot2` functions, as well as offering several other plotting types that are not immediately available from `matplot` directly, such as comparative violin plots.
 
@@
 Input data to construct plots with `ggplot2` require data to be organised in a manner that maps data columns to aesthetic elements.
-This generally works well where data is tidied in a rectangular form, referred to as "tidy data" [@Wickham2014-gy], where each row represents an observational unit, each column represents a variable, and each cell represents a value.
-In some cases, what constitutes a variable (or observational unit), hence a column (or row), in a tidy data can be dependent upon interpretation or downstream interest (e.g.
+This generally works well where data is tidied in a long rectangular form, often referred to as "tidy data" [@Wickham2014-gy], where each row represents an observational unit, each column represents a variable, and each cell represents a value.
+In some cases, what constitutes a variable (or observational unit), and hence a column (or row), in tidy data can be dependent upon interpretation or downstream interest (e.g.
 Tables \ref{tab:tab1} and \ref{tab:tab2} can be both considered as tidy data), but a clear violation of tidy data principles is when the column names contain data values, e.g.
-Table \ref{tab:tab3} contain the name of the species across a number of column names.
+*Table \ref{tab:tab3} contain the name of the species across a number of column names.* THIS IS WRONG?
+THERE ARE NO SPECIES HERE?
 
@@
 
-The organisation of the data is largely dependent on the downstream analysis and there is no one correct way to do this.
+The organisation of the data is largely dependent on the downstream analysis, and there is no one correct way to do this.
 Some forms of multivariate data, e.g.
-Table \ref{tab:tab3}, are prevalent in the field because it aligns as an input data for a modelling software and/or the format is more convenient for input or view of the data in spreadsheet format.
-However, this format is not consistent with the required format for `ggplot2`, and consequently, plotting with `ggplot2` interrupts the workflow of a user that is trying to quickly visualise these types of data.
-The `ggmatplot` R-package provides a solution to this common friction in producing plots with `ggplot2`.
+Table \ref{tab:tab3}, are prevalent in many scientific fields because it aligns with the input data for a particular modelling software, and/or the format is more convenient for input or view of the data in spreadsheet format (say).
+Unfortunately, this format is not consistent with the required format for `ggplot2`, and consequently plotting with `ggplot2` interrupts the workflow of a user that is trying to quickly visualise these types of data (as part of their exploratory data analysis, say).
+The `ggmatplot` R-package seeks to provide a solution to this common friction in producing plots with `ggplot2`.
 
@@
 
-In this section we demonstrate the use of the `ggmatplot` package and contrast the specification with `ggplot2` after data wrangling using `dplyr` and `tidyr` [@Wickham2019] using the example data in Tables \ref{tab:tab1} and \ref{tab:tab3}, which are stored in the objects `wide_df` and `SnowGR`, respectively.
+In this section, we demonstrate the use of the `ggmatplot` package and contrast the specification with `ggplot2` after data wrangling using `dplyr` and `tidyr` [@Wickham2019].
+We will use the example data in Tables \ref{tab:tab1} and \ref{tab:tab3}, which are stored in the objects `wide_df` and `SnowGR`, respectively.
 
@@
 
-The code below constructs a line plot (superimposed with a point) of the various types (food, service and ambience) of rating, contained in columns 2 to 4 of `wide_df`, against the overall restaurant rating in column 5 of `wide_df` as shown in Figure 1.
+The code below constructs a line plot (superimposed with a point) of the various types (food, service and ambience) of ratings, contained in columns 2 to 4 of `wide_df`, against the overall restaurant rating in column 5 of `wide_df` as shown in Figure 1.
 
@@
 
-In contrast to the above, the data must be wrangled to a long form first before plotting in the code below to obtain the same figure as Figure 1.
-This adds a small, but noticable, friction to the workflow for the user that is looking to quickly explore their data.
+In contrast to the above, using `ggplot2` alone, the data must be wrangled to a long form first before plotting, as exemplified in the code below, in order to obtain the same result as Figure 1.
+This adds a small, but noticeable, friction to the workflow for the practitioner that is looking to promptly explore their data.
 
@@
 removal of the legend).
+CAN WE `fct_inorder` SO THAT THE MONTHS APPEAR IN ORDER OF THEIR COLUMN NAMES?
+EMI HAS ALREADY PICKED THIS UP AS AN ISSUE...
 
@@
 The equivalent code for the above to produce Figure 2 without using `ggmatplot` is given below.
-There is again a slight friction in putting the data in the right format before plotting.
-The original wide data format like those shown in Table \ref{tab:tab3} is common so an analyst who has to repeat these tasks can benefit from a quick approach.
+Again, we observe a slight but non-negligible friction in putting the data in the right format prior to plotting.
+The original wide data format like those shown in Table \ref{tab:tab3} is common in the environmental sciences among other disciplines, and thus an analyst who has to repeat these tasks can benefit from a quick approach as `ggmatplot` offers.
 
@@
 
+WOULD IT BENEFIT FROM ALSO SHOWING A PLOT THAT BASE MATPLOT CAN NOT?
+THE ABOVE CAN BE DONE BY APPLY `boxplot`, AND HENCE WHY I WROTE VIOLIN PLOT AT THE BEGINNING AS IT IS NOT AVAILABLE AT LEAST IN CORE R?
+
 # Discussion
 
-The `ggmatplot` R-package provides a solution to a common friction to quickly plotting multivariate data where the primary interest is mapping the column names as an aesthetic element.
-The solution provided however is a recipe-driven approach where the user can only produce plot types as many there are included in the `plot_type` option.
-Future development of the package could benefit from using a grammar approach, like in @Wilkinson2005-oz and @Wickham2010-kt, where plot types can be extensible.
+The `ggmatplot` R-package provides a solution to a common friction encountered when wanting to quickly plot multivariate data, where the primary interest is mapping the column names as an aesthetic element.
+While an excellent start, we also acknowledge that solution provided is a recipe-driven approach, where the user can only produce plot types as many there are included in the `plot_type` option.
+Future developments of the package could benefit from using a grammar approach, like in @Wilkinson2005-oz and @Wickham2010-kt, where plot types can be extensible.
 
@@
 
-FKCH was supported by ARC DECRA XXX.
+FKCH was supported by an Australian Research Council Discovery Fellowship DE200100435.
 

### 2022-05-05 Paper/paper.tex
Adding edits and comments to JOSS ms

@@
 \renewcommand{\headrulewidth}{0pt}
-\fancyhead[L]{\hspace{-0.75cm}\includegraphics[width=5.5cm]{/Library/Frameworks/R.framework/Versions/4.1/Resources/library/rticles/rmarkdown/templates/joss/resources/JOSS-logo.png}}
+\fancyhead[L]{\hspace{-0.75cm}\includegraphics[width=5.5cm]{/home/fkch/R/x86_64-pc-linux-gnu-library/3.6/rticles/rmarkdown/templates/joss/resources/JOSS-logo.png}}
 \fancyhead[C]{}
@@
 \urlstyle{same}  % don't use monospace font for urls
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\setcounter{secnumdepth}{0}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+
+% Pandoc syntax highlighting
 \usepackage{color}
@@
 \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
+
+% tightlist command for lists without linebreak
 \providecommand{\tightlist}{%
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
 
 % Pandoc citation processing
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
 \newlength{\csllabelwidth}
 \setlength{\csllabelwidth}{3em}
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
 % for Pandoc 2.8 to 2.10.1
@@
   % turn on hanging indent if param 1 is 1
-  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  \ifodd #1
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
+  \fi
   % set entry spacing
-  \ifnum #2 > 0
-  \setlength{\parskip}{#2\baselineskip}
-  \fi
+  \setlength{\parskip}{#2\cslentryspacingunit}
  }%
  {}
-\usepackage{calc} % for calculating minipage widths
+\usepackage{calc}
 \newcommand{\CSLBlock}[1]{#1\hfill\break}
@@
 (R Core Team, 2021), is a powerful and popular tool to create versatile
-statistical graphics. This graphical system, however, requires input
-data to be organised in a manner that a data column is mapped to an
-aesthetic element (e.g.~x-coordinate, y-coordinate, color, size), which
-create friction in constructing plots with an aesthetic element that
-span multiple columns in the original data by requiring users to
-re-organise the data.
+statistical graphics. However, this graphical system requires input data
+to be organised in a manner that a data column is mapped to an aesthetic
+element (e.g.~x-coordinate, y-coordinate, color, size), which creates
+friction in constructing plots with an aesthetic element that span
+multiple columns in the original data by requiring users to re-organise
+the data.\\
 
@@
 by the function \texttt{matplot()} in the core R \texttt{graphics}
-system, thus \texttt{ggmatplot} can be considered as a \texttt{ggplot}
-version of \texttt{matplot} with the benefits of customising the plots
-as any other \texttt{ggplot} objects via \texttt{ggplot2} functions.
+system -- as such, \texttt{ggmatplot} may be considered as a
+\texttt{ggplot} version of \texttt{matplot} with the benefits of
+customising the plots as any other \texttt{ggplot} objects via
+\texttt{ggplot2} functions, as well as offering several other plotting
+types that are not immediately available from \texttt{matplot} directly,
+such as comparative violin plots.
 
@@
 organised in a manner that maps data columns to aesthetic elements. This
-generally works well where data is tidied in a rectangular form,
-referred to as ``tidy data'' (Wickham, 2014), where each row represents
-an observational unit, each column represents a variable, and each cell
-represents a value. In some cases, what constitutes a variable (or
-observational unit), hence a column (or row), in a tidy data can be
-dependent upon interpretation or downstream interest (e.g.~Tables
+generally works well where data is tidied in a long rectangular form,
+often referred to as ``tidy data'' (Wickham, 2014), where each row
+represents an observational unit, each column represents a variable, and
+each cell represents a value. In some cases, what constitutes a variable
+(or observational unit), and hence a column (or row), in tidy data can
+be dependent upon interpretation or downstream interest (e.g.~Tables
 \ref{tab:tab1} and \ref{tab:tab2} can be both considered as tidy data),
 but a clear violation of tidy data principles is when the column names
-contain data values, e.g.~Table \ref{tab:tab3} contain the name of the
-species across a number of column names.
+contain data values, e.g.~\emph{Table \ref{tab:tab3} contain the name of
+the species across a number of column names.} THIS IS WRONG? THERE ARE
+NO SPECIES HERE?
 
@@
 The organisation of the data is largely dependent on the downstream
-analysis and there is no one correct way to do this. Some forms of
-multivariate data, e.g.~Table \ref{tab:tab3}, are prevalent in the field
-because it aligns as an input data for a modelling software and/or the
-format is more convenient for input or view of the data in spreadsheet
-format. However, this format is not consistent with the required format
-for \texttt{ggplot2}, and consequently, plotting with \texttt{ggplot2}
-interrupts the workflow of a user that is trying to quickly visualise
-these types of data. The \texttt{ggmatplot} R-package provides a
-solution to this common friction in producing plots with
-\texttt{ggplot2}.
+analysis, and there is no one correct way to do this. Some forms of
+multivariate data, e.g.~Table \ref{tab:tab3}, are prevalent in many
+scientific fields because it aligns with the input data for a particular
+modelling software, and/or the format is more convenient for input or
+view of the data in spreadsheet format (say). Unfortunately, this format
+is not consistent with the required format for \texttt{ggplot2}, and
+consequently plotting with \texttt{ggplot2} interrupts the workflow of a
+user that is trying to quickly visualise these types of data (as part of
+their exploratory data analysis, say). The \texttt{ggmatplot} R-package
+seeks to provide a solution to this common friction in producing plots
+with \texttt{ggplot2}.
 
@@
 
-In this section we demonstrate the use of the \texttt{ggmatplot} package
-and contrast the specification with \texttt{ggplot2} after data
-wrangling using \texttt{dplyr} and \texttt{tidyr} (Wickham et al., 2019)
-using the example data in Tables \ref{tab:tab1} and \ref{tab:tab3},
-which are stored in the objects \texttt{wide\_df} and \texttt{SnowGR},
-respectively.
+In this section, we demonstrate the use of the \texttt{ggmatplot}
+package and contrast the specification with \texttt{ggplot2} after data
+wrangling using \texttt{dplyr} and \texttt{tidyr} (Wickham et al.,
+2019). We will use the example data in Tables \ref{tab:tab1} and
+\ref{tab:tab3}, which are stored in the objects \texttt{wide\_df} and
+\texttt{SnowGR}, respectively.
 
@@
 The code below constructs a line plot (superimposed with a point) of the
-various types (food, service and ambience) of rating, contained in
+various types (food, service and ambience) of ratings, contained in
 columns 2 to 4 of \texttt{wide\_df}, against the overall restaurant
@@
 
-In contrast to the above, the data must be wrangled to a long form first
-before plotting in the code below to obtain the same figure as Figure 1.
-This adds a small, but noticable, friction to the workflow for the user
-that is looking to quickly explore their data.
+In contrast to the above, using \texttt{ggplot2} alone, the data must be
+wrangled to a long form first before plotting, as exemplified in the
+code below, in order to obtain the same result as Figure 1. This adds a
+small, but noticeable, friction to the workflow for the practitioner
+that is looking to promptly explore their data.
 
@@
 \texttt{ggplot} functions to modify the output (e.g.~removal of the
-legend).
+legend). CAN WE \texttt{fct\_inorder} SO THAT THE MONTHS APPEAR IN ORDER
+OF THEIR COLUMN NAMES? EMI HAS ALREADY PICKED THIS UP AS AN
+ISSUE\ldots{}
 
@@
 The equivalent code for the above to produce Figure 2 without using
-\texttt{ggmatplot} is given below. There is again a slight friction in
-putting the data in the right format before plotting. The original wide
-data format like those shown in Table \ref{tab:tab3} is common so an
-analyst who has to repeat these tasks can benefit from a quick approach.
+\texttt{ggmatplot} is given below. Again, we observe a slight but
+non-negligible friction in putting the data in the right format prior to
+plotting. The original wide data format like those shown in Table
+\ref{tab:tab3} is common in the environmental sciences among other
+disciplines, and thus an analyst who has to repeat these tasks can
+benefit from a quick approach as \texttt{ggmatplot} offers.
 
@@
 
+WOULD IT BENEFIT FROM ALSO SHOWING A PLOT THAT BASE MATPLOT CAN NOT? THE
+ABOVE CAN BE DONE BY APPLY \texttt{boxplot}, AND HENCE WHY I WROTE
+VIOLIN PLOT AT THE BEGINNING AS IT IS NOT AVAILABLE AT LEAST IN CORE R?
+
 \hypertarget{discussion}{%
@@
 The \texttt{ggmatplot} R-package provides a solution to a common
-friction to quickly plotting multivariate data where the primary
-interest is mapping the column names as an aesthetic element. The
-solution provided however is a recipe-driven approach where the user can
-only produce plot types as many there are included in the
-\texttt{plot\_type} option. Future development of the package could
-benefit from using a grammar approach, like in Wilkinson (2005) and
-Wickham (2010), where plot types can be extensible.
+friction encountered when wanting to quickly plot multivariate data,
+where the primary interest is mapping the column names as an aesthetic
+element. While an excellent start, we also acknowledge that solution
+provided is a recipe-driven approach, where the user can only produce
+plot types as many there are included in the \texttt{plot\_type} option.
+Future developments of the package could benefit from using a grammar
+approach, like in Wilkinson (2005) and Wickham (2010), where plot types
+can be extensible.
 
@@
 
-FKCH was supported by ARC DECRA XXX.
+FKCH was supported by an Australian Research Council Discovery
+Fellowship DE200100435.
 

### 2022-05-07 R/ggmatplot.R
Removing default coloring by groups for violin and boxplots

@@ -368,9 +368,18 @@ ggmatplot <- function(x, y, plot_type = "point", color = NULL, fill = NULL,
 
   # if both color and fill values are not defined, use default values
   if (is.null(color) & is.null(fill)) {
-    p <- p +
-      scale_fill_discrete(name = legend_title, labels = legend_label) +
-      scale_color_discrete(name = legend_title, labels = legend_label)
+    # removing default coloring by groups for violin and boxplots
+    if(plot_type %in% c("violin","boxplot")){
+      p <- p +
+        scale_fill_manual(name = legend_title, labels = legend_label, values = rep("white",numGroups)) +
+        scale_color_manual(name = legend_title, labels = legend_label, values = rep("black",numGroups)) +
+        theme(legend.position = "none")
+    } else{
+      p <- p +
+        scale_fill_discrete(name = legend_title, labels = legend_label) +
+        scale_color_discrete(name = legend_title, labels = legend_label)
+    }
+
   }
 
   if (!is.null(shape)) {

### 2022-05-07 R/ggmatplot.R
Removed violin and boxplot x axis alphabetic ordering

@@ -244,7 +244,7 @@ ggmatplot <- function(x, y, plot_type = "point", color = NULL, fill = NULL,
         "geom_violin",
         c(
           list(mapping = aes(
-            x = Group,
+            x = factor(Group, levels = unique(Group)),
             y = .data[[yname]]
           )),
           params
@@ -257,7 +257,7 @@ ggmatplot <- function(x, y, plot_type = "point", color = NULL, fill = NULL,
         "geom_boxplot",
         c(
           list(mapping = aes(
-            x = Group,
+            x = factor(Group, levels = unique(Group)),
             y = .data[[yname]]
           )),
           params

### 2022-05-07 Paper/paper.Rmd
add comments in the draft

@@
 
+*Regarding the question from Francis, I don't know if long format or wide format data is well defined in the literature.
+If it is, we definitely can mention it.
+If not, do we need to define it?
+--commented by XL*
+
 The `ggmatplot`, built upon `ggplot2`, is an R-package that allows quick plotting across the columns of matrices or data with the result returned as a `ggplot` object.
@@
   replace_na(list(Jan = ".")) %>% 
-  table_styling(caption = 'The first 6 rows and 11 columns of the snowfall data for Grand Rapids, Michigan in the `mosaicData` R pacakge (Prium, Kaplan \\& Horton, 2021).',
+  table_styling(caption = 'The first 6 rows and 11 columns of the snowfall data for Grand Rapids, Michigan in the R pacakge  \texttt{mosaicData} (Prium, Kaplan \\& Horton, 2021).',
                 align = "r")
@@
 The code below constructs a line plot (superimposed with a point) of the various types (food, service and ambience) of ratings, contained in columns 2 to 4 of `wide_df`, against the overall restaurant rating in column 5 of `wide_df` as shown in Figure 1.
+
+*In example 1, should X lab be restaurant or one type rating?
+The one made by ggmatplot is different from the one made by ggplot.
+Currently, the plot does not show a clear relationship between various types and the overall rating.
+Is it better to compare the ratings among restaurants?
+-- commented by XL*
 
@@
 
+*Now the color and fill has been removed by default for boxplot by default, there is no legend. We may need to change the demonstration of Example 2. -- commented by XL*
+*Another question is that only the github version has such setting. Do I need to submit the new version to CRAN? -- commented by XL*
+
 ```{r matplot3, fig.height = 2, fig.cap="The distribution of the amount of snowfall at Grand Rapids, Michigan, across months from 1893-2011."}
@@
 
+*The other option is the density plot which can not be done by matplot--commented by XL. But since there are too many months, it might not be a good idea. -- commented by XL*
+
 # Discussion

### 2022-05-07 Paper/paper.tex
add comments in the draft

@@
 \renewcommand{\headrulewidth}{0pt}
-\fancyhead[L]{\hspace{-0.75cm}\includegraphics[width=5.5cm]{/home/fkch/R/x86_64-pc-linux-gnu-library/3.6/rticles/rmarkdown/templates/joss/resources/JOSS-logo.png}}
+\fancyhead[L]{\hspace{-0.75cm}\includegraphics[width=5.5cm]{/Library/Frameworks/R.framework/Versions/4.1/Resources/library/rticles/rmarkdown/templates/joss/resources/JOSS-logo.png}}
 \fancyhead[C]{}
@@
 
+\emph{Regarding the question from Francis, I don't know if long format
+or wide format data is well defined in the literature. If it is, we
+definitely can mention it. If not, do we need to define it? --commented
+by XL}
+
 The \texttt{ggmatplot}, built upon \texttt{ggplot2}, is an R-package
@@
 
-\caption{\label{tab:tab3}The first 6 rows and 11 columns of the snowfall data for Grand Rapids, Michigan in the `mosaicData` R pacakge (Prium, Kaplan \& Horton, 2021).}
+\caption{\label{tab:tab3}The first 6 rows and 11 columns of the snowfall data for Grand Rapids, Michigan in the R pacakge   exttt{mosaicData} (Prium, Kaplan \& Horton, 2021).}
 \centering
@@
 rating in column 5 of \texttt{wide\_df} as shown in Figure 1.
+
+\emph{In example 1, should X lab be restaurant or one type rating? The
+one made by ggmatplot is different from the one made by ggplot.
+Currently, the plot does not show a clear relationship between various
+types and the overall rating. Is it better to compare the ratings among
+restaurants? -- commented by XL}
 
@@
 
+\emph{Now the color and fill has been removed by default for boxplot by
+default, there is no legend. We may need to change the demonstration of
+Example 2. -- commented by XL} \emph{Another question is that only the
+github version has such setting. Do I need to submit the new version to
+CRAN? -- commented by XL}
+
 \begin{Shaded}
@@
 
+\emph{The other option is the density plot which can not be done by
+matplot--commented by XL. But since there are too many months, it might
+not be a good idea. -- commented by XL}
+
 \hypertarget{discussion}{%
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-mosaicData}{}}%
+\leavevmode\hypertarget{ref-mosaicData}{}%
 Pruim, R., Kaplan, D., \& Horton, N. (2021). \emph{mosaicData: Project
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-rstats}{}}%
+\leavevmode\hypertarget{ref-rstats}{}%
 R Core Team. (2021). \emph{R: A language and environment for statistical
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Wickham2010-kt}{}}%
+\leavevmode\hypertarget{ref-Wickham2010-kt}{}%
 Wickham, H. (2010). A layered grammar of graphics. \emph{Journal of
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Wickham2014-gy}{}}%
+\leavevmode\hypertarget{ref-Wickham2014-gy}{}%
 Wickham, H. (2014). Tidy data. \emph{Journal of Statistical Software},
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Wickham2016}{}}%
+\leavevmode\hypertarget{ref-Wickham2016}{}%
 Wickham, H. (2016). \emph{ggplot2: Elegant graphics for data analysis}.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Wickham2019}{}}%
+\leavevmode\hypertarget{ref-Wickham2019}{}%
 Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D.,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Wilkinson2005-oz}{}}%
+\leavevmode\hypertarget{ref-Wilkinson2005-oz}{}%
 Wilkinson, L. (2005). \emph{The grammar of graphics}. Springer.

### 2022-05-07 Paper/paper.Rmd
modify the comments in the draft

@@
 
-*Regarding the question from Francis, I don't know if long format or wide format data is well defined in the literature.
+*Regarding Francis' question, I don't know if the long format or wide format data are well defined in the literature.
 If it is, we definitely can mention it.
@@
 
-*In example 1, should X lab be restaurant or one type rating?
+*In example 1, should X lab be restaurant or rating?
 The one made by ggmatplot is different from the one made by ggplot.
@@
 
-*Now the color and fill has been removed by default for boxplot by default, there is no legend. We may need to change the demonstration of Example 2. -- commented by XL*
+*Now the color and fill has been removed for boxplot. So by default, there is no legend. We may need to change the demonstration of Example 2. -- commented by XL*
 *Another question is that only the github version has such setting. Do I need to submit the new version to CRAN? -- commented by XL*

### 2022-05-07 Paper/paper.tex
modify the comments in the draft

@@
 
-\emph{Regarding the question from Francis, I don't know if long format
-or wide format data is well defined in the literature. If it is, we
+\emph{Regarding Francis' question, I don't know if the long format or
+wide format data are well defined in the literature. If it is, we
 definitely can mention it. If not, do we need to define it? --commented
@@
 
-\emph{In example 1, should X lab be restaurant or one type rating? The
-one made by ggmatplot is different from the one made by ggplot.
-Currently, the plot does not show a clear relationship between various
-types and the overall rating. Is it better to compare the ratings among
-restaurants? -- commented by XL}
+\emph{In example 1, should X lab be restaurant or rating? The one made
+by ggmatplot is different from the one made by ggplot. Currently, the
+plot does not show a clear relationship between various types and the
+overall rating. Is it better to compare the ratings among restaurants?
+-- commented by XL}
 
@@
 
-\emph{Now the color and fill has been removed by default for boxplot by
+\emph{Now the color and fill has been removed for boxplot. So by
 default, there is no legend. We may need to change the demonstration of

### 2022-05-09 Paper/paper.Rmd
Updating ms and adding more comments that MUST be addressed

--- 2022-05-05 Paper/paper.Rmd
@@
 x-coordinate, y-coordinate, color, size), which creates friction in constructing plots with an aesthetic element that span multiple columns in the original data by requiring users to re-organise the data.
-<!-- Should we use the words long and wide format data frames here, or is that wrong?
+<!-- FKCH: Should we use the words long and wide format data frames here, or is that wrong?
 I only ask because I (maybe others?) think about ggplot requiring long format data, but we want to wide formatting?
+
+XL: Regarding Francis' question, I don't know if the long format or wide format data are well defined in the literature.
+If it is, we definitely can mention it.
+If not, do we need to define it?
+--commented by XL
+
+FKCH: @Emi you know the literature best. Can you make a decision on this? Thanks!
 -->
@@
 Tables \ref{tab:tab1} and \ref{tab:tab2} can be both considered as tidy data), but a clear violation of tidy data principles is when the column names contain data values, e.g.
-*Table \ref{tab:tab3} contain the name of the species across a number of column names.* THIS IS WRONG?
-THERE ARE NO SPECIES HERE?
+Table \ref{tab:tab3} contains months of the year across a number of column names.
 
@@
 
-The code below constructs a line plot (superimposed with a point) of the various types (food, service and ambience) of ratings, contained in columns 2 to 4 of `wide_df`, against the overall restaurant rating in column 5 of `wide_df` as shown in Figure 1.
+The code below constructs a line plot (superimposed with a point) of the various types (food, service and ambience) of ratings, contained in columns 2 to 4 of `wide_df`, against the overall rating in column 5 of `wide_df` as shown in Figure 1.
 
-```{r matplot2, fig.height = 2, fig.cap = "Line plot of the food, service and ambience rating against the overall restaurant rating."}
+<!-- Something is wrong in ggmatplot CRAN version with order of legend!!!
+The legend is displaying things in order of appearance, but the lines are plotting in alphabetical order.
+See ggplot below for correct order -->
+
+```{r matplot2, fig.height = 2, fig.cap = "Line plot of the food, service, and ambience ratings versus overall rating, for five restaurants."}
 library(ggmatplot)
-ggmatplot(x = wide_df[, 2:4], y = wide_df[, 5], plot_type = "both",
-          xlab = "Rating",  ylab = "Overall rating", legend_title = "Type") 
+ggmatplot(x = wide_df[, 5], y = wide_df[, 2:4], plot_type = "both",
+          xlab = "Overall rating",  ylab = "Rating", legend_title = "Type") 
 ```
@@
                values_to = "rating") %>% 
-  ggplot(aes(rating, overall_rating, color = rating_type)) + 
+  ggplot(aes(x = overall_rating, y = rating, color = rating_type)) + 
   geom_point(aes(shape = rating_type)) +
@@
 removal of the legend).
-CAN WE `fct_inorder` SO THAT THE MONTHS APPEAR IN ORDER OF THEIR COLUMN NAMES?
-EMI HAS ALREADY PICKED THIS UP AS AN ISSUE...
 
@@
 ggmatplot(x = SnowGR[, 3:14], plot_type = "boxplot",
-          xlab = "Month",  ylab = "Snowfall") +
-  guides(color = "none", fill = "none")
+          xlab = "Month",  ylab = "Snowfall")
 ```
@@
                values_to = "Snowfall") %>% 
+  mutate(Month = fct_inorder(Month)) %>% 
   ggplot(aes(Month, Snowfall)) + 
-  geom_boxplot(aes(color = Month, fill = Month), alpha = 0.5) +
-  guides(color = "none", fill = "none")
+  geom_boxplot()
 ```
 
-WOULD IT BENEFIT FROM ALSO SHOWING A PLOT THAT BASE MATPLOT CAN NOT?
+<!-- FKCH: WOULD IT BENEFIT FROM ALSO SHOWING A PLOT THAT BASE MATPLOT CAN NOT?
 THE ABOVE CAN BE DONE BY APPLY `boxplot`, AND HENCE WHY I WROTE VIOLIN PLOT AT THE BEGINNING AS IT IS NOT AVAILABLE AT LEAST IN CORE R?
+
+XL: The other option is the density plot which can not be done by matplot--commented by XL. But since there are too many months, it might not be a good idea. -
+-->
 

### 2022-05-09 Paper/paper.tex
Updating ms and adding more comments that MUST be addressed

--- 2022-05-05 Paper/paper.tex
@@
 but a clear violation of tidy data principles is when the column names
-contain data values, e.g.~\emph{Table \ref{tab:tab3} contain the name of
-the species across a number of column names.} THIS IS WRONG? THERE ARE
-NO SPECIES HERE?
+contain data values, e.g.~Table \ref{tab:tab3} contains months of the
+year across a number of column names.
 
@@
 various types (food, service and ambience) of ratings, contained in
-columns 2 to 4 of \texttt{wide\_df}, against the overall restaurant
-rating in column 5 of \texttt{wide\_df} as shown in Figure 1.
+columns 2 to 4 of \texttt{wide\_df}, against the overall rating in
+column 5 of \texttt{wide\_df} as shown in Figure 1.
 
@@
 \FunctionTok{library}\NormalTok{(ggmatplot)}
-\FunctionTok{ggmatplot}\NormalTok{(}\AttributeTok{x =}\NormalTok{ wide\_df[, }\DecValTok{2}\SpecialCharTok{:}\DecValTok{4}\NormalTok{], }\AttributeTok{y =}\NormalTok{ wide\_df[, }\DecValTok{5}\NormalTok{], }\AttributeTok{plot\_type =} \StringTok{"both"}\NormalTok{,}
-          \AttributeTok{xlab =} \StringTok{"Rating"}\NormalTok{,  }\AttributeTok{ylab =} \StringTok{"Overall rating"}\NormalTok{, }\AttributeTok{legend\_title =} \StringTok{"Type"}\NormalTok{) }
+\FunctionTok{ggmatplot}\NormalTok{(}\AttributeTok{x =}\NormalTok{ wide\_df[, }\DecValTok{5}\NormalTok{], }\AttributeTok{y =}\NormalTok{ wide\_df[, }\DecValTok{2}\SpecialCharTok{:}\DecValTok{4}\NormalTok{], }\AttributeTok{plot\_type =} \StringTok{"both"}\NormalTok{,}
+          \AttributeTok{xlab =} \StringTok{"Overall rating"}\NormalTok{,  }\AttributeTok{ylab =} \StringTok{"Rating"}\NormalTok{, }\AttributeTok{legend\_title =} \StringTok{"Type"}\NormalTok{) }
 \end{Highlighting}
@@
 \includegraphics{paper_files/figure-latex/matplot2-1.pdf}
-\caption{Line plot of the food, service and ambience rating against the
-overall restaurant rating.}
+\caption{Line plot of the food, service, and ambience ratings versus
+overall rating, for five restaurants.}
 \end{figure}
@@
                \AttributeTok{values\_to =} \StringTok{"rating"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(rating, overall\_rating, }\AttributeTok{color =}\NormalTok{ rating\_type)) }\SpecialCharTok{+} 
+  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ overall\_rating, }\AttributeTok{y =}\NormalTok{ rating, }\AttributeTok{color =}\NormalTok{ rating\_type)) }\SpecialCharTok{+}
   \FunctionTok{geom\_point}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{shape =}\NormalTok{ rating\_type)) }\SpecialCharTok{+}
@@
 \texttt{ggplot} functions to modify the output (e.g.~removal of the
-legend). CAN WE \texttt{fct\_inorder} SO THAT THE MONTHS APPEAR IN ORDER
-OF THEIR COLUMN NAMES? EMI HAS ALREADY PICKED THIS UP AS AN
-ISSUE\ldots{}
+legend).
 
@@
 \FunctionTok{ggmatplot}\NormalTok{(}\AttributeTok{x =}\NormalTok{ SnowGR[, }\DecValTok{3}\SpecialCharTok{:}\DecValTok{14}\NormalTok{], }\AttributeTok{plot\_type =} \StringTok{"boxplot"}\NormalTok{,}
-          \AttributeTok{xlab =} \StringTok{"Month"}\NormalTok{,  }\AttributeTok{ylab =} \StringTok{"Snowfall"}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{guides}\NormalTok{(}\AttributeTok{color =} \StringTok{"none"}\NormalTok{, }\AttributeTok{fill =} \StringTok{"none"}\NormalTok{)}
+          \AttributeTok{xlab =} \StringTok{"Month"}\NormalTok{,  }\AttributeTok{ylab =} \StringTok{"Snowfall"}\NormalTok{)}
 \end{Highlighting}
@@
                \AttributeTok{values\_to =} \StringTok{"Snowfall"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Month =} \FunctionTok{fct\_inorder}\NormalTok{(Month)) }\SpecialCharTok{\%\textgreater{}\%} 
   \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(Month, Snowfall)) }\SpecialCharTok{+} 
-  \FunctionTok{geom\_boxplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{color =}\NormalTok{ Month, }\AttributeTok{fill =}\NormalTok{ Month), }\AttributeTok{alpha =} \FloatTok{0.5}\NormalTok{) }\SpecialCharTok{+}
-  \FunctionTok{guides}\NormalTok{(}\AttributeTok{color =} \StringTok{"none"}\NormalTok{, }\AttributeTok{fill =} \StringTok{"none"}\NormalTok{)}
+  \FunctionTok{geom\_boxplot}\NormalTok{()}
 \end{Highlighting}
 \end{Shaded}
-
-WOULD IT BENEFIT FROM ALSO SHOWING A PLOT THAT BASE MATPLOT CAN NOT? THE
-ABOVE CAN BE DONE BY APPLY \texttt{boxplot}, AND HENCE WHY I WROTE
-VIOLIN PLOT AT THE BEGINNING AS IT IS NOT AVAILABLE AT LEAST IN CORE R?
 

### 2022-05-09 Paper/paper.Rmd
minor changes

@@
 FKCH: @Emi you know the literature best. Can you make a decision on this? Thanks!
+
+ET: I don't think we should mention long/wide -- there's more nuance than that. 
 -->
@@
   replace_na(list(Jan = ".")) %>% 
-  table_styling(caption = 'The first 6 rows and 11 columns of the snowfall data for Grand Rapids, Michigan in the R pacakge  \texttt{mosaicData} (Prium, Kaplan \\& Horton, 2021).',
+  table_styling(caption = 'The first 6 rows and 11 columns of the snowfall data for Grand Rapids, Michigan in the R pacakge  \\texttt{mosaicData} (Prium, Kaplan \\& Horton, 2021).',
                 align = "r")
@@
 The legend is displaying things in order of appearance, but the lines are plotting in alphabetical order.
-See ggplot below for correct order -->
+See ggplot below for correct order
+
+ET: fixed in the latest github version.
+
+-->
 
@@
 As the resulting object is a `ggplot` object, the user can leverage the `ggplot` functions to modify the output (e.g.
-removal of the legend).
+addition of a title).
 
@@
 ggmatplot(x = SnowGR[, 3:14], plot_type = "boxplot",
-          xlab = "Month",  ylab = "Snowfall")
+          xlab = "Month",  ylab = "Snowfall") + 
+          ggtitle("Grand Rapids, Michigan, 1893-2011")
 ```
@@
 library(ggplot2)
-library(tidyr) # or library(tidyverse)
+library(tidyr) 
+library(forcats) # or library(tidyverse)
 SnowGR %>% 
@@
   ggplot(aes(Month, Snowfall)) + 
-  geom_boxplot()
+  geom_boxplot() + 
+  ggtitle("Grand Rapids, Michigan, 1893-2011")
 ```
@@
 XL: The other option is the density plot which can not be done by matplot--commented by XL. But since there are too many months, it might not be a good idea. -
+
+ET: violin plot doesn't look good here so I'd rather stick with boxplot.
+I don't think we need to emphasise how it can provide more plots than matplot because it already offers an advantage of being a ggplot object, and thus leverage existing ggplot customisation
 -->

### 2022-05-09 Paper/paper.tex
minor changes

@@
 \urlstyle{same}  % don't use monospace font for urls
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\setcounter{secnumdepth}{0}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-% Pandoc syntax highlighting
 \usepackage{color}
@@
 \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-
-% tightlist command for lists without linebreak
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
 \providecommand{\tightlist}{%
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{0}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 
 % Pandoc citation processing
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
 \newlength{\cslhangindent}
 \setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslentryspacingunit} % times entry-spacing
-\setlength{\cslentryspacingunit}{\parskip}
 % for Pandoc 2.8 to 2.10.1
@@
   % turn on hanging indent if param 1 is 1
-  \ifodd #1
-  \let\oldpar\par
-  \def\par{\hangindent=\cslhangindent\oldpar}
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
   \fi
-  % set entry spacing
-  \setlength{\parskip}{#2\cslentryspacingunit}
  }%
  {}
-\usepackage{calc}
+\usepackage{calc} % for calculating minipage widths
 \newcommand{\CSLBlock}[1]{#1\hfill\break}
@@
 
-\caption{\label{tab:tab3}The first 6 rows and 11 columns of the snowfall data for Grand Rapids, Michigan in the R pacakge   exttt{mosaicData} (Prium, Kaplan \& Horton, 2021).}
+\caption{\label{tab:tab3}The first 6 rows and 11 columns of the snowfall data for Grand Rapids, Michigan in the R pacakge  \texttt{mosaicData} (Prium, Kaplan \& Horton, 2021).}
 \centering
@@
 resulting object is a \texttt{ggplot} object, the user can leverage the
-\texttt{ggplot} functions to modify the output (e.g.~removal of the
-legend).
+\texttt{ggplot} functions to modify the output (e.g.~addition of a
+title).
 
@@
 \FunctionTok{ggmatplot}\NormalTok{(}\AttributeTok{x =}\NormalTok{ SnowGR[, }\DecValTok{3}\SpecialCharTok{:}\DecValTok{14}\NormalTok{], }\AttributeTok{plot\_type =} \StringTok{"boxplot"}\NormalTok{,}
-          \AttributeTok{xlab =} \StringTok{"Month"}\NormalTok{,  }\AttributeTok{ylab =} \StringTok{"Snowfall"}\NormalTok{)}
+          \AttributeTok{xlab =} \StringTok{"Month"}\NormalTok{,  }\AttributeTok{ylab =} \StringTok{"Snowfall"}\NormalTok{) }\SpecialCharTok{+} 
+          \FunctionTok{ggtitle}\NormalTok{(}\StringTok{"Grand Rapids, Michigan, 1893{-}2011"}\NormalTok{)}
 \end{Highlighting}
@@
 \FunctionTok{library}\NormalTok{(ggplot2)}
-\FunctionTok{library}\NormalTok{(tidyr) }\CommentTok{\# or library(tidyverse)}
+\FunctionTok{library}\NormalTok{(tidyr) }
+\FunctionTok{library}\NormalTok{(forcats) }\CommentTok{\# or library(tidyverse)}
 \NormalTok{SnowGR }\SpecialCharTok{\%\textgreater{}\%} 
@@
   \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(Month, Snowfall)) }\SpecialCharTok{+} 
-  \FunctionTok{geom\_boxplot}\NormalTok{()}
+  \FunctionTok{geom\_boxplot}\NormalTok{() }\SpecialCharTok{+} 
+  \FunctionTok{ggtitle}\NormalTok{(}\StringTok{"Grand Rapids, Michigan, 1893{-}2011"}\NormalTok{)}
 \end{Highlighting}

### 2022-05-09 Paper/paper.Rmd
Minor changes to ms

@@
 The layered grammar of graphics [@Wickham2010-kt], implemented as the `ggplot2` package [@Wickham2016] in the statistical language R [@rstats], is a powerful and popular tool to create versatile statistical graphics.
-However, this graphical system requires input data to be organised in a manner that a data column is mapped to an aesthetic element (e.g.
-x-coordinate, y-coordinate, color, size), which creates friction in constructing plots with an aesthetic element that span multiple columns in the original data by requiring users to re-organise the data.
+However, this graphical system requires input data to be organised in a manner that a data column is mapped to an aesthetic element (e.g., x-coordinate, y-coordinate, color, size), which creates friction in constructing plots with an aesthetic element that spans multiple columns in the original data, as it requires the user to reorganise the data.
 <!-- FKCH: Should we use the words long and wide format data frames here, or is that wrong?
@@
 The `ggmatplot`, built upon `ggplot2`, is an R-package that allows quick plotting across the columns of matrices or data with the result returned as a `ggplot` object.
-The package is inspired by the function `matplot()` in the core R `graphics` system -- as such, `ggmatplot`  may be considered as a `ggplot` version of `matplot` with the benefits of customising the plots as any other `ggplot` objects via `ggplot2` functions, as well as offering several other plotting types that are not immediately available from `matplot` directly, such as comparative violin plots.
+The package is inspired by the function `matplot()` in the core R `graphics` system -- as such, `ggmatplot` may be considered as a `ggplot` version of `matplot`, with the benefits of customising the plots as any other `ggplot` objects via `ggplot2` functions, as well as offering several other plotting types that are not immediately available from `matplot` directly e.g., comparative violin plots.
 
@@
 This generally works well where data is tidied in a long rectangular form, often referred to as "tidy data" [@Wickham2014-gy], where each row represents an observational unit, each column represents a variable, and each cell represents a value.
-In some cases, what constitutes a variable (or observational unit), and hence a column (or row), in tidy data can be dependent upon interpretation or downstream interest (e.g.
-Tables \ref{tab:tab1} and \ref{tab:tab2} can be both considered as tidy data), but a clear violation of tidy data principles is when the column names contain data values, e.g.
+However, in some cases what constitutes a variable (or observational unit), and hence a column (or row), in tidy data can be dependent upon interpretation or downstream interest (e.g.
+Tables \ref{tab:tab1} and \ref{tab:tab2} can both be considered as tidy data), but a clear violation of tidy data principles is when the column names contain data values, e.g.
 Table \ref{tab:tab3} contains months of the year across a number of column names.
@@
 
-The organisation of the data is largely dependent on the downstream analysis, and there is no one correct way to do this.
+The organisation of the data is largely dependent on the subsequent analysis, and there is no one correct way to do this.
 Some forms of multivariate data, e.g.
 Table \ref{tab:tab3}, are prevalent in many scientific fields because it aligns with the input data for a particular modelling software, and/or the format is more convenient for input or view of the data in spreadsheet format (say).
-Unfortunately, this format is not consistent with the required format for `ggplot2`, and consequently plotting with `ggplot2` interrupts the workflow of a user that is trying to quickly visualise these types of data (as part of their exploratory data analysis, say).
-The `ggmatplot` R-package seeks to provide a solution to this common friction in producing plots with `ggplot2`.
+Unfortunately, this format is not consistent with the required format for `ggplot2`, and consequently plotting with `ggplot2` interrupts the workflow of a user that is trying to quickly visualise these types of data (as part of their exploratory data analysis, for example).
+The `ggmatplot` R-package seeks to provide a solution to this common friction.
+<!-- in producing plots with `ggplot2`.  -->
 
@@
 
-In contrast to the above, using `ggplot2` alone, the data must be wrangled to a long form first before plotting, as exemplified in the code below, in order to obtain the same result as Figure 1.
-This adds a small, but noticeable, friction to the workflow for the practitioner that is looking to promptly explore their data.
+In contrast to the above, using `ggplot2` alone to obtain the same result as Figure 1 requires the data be wrangled to a long form first before plotting.
+This is exemplified in the code below, from which we identify a small, but noticeable, friction to the workflow for the practitioner that is looking to promptly explore their data.
 
@@
 
-The example code draws the boxplot of each column of amount of snowfall across months in the `SnowGR` data as shown in Figure 2.
-As the resulting object is a `ggplot` object, the user can leverage the `ggplot` functions to modify the output (e.g.
-addition of a title).
+The example code draws the boxplot of each column of amount of snowfall across months in the `SnowGR` data, as presented in Figure 2.
+As the resulting object is a `ggplot` object, the user can leverage the `ggplot` functions to modify the output (e.g., addition of a title).
 
@@
 
-The equivalent code for the above to produce Figure 2 without using `ggmatplot` is given below.
+The equivalent code to produce Figure 2 without using `ggmatplot` is given below.
 Again, we observe a slight but non-negligible friction in putting the data in the right format prior to plotting.
-The original wide data format like those shown in Table \ref{tab:tab3} is common in the environmental sciences among other disciplines, and thus an analyst who has to repeat these tasks can benefit from a quick approach as `ggmatplot` offers.
+The original wide data format as seen in Table \ref{tab:tab3} is common in the environmental sciences, among other disciplines, and thus an analyst who has to repeat these tasks can benefit from a quick approach as `ggmatplot` offers.
 
@@
 The `ggmatplot` R-package provides a solution to a common friction encountered when wanting to quickly plot multivariate data, where the primary interest is mapping the column names as an aesthetic element.
-While an excellent start, we also acknowledge that solution provided is a recipe-driven approach, where the user can only produce plot types as many there are included in the `plot_type` option.
+While a important start, we also acknowledge that solution provided is a recipe-driven approach, where the user can only produce plot types as many there are included in the `plot_type` option.
 Future developments of the package could benefit from using a grammar approach, like in @Wilkinson2005-oz and @Wickham2010-kt, where plot types can be extensible.

### 2022-05-09 Paper/paper.tex
Minor changes to ms

@@
 to be organised in a manner that a data column is mapped to an aesthetic
-element (e.g.~x-coordinate, y-coordinate, color, size), which creates
-friction in constructing plots with an aesthetic element that span
-multiple columns in the original data by requiring users to re-organise
-the data.\\
+element (e.g., x-coordinate, y-coordinate, color, size), which creates
+friction in constructing plots with an aesthetic element that spans
+multiple columns in the original data, as it requires the user to
+reorganise the data.\\
 
@@
 system -- as such, \texttt{ggmatplot} may be considered as a
-\texttt{ggplot} version of \texttt{matplot} with the benefits of
+\texttt{ggplot} version of \texttt{matplot}, with the benefits of
 customising the plots as any other \texttt{ggplot} objects via
 \texttt{ggplot2} functions, as well as offering several other plotting
-types that are not immediately available from \texttt{matplot} directly,
-such as comparative violin plots.
+types that are not immediately available from \texttt{matplot} directly
+e.g., comparative violin plots.
 
@@
 represents an observational unit, each column represents a variable, and
-each cell represents a value. In some cases, what constitutes a variable
-(or observational unit), and hence a column (or row), in tidy data can
-be dependent upon interpretation or downstream interest (e.g.~Tables
-\ref{tab:tab1} and \ref{tab:tab2} can be both considered as tidy data),
-but a clear violation of tidy data principles is when the column names
-contain data values, e.g.~Table \ref{tab:tab3} contains months of the
-year across a number of column names.
+each cell represents a value. However, in some cases what constitutes a
+variable (or observational unit), and hence a column (or row), in tidy
+data can be dependent upon interpretation or downstream interest
+(e.g.~Tables \ref{tab:tab1} and \ref{tab:tab2} can both be considered as
+tidy data), but a clear violation of tidy data principles is when the
+column names contain data values, e.g.~Table \ref{tab:tab3} contains
+months of the year across a number of column names.
 
@@
 
-The organisation of the data is largely dependent on the downstream
+The organisation of the data is largely dependent on the subsequent
 analysis, and there is no one correct way to do this. Some forms of
@@
 user that is trying to quickly visualise these types of data (as part of
-their exploratory data analysis, say). The \texttt{ggmatplot} R-package
-seeks to provide a solution to this common friction in producing plots
-with \texttt{ggplot2}.
+their exploratory data analysis, for example). The \texttt{ggmatplot}
+R-package seeks to provide a solution to this common friction.
 
@@
 
-In contrast to the above, using \texttt{ggplot2} alone, the data must be
-wrangled to a long form first before plotting, as exemplified in the
-code below, in order to obtain the same result as Figure 1. This adds a
-small, but noticeable, friction to the workflow for the practitioner
-that is looking to promptly explore their data.
+In contrast to the above, using \texttt{ggplot2} alone to obtain the
+same result as Figure 1 requires the data be wrangled to a long form
+first before plotting. This is exemplified in the code below, from which
+we identify a small, but noticeable, friction to the workflow for the
+practitioner that is looking to promptly explore their data.
 
@@
 The example code draws the boxplot of each column of amount of snowfall
-across months in the \texttt{SnowGR} data as shown in Figure 2. As the
-resulting object is a \texttt{ggplot} object, the user can leverage the
-\texttt{ggplot} functions to modify the output (e.g.~addition of a
+across months in the \texttt{SnowGR} data, as presented in Figure 2. As
+the resulting object is a \texttt{ggplot} object, the user can leverage
+the \texttt{ggplot} functions to modify the output (e.g., addition of a
 title).
@@
 
-The equivalent code for the above to produce Figure 2 without using
-\texttt{ggmatplot} is given below. Again, we observe a slight but
-non-negligible friction in putting the data in the right format prior to
-plotting. The original wide data format like those shown in Table
-\ref{tab:tab3} is common in the environmental sciences among other
-disciplines, and thus an analyst who has to repeat these tasks can
-benefit from a quick approach as \texttt{ggmatplot} offers.
+The equivalent code to produce Figure 2 without using \texttt{ggmatplot}
+is given below. Again, we observe a slight but non-negligible friction
+in putting the data in the right format prior to plotting. The original
+wide data format as seen in Table \ref{tab:tab3} is common in the
+environmental sciences, among other disciplines, and thus an analyst who
+has to repeat these tasks can benefit from a quick approach as
+\texttt{ggmatplot} offers.
 
@@
 where the primary interest is mapping the column names as an aesthetic
-element. While an excellent start, we also acknowledge that solution
+element. While a important start, we also acknowledge that solution
 provided is a recipe-driven approach, where the user can only produce

### 2022-05-09 Paper/paper.Rmd
minor changes in ms

@@
   geom_line(aes(group = rating_type, linetype = rating_type)) +
-  labs(x = "Restaurant", y = "Rating", 
+  labs(x = "Overall rating", y = "Rating", 
        color = "Type", linetype = "Type", shape = "Type")
@@
 The `ggmatplot` R-package provides a solution to a common friction encountered when wanting to quickly plot multivariate data, where the primary interest is mapping the column names as an aesthetic element.
-While a important start, we also acknowledge that solution provided is a recipe-driven approach, where the user can only produce plot types as many there are included in the `plot_type` option.
+While an important start, we also acknowledge that solution provided is a recipe-driven approach, where the user can only produce plot types as many there are included in the `plot_type` option.
 Future developments of the package could benefit from using a grammar approach, like in @Wilkinson2005-oz and @Wickham2010-kt, where plot types can be extensible.

### 2022-05-09 Paper/paper.tex
minor changes in ms

@@
   \FunctionTok{geom\_line}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{group =}\NormalTok{ rating\_type, }\AttributeTok{linetype =}\NormalTok{ rating\_type)) }\SpecialCharTok{+}
-  \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Restaurant"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Rating"}\NormalTok{, }
+  \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \StringTok{"Overall rating"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Rating"}\NormalTok{, }
        \AttributeTok{color =} \StringTok{"Type"}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"Type"}\NormalTok{, }\AttributeTok{shape =} \StringTok{"Type"}\NormalTok{)}
@@
 where the primary interest is mapping the column names as an aesthetic
-element. While a important start, we also acknowledge that solution
+element. While an important start, we also acknowledge that solution
 provided is a recipe-driven approach, where the user can only produce

### 2022-05-10 R/ggmatplot.R
Modify the legend order alphabetically.

@@ -333,7 +333,7 @@ ggmatplot <- function(x, y, plot_type = "point", color = NULL, fill = NULL,
       ), call. = FALSE)
     }
   } else {
-    legend_label <- unique(data$Group)
+    legend_label <- levels(factor(data$Group))
   }
 
   if (!is.null(color)) {

### 2022-05-10 R/ggmatplot.R
correct the order of manual settings for color, shape, linetype, fill.

@@ -338,6 +338,7 @@ ggmatplot <- function(x, y, plot_type = "point", color = NULL, fill = NULL,
 
   if (!is.null(color)) {
     color <- validateNumParams(color, numGroups)
+    color <- color[order(unique(data$Group))]
     p <- p + scale_color_manual(
       name = legend_title, labels = legend_label,
       values = color
@@ -353,6 +354,7 @@ ggmatplot <- function(x, y, plot_type = "point", color = NULL, fill = NULL,
 
   if (!is.null(fill)) {
     fill <- validateNumParams(fill, numGroups)
+    fill <- fill[order(unique(data$Group))]
     p <- p + scale_fill_manual(
       name = legend_title, labels = legend_label,
       values = fill
@@ -392,6 +394,7 @@ ggmatplot <- function(x, y, plot_type = "point", color = NULL, fill = NULL,
         plot_type
       ), call. = FALSE)
     }
+    shape <- shape[order(unique(data$Group))]
     p <- p + scale_shape_manual(
       name = legend_title, labels = legend_label,
       values = shape
@@ -410,6 +413,7 @@ ggmatplot <- function(x, y, plot_type = "point", color = NULL, fill = NULL,
         plot_type
       ), call. = FALSE)
     }
+    linetype <- linetype[order(unique(data$Group))]
     p <- p + scale_linetype_manual(
       name = legend_title, labels = legend_label,
       values = linetype

### 2022-05-13 R/errorplotstats.R
major overhaul of internal code

@@ -18,7 +18,7 @@
 #' errorplotstats(data, desc_stat = "mean_se")
 errorplotstats <- function(data, desc_stat = "") {
   if (desc_stat == "mean_se") {
-    errorplot_data <- do.call(rbind, by(data, list(data$Group), function(x) {
+    errorplot_data <- do.call(rbind, by(data, list(data[[GROUP_NAME]]), function(x) {
       c(
         y = mean(x$x, na.rm = TRUE),
         ymin = mean(x$x, na.rm = TRUE) -
@@ -28,7 +28,7 @@ errorplotstats <- function(data, desc_stat = "") {
       )
     }))
   } else if (desc_stat == "mean_sd") {
-    errorplot_data <- do.call(rbind, by(data, list(data$Group), function(x) {
+    errorplot_data <- do.call(rbind, by(data, list(data[[GROUP_NAME]]), function(x) {
       c(
         y = mean(x$x, na.rm = TRUE),
         ymin = mean(x$x, na.rm = TRUE) - sd(x$x, na.rm = TRUE),
@@ -36,7 +36,7 @@ errorplotstats <- function(data, desc_stat = "") {
       )
     }))
   } else if (desc_stat == "mean_range") {
-    errorplot_data <- do.call(rbind, by(data, list(data$Group), function(x) {
+    errorplot_data <- do.call(rbind, by(data, list(data[[GROUP_NAME]]), function(x) {
       c(
         y = mean(x$x, na.rm = TRUE),
         ymin = min(x$x, na.rm = TRUE),
@@ -44,7 +44,7 @@ errorplotstats <- function(data, desc_stat = "") {
       )
     }))
   } else if (desc_stat == "median_iqr") {
-    errorplot_data <- do.call(rbind, by(data, list(data$Group), function(x) {
+    errorplot_data <- do.call(rbind, by(data, list(data[[GROUP_NAME]]), function(x) {
       c(
         y = median(x$x, na.rm = TRUE),
         ymin = quantile(x$x, probs = 0.25, names = FALSE),
@@ -52,7 +52,7 @@ errorplotstats <- function(data, desc_stat = "") {
       )
     }))
   } else if (desc_stat == "median_range") {
-    errorplot_data <- do.call(rbind, by(data, list(data$Group), function(x) {
+    errorplot_data <- do.call(rbind, by(data, list(data[[GROUP_NAME]]), function(x) {
       c(
[…]

### 2022-05-13 R/ggmatplot.R

@@ -96,370 +96,122 @@
 #'   legend_label = c("Sepal", "Petal"), legend_title = "",
 #'   xlab = "Length", ylab = "Width"
 #' )
-ggmatplot <- function(x, y, plot_type = "point", color = NULL, fill = NULL,
+ggmatplot <- function(x = NULL, y = NULL,
+                      plot_type = c("point", "line", "both", "density", "histogram",
+                                    "boxplot", "dotplot", "errorplot", "violin", "ecdf"),
+                      color = NULL, fill = NULL,
                       shape = NULL, linetype = NULL, xlim = c(NA, NA),
                       ylim = c(NA, NA), log = NULL, main = NULL, xlab = NULL,
                       ylab = NULL, legend_label = NULL, legend_title = NULL,
                       desc_stat = "mean_se", asp = NA, ...) {
 
   # binding global variables to objects
-  Group <- ymin <- ymax <- NULL
+  Variable <- ymin <- ymax <- NULL
+  params_input <- list(...)
 
-  # valid plot types
-  if (!plot_type %in% c(
-    "point", "line", "both", "density", "histogram",
-    "boxplot", "dotplot", "errorplot", "violin", "ecdf"
-  )) {
-    stop("plot_type can not take this value", call. = FALSE)
-  }
-  if (!missing(x) & !missing(y)) {
-    # only single matrix input allowed for the following plot types
-    if (plot_type %in% c(
-      "density", "histogram", "boxplot", "dotplot",
-      "errorplot", "violin", "ecdf"
-    )) {
-      stop("This plot type only uses a single matrix input", call. = FALSE)
-    }
-    data.list <- matclean(x = x, y = y)
-  } else if (!missing(x)) {
-    data.list <- matclean(x = x)
-  } else {
-    stop("x can not be missing", call. = FALSE)
+  plot_type <- match.arg(plot_type)
+  validate_input(x, y, plot_type, desc_stat,
+                 color, fill, shape, linetype, xlim, ylim,
+                 log, asp)
+
+  data_list <- construct_data_list(x, y)
+  data <- get_data_for_plot(data_list, plot_type, desc_stat)
+  params <- setup_params(params_input, plot_type)
+
+  xname <- data_list$xname
+  yname <- data_list$yname
+
+  plot_geom <- function(geom, aes) {
+    do.call(geom, c(list(mapping = aes), params))
   }
 
-  params <- list(...)
-
[…]

### 2022-05-13 R/matclean.R

@@ -7,7 +7,7 @@
 #' @return A list containing:
 #'  * data - the long format data frame.
 #'  * xname - name(s) of the ID column(s)
-#'  * yname - names of the pivoted columns(columns with group names and values)
+#'  * yname - names of the pivoted columns (columns with group names and values)
 #' @noRd
 #'
 #' @examples
@@ -29,19 +29,19 @@ matclean <- function(x, y) {
   }
 
   if (missing(y)) {
-    xname <- "Observation_number"
+    xname <- OBSERVATION_NAME
     yname <- "x"
-    x$Observation_number <- 1:nrow(x)
-    data <- pivotlonger(x, names_to = "Group", values_to = yname, ncolx + 1)
+    x[[OBSERVATION_NAME]] <- 1:nrow(x)
+    data <- pivotlonger(x, names_to = GROUP_NAME, values_to = yname, ncolx + 1)
   } else if (ncolx > ncoly & ncoly == 1) {
     xname <- "x"
     yname <- colnames(y)
     data <- data.frame(x, y)
     ncol <- ncol(data)
     yname <- colnames(data)[ncol]
-    data$Observation_number <- 1:nrow(data)
+    data[[OBSERVATION_NAME]] <- 1:nrow(data)
     data <- pivotlonger(data,
-      names_to = "Group", values_to = xname,
+      names_to = GROUP_NAME, values_to = xname,
       c(ncol, (ncol + 1))
     )
   } else if (ncoly > ncolx & ncolx == 1) {
@@ -49,29 +49,32 @@ matclean <- function(x, y) {
     yname <- "y"
     data <- data.frame(x, y)
     ncol <- ncol(data)
-    data$Observation_number <- 1:nrow(data)
+    data[[OBSERVATION_NAME]] <- 1:nrow(data)
     data <- pivotlonger(data,
-      names_to = "Group", values_to = yname,
+      names_to = GROUP_NAME, values_to = yname,
       c(1, (ncol + 1))
     )
   } else if (ncolx == ncoly) {
     xname <- "x"
     yname <- "y"
     colnames(x) <- colnames(y) <- paste0("Column ", 1:ncolx)
-    x <- pivotlonger(x, names_to = "Group", values_to = xname)
-    x <- x[with(x, order(Group)), ]
-    x$Observation_number <- 1:nrow(x)
-    y <- pivotlonger(y, names_to = "Group", values_to = yname)
[…]

### 2022-05-13 R/utils.R

@@ -1,0 +1,200 @@
+#' Function to validate the number of parameter values against the number of
+#' groups
+#'
+#' @param param_values A list of values.
+#' @param ngroup An integer denoting the number of unique values of the
+#' grouping variable.
+#' @param strict A logical value to indicate that the number of param_values has
+#'  to match the number of groups. Default is FALSE.
+#'
+#' @return If valid, a list of parameter values. If invalid, an error message.
+#' @noRd
+#'
+validate_nparam <- function(param_values, ngroup, strict = FALSE) {
+  if(is.null(param_values)) {
+    return(param_values)
+  }
+  # set same value for all groups
+  if (!strict & !ngroup == 1 & length(param_values) == 1) {
+    return(rep(param_values, ngroup))
+  }
+  # values > number of unique groups
+  else if (length(param_values) > ngroup) {
+    stop(paste0(
+      "Too many ", substitute(param_values), " values. Only ",
+      ngroup, " needed but ", length(param_values), " provided."
+    ), call. = FALSE)
+  }
+  # values < number of unique groups
+  else if (length(param_values) < ngroup) {
+    stop(paste0(
+      "Insufficient ", substitute(param_values), " values. ",
+      ngroup, " needed but only ", length(param_values), " provided."
+    ), call. = FALSE)
+  }
+  # values = number of unique groups
+  else {
+    return(param_values)
+  }
+}
+
+
+validate_input <- function(x, y, plot_type, desc_stat,
+                           color, fill, shape, linetype, xlim, ylim,
+                           log, asp) {
+  if(!is.null(x) & !is.null(y) & plot_type %in% SINGLE_INPUT_PLOT_TYPE) {
+    stop("This plot type only uses a single matrix input", call. = FALSE)
+  } else if (is.null(x)) {
+    stop("x can not be missing", call. = FALSE)
+  }
+
+  if(desc_stat != "mean_se" & plot_type != "errorplot") {
+    warning(paste0(
+      "desc_stat is an invalid parameter for plot type: ", plot_type),
+      call. = FALSE)
+  }
+
+  if(plot_type == "errorplot" && !(desc_stat %in% STATS_WITH_RANGE)) {
+    stop("desc_stat can not take this value", call. = FALSE)
+  }
+
+  if(!is.null(log) && !log %in% c("x", "y", "xy")) {
[…]

### 2022-05-13 R/validateNumParams.R

@@ -1,34 +1,0 @@
-#' Function to validate the number of parameter values against the number of
-#' groups
-#'
-#' @param parameterValues A list of values.
-#' @param numGroups An integer denoting the number of unique values of the
-#' grouping variable.
-#'
-#' @return If valid, a list of parameter values. If invalid, an error message.
-#' @noRd
-#'
-validateNumParams <- function(parameterValues, numGroups) {
-  # set same value for all groups
-  if (!numGroups == 1 & length(parameterValues) == 1) {
-    return(rep(parameterValues, numGroups))
-  }
-  # values > number of unique groups
-  else if (length(parameterValues) > numGroups) {
-    stop(paste0(
-      "Too many ", substitute(parameterValues), " values. Only ",
-      numGroups, " needed but ", length(parameterValues), " provided."
-    ), call. = FALSE)
-  }
-  # values < number of unique groups
-  else if (length(parameterValues) < numGroups) {
-    stop(paste0(
-      "Insufficient ", substitute(parameterValues), " values. ",
-      numGroups, " needed but only ", length(parameterValues), " provided."
-    ), call. = FALSE)
-  }
-  # values = number of unique groups
-  else {
-    return(parameterValues)
-  }
-}

### 2022-05-13 tests/testthat/test_ggmatplot_parameters.R
fix tests

@@ -161,7 +161,7 @@ test_that("invalid plot types are not allowed", {
   skip_on_os("windows")
   expect_error(
     ggmatplot(x, y, plot_type = "scatterplot"),
-    "plot_type can not take this value"
+    "'arg' should be one of “point”, “line”, “both”, “density”, “histogram”, “boxplot”, “dotplot”, “errorplot”, “violin”, “ecdf”"
   )
 })
 
@@ -270,7 +270,7 @@ test_that("invalid parameter values throw errors", {
   # invalid log value
   expect_error(
     ggmatplot(y, plot_type = "point", log = 1),
-    "invalid log value provided"
+    "invalid `log` value provided"
   )
   # invalid des_stat value
   expect_error(

### 2022-05-13 tests/testthat/test_matclean.R

@@ -1,36 +1,35 @@
-library(tibble)
 
 test_that("converts a single matrix into long format", {
-  x <- tibble(a = 1:3, b = 11:13, c = 31:33)
+  x <- data.frame(a = 1:3, b = 11:13, c = 31:33)
   outputList <- matclean(x)
 
-  expect_equal(outputList$xname, "Observation_number")
+  expect_equal(outputList$xname, "Observation")
   expect_equal(outputList$yname, "x")
-  expect_named(outputList$data, c("Observation_number", "Group", "x"))
+  expect_named(outputList$data, c("Observation", "Group", "x"))
 
 
-  expect_equal(outputList$data$Group, rep(names(x), nrow(x)))
+  expect_equal(outputList$data$Group, factor(rep(names(x), nrow(x))))
   expect_equal(outputList$data$x, c(1, 11, 31, 2, 12, 32, 3, 13, 33))
   expect_equal(
-    outputList$data$Observation_number,
+    outputList$data$Observation,
     unlist(lapply(1:nrow(x), function(var) rep(var, ncol(x))))
   )
 })
 
 test_that("converts two matrices(x and y) into long format
           (ncolx>ncoly & ncoly=1)", {
-  x <- tibble(a = 1:3, b = 11:13, c = 31:33)
-  y <- tibble(d = 51:53)
+  x <- data.frame(a = 1:3, b = 11:13, c = 31:33)
+  y <- data.frame(d = 51:53)
   outputList <- matclean(x, y)
 
   expect_equal(outputList$xname, "x")
   expect_equal(outputList$yname, names(y))
-  expect_named(outputList$data, c(names(y), "Observation_number", "Group", "x"))
+  expect_named(outputList$data, c(names(y), "Observation", "Group", "x"))
 
-  expect_equal(outputList$data$Group, rep(names(x), nrow(x)))
+  expect_equal(outputList$data$Group, factor(rep(names(x), nrow(x))))
   expect_equal(outputList$data$x, c(1, 11, 31, 2, 12, 32, 3, 13, 33))
   expect_equal(
-    outputList$data$Observation_number,
+    outputList$data$Observation,
     unlist(lapply(1:nrow(x), function(var) rep(var, ncol(x))))
   )
   expect_equal(
@@ -41,18 +40,18 @@ test_that("converts two matrices(x and y) into long format
 
[…]

### 2022-05-13 tests/testthat/test_pivotlonger.R

@@ -1,5 +1,5 @@
 test_that("can pivot columns to long format based on a single id column", {
-  df <- tibble(x = 1:3, y = 11:13, z = 31:33)
+  df <- data.frame(x = 1:3, y = 11:13, z = 31:33)
   pv <- pivotlonger(df, names_to = "Group", values_to = "Value", id_cols = 1)
 
   expect_named(pv, c("x", "Group", "Value"))
@@ -8,7 +8,7 @@ test_that("can pivot columns to long format based on a single id column", {
 })
 
 test_that("can pivot columns to long format based on a multiple id columns", {
-  df <- tibble(x = 1:3, y = 11:13, z = 31:33, a = 41:43, b = 51:53)
+  df <- data.frame(x = 1:3, y = 11:13, z = 31:33, a = 41:43, b = 51:53)
   pv <- pivotlonger(df,
     names_to = "Group", values_to = "Value",
     id_cols = c(1, 2, 3)
@@ -30,7 +30,7 @@ test_that("NA values are preserved in both id columns and pivoted columns", {
 })
 
 test_that("id columns can have different data types", {
-  df <- tibble(x = 1:3, y = c("a", "b", "c"), z = 31:33)
+  df <- data.frame(x = 1:3, y = c("a", "b", "c"), z = 31:33)
   pv <- pivotlonger(df,
     names_to = "Group", values_to = "Value",
     id_cols = c(1, 2)
@@ -39,6 +39,6 @@ test_that("id columns can have different data types", {
   expect_named(pv, c("x", "y", "Group", "Value"))
   expect_equal(pv$x, c(1, 2, 3))
   expect_equal(pv$y, c("a", "b", "c"))
-  expect_equal(pv$Group, rep(names(df[, 3]), nrow(df)))
+  expect_equal(pv$Group, rep(names(df[, 3, drop = FALSE]), nrow(df)))
   expect_equal(pv$Value, c(31, 32, 33))
 })

### 2022-05-13 R/utils.R
change to color legend where single color/fill supplied by user (display turns off for legend)

@@ -141,13 +141,22 @@ update_legend_aes <- function(plot, plot_type, color, fill, shape, linetype, leg
       name = legend_title, labels = legend_label,
       values = color
     )
+    ncolor <- length(unique(color))
+    if(ncolor==1) {
+      plot <- plot + guides(color = "none")
+    }
+
     # if color is defined and fill isn't, update both using color values
     if (is.null(fill)) {
       plot <- plot + scale_fill_manual(
         name = legend_title, labels = legend_label,
         values = color
       )
+      if(ncolor==1) {
+        plot <- plot + guides(fill = "none")
+      }
     }
+
   }
 
   if (!is.null(fill)) {
@@ -155,12 +164,19 @@ update_legend_aes <- function(plot, plot_type, color, fill, shape, linetype, leg
       name = legend_title, labels = legend_label,
       values = fill
     )
+    nfill <- length(unique(fill))
+    if(nfill==1) {
+      plot <- plot + guides(fill = "none")
+    }
     # if fill is defined and color isn't, update both using fill values
     if (is.null(color)) {
       plot <- plot + scale_color_manual(
         name = legend_title, labels = legend_label,
         values = fill
       )
+      if(nfill==1) {
+        plot <- plot + guides(color = "none")
+      }
     }
   }
 
@@ -171,7 +187,7 @@ update_legend_aes <- function(plot, plot_type, color, fill, shape, linetype, leg
       plot <- plot +
         scale_fill_manual(name = legend_title, labels = legend_label, values = rep("white", length(legend_label))) +
         scale_color_manual(name = legend_title, labels = legend_label, values = rep("black", length(legend_label))) +
-        theme(legend.position = "none")
+        guides(fill = "none", color = "none")
     } else{
       plot <- plot +
         scale_fill_discrete(name = legend_title, labels = legend_label) +

### 2022-05-13 Paper/paper.tex
updated paper with latest update of package

@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-mosaicData}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-mosaicData}{}}%
 Pruim, R., Kaplan, D., \& Horton, N. (2021). \emph{mosaicData: Project
@@
 
-\leavevmode\hypertarget{ref-rstats}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-rstats}{}}%
 R Core Team. (2021). \emph{R: A language and environment for statistical
@@
 
-\leavevmode\hypertarget{ref-Wickham2010-kt}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Wickham2010-kt}{}}%
 Wickham, H. (2010). A layered grammar of graphics. \emph{Journal of
@@
 
-\leavevmode\hypertarget{ref-Wickham2014-gy}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Wickham2014-gy}{}}%
 Wickham, H. (2014). Tidy data. \emph{Journal of Statistical Software},
@@
 
-\leavevmode\hypertarget{ref-Wickham2016}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Wickham2016}{}}%
 Wickham, H. (2016). \emph{ggplot2: Elegant graphics for data analysis}.
@@
 
-\leavevmode\hypertarget{ref-Wickham2019}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Wickham2019}{}}%
 Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D.,
@@
 
-\leavevmode\hypertarget{ref-Wilkinson2005-oz}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Wilkinson2005-oz}{}}%
 Wilkinson, L. (2005). \emph{The grammar of graphics}. Springer.

### 2022-05-15 R/ggmatplot.R
modify legend settings

@@ -191,8 +191,7 @@ ggmatplot <- function(x = NULL, y = NULL,
   shape <- validate_nparam(shape, nvar)
   linetype <- validate_nparam(linetype, nvar)
 
-  p <- update_legend_aes(p, plot_type, color, fill, shape, linetype, legend_title, legend_label)
-
+  p <- update_legend_aes(p, plot_type, color, fill, shape, linetype, legend_title, legend_label, nvar)
 
   # labels
   if (!is.null(main)) p <- p + ggtitle(main)

### 2022-05-15 R/utils.R

@@ -135,14 +135,14 @@ setup_params <- function(params_input, plot_type) {
 }
 
 
-update_legend_aes <- function(plot, plot_type, color, fill, shape, linetype, legend_title, legend_label) {
+update_legend_aes <- function(plot, plot_type, color, fill, shape, linetype, legend_title, legend_label, nvar) {
   if (!is.null(color)) {
     plot <- plot + scale_color_manual(
       name = legend_title, labels = legend_label,
       values = color
     )
     ncolor <- length(unique(color))
-    if(ncolor==1) {
+    if(ncolor==1 || plot_type %in% c("violin", "boxplot", "errorplot")) {
       plot <- plot + guides(color = "none")
     }
 
@@ -152,7 +152,7 @@ update_legend_aes <- function(plot, plot_type, color, fill, shape, linetype, leg
         name = legend_title, labels = legend_label,
         values = color
       )
-      if(ncolor==1) {
+      if(ncolor==1 || plot_type %in% c("violin", "boxplot", "errorplot")) {
         plot <- plot + guides(fill = "none")
       }
     }
@@ -165,7 +165,7 @@ update_legend_aes <- function(plot, plot_type, color, fill, shape, linetype, leg
       values = fill
     )
     nfill <- length(unique(fill))
-    if(nfill==1) {
+    if(nfill==1 || plot_type %in% c("violin", "boxplot", "errorplot")) {
       plot <- plot + guides(fill = "none")
     }
     # if fill is defined and color isn't, update both using fill values
@@ -174,7 +174,7 @@ update_legend_aes <- function(plot, plot_type, color, fill, shape, linetype, leg
         name = legend_title, labels = legend_label,
         values = fill
       )
-      if(nfill==1) {
+      if(nfill==1 || plot_type %in% c("violin", "boxplot", "errorplot")) {
         plot <- plot + guides(color = "none")
       }
     }
@@ -182,13 +182,19 @@ update_legend_aes <- function(plot, plot_type, color, fill, shape, linetype, leg
 
   # if both color and fill values are not defined, use default values
   if (is.null(color) & is.null(fill)) {
-    # removing default coloring by groups for violin and boxplots
+
+    # removing default coloring by groups for violin, boxplots and errorplot
     if(plot_type %in% c("violin","boxplot")){
       plot <- plot +
         scale_fill_manual(name = legend_title, labels = legend_label, values = rep("white", length(legend_label))) +
[…]

### 2022-05-15 R/utils.R
modify legend settings

@@ -184,17 +184,17 @@ update_legend_aes <- function(plot, plot_type, color, fill, shape, linetype, leg
   if (is.null(color) & is.null(fill)) {
 
     # removing default coloring by groups for violin, boxplots and errorplot
-    if(plot_type %in% c("violin","boxplot")){
+    if(plot_type %in% c("violin","boxplot")) {
       plot <- plot +
         scale_fill_manual(name = legend_title, labels = legend_label, values = rep("white", length(legend_label))) +
         scale_color_manual(name = legend_title, labels = legend_label, values = rep("black", length(legend_label))) +
         guides(fill = "none", color = "none")
-    } else if (plot_type %in% c("errorplot") || nvar == 1){
+    } else if (plot_type %in% c("errorplot") || nvar == 1) {
       plot <- plot +
         scale_fill_manual(name = legend_title, labels = legend_label, values = rep("black", length(legend_label))) +
         scale_color_manual(name = legend_title, labels = legend_label, values = rep("black", length(legend_label))) +
         guides(fill = "none", color = "none")
-    }else{
+    }else {
       plot <- plot +
         scale_fill_discrete(name = legend_title, labels = legend_label) +
         scale_color_discrete(name = legend_title, labels = legend_label)
@@ -207,7 +207,7 @@ update_legend_aes <- function(plot, plot_type, color, fill, shape, linetype, leg
       name = legend_title, labels = legend_label,
       values = shape
     )
-  } else{
+  } else {
     plot <- plot + scale_shape_discrete(name = legend_title, labels = legend_label)
   }
 
@@ -218,7 +218,7 @@ update_legend_aes <- function(plot, plot_type, color, fill, shape, linetype, leg
     plot <- plot + scale_linetype_discrete(name = legend_title, labels = legend_label)
   }
 
-  if(nvar == 1){
+  if(nvar == 1) {
     plot <- plot + guides(fill = "none", color = "none", shape = "none", linetype = "none")
   }
 

### 2022-05-15 R/utils.R
Fixing two CRAN checks issues [stringasFactors and errors for match.arg], and dotplot now removes color and legend

@@ -142,7 +142,7 @@ update_legend_aes <- function(plot, plot_type, color, fill, shape, linetype, leg
       values = color
     )
     ncolor <- length(unique(color))
-    if(ncolor==1 || plot_type %in% c("violin", "boxplot", "errorplot")) {
+    if(ncolor==1 || plot_type %in% c("violin", "boxplot", "errorplot", "dotplot")) {
       plot <- plot + guides(color = "none")
     }
 
@@ -152,7 +152,7 @@ update_legend_aes <- function(plot, plot_type, color, fill, shape, linetype, leg
         name = legend_title, labels = legend_label,
         values = color
       )
-      if(ncolor==1 || plot_type %in% c("violin", "boxplot", "errorplot")) {
+      if(ncolor==1 || plot_type %in% c("violin", "boxplot", "errorplot", "dotplot")) {
         plot <- plot + guides(fill = "none")
       }
     }
@@ -165,7 +165,7 @@ update_legend_aes <- function(plot, plot_type, color, fill, shape, linetype, leg
       values = fill
     )
     nfill <- length(unique(fill))
-    if(nfill==1 || plot_type %in% c("violin", "boxplot", "errorplot")) {
+    if(nfill==1 || plot_type %in% c("violin", "boxplot", "errorplot", "dotplot")) {
       plot <- plot + guides(fill = "none")
     }
     # if fill is defined and color isn't, update both using fill values
@@ -174,7 +174,7 @@ update_legend_aes <- function(plot, plot_type, color, fill, shape, linetype, leg
         name = legend_title, labels = legend_label,
         values = fill
       )
-      if(nfill==1 || plot_type %in% c("violin", "boxplot", "errorplot")) {
+      if(nfill==1 || plot_type %in% c("violin", "boxplot", "errorplot", "dotplot")) {
         plot <- plot + guides(color = "none")
       }
     }
@@ -189,7 +189,7 @@ update_legend_aes <- function(plot, plot_type, color, fill, shape, linetype, leg
         scale_fill_manual(name = legend_title, labels = legend_label, values = rep("white", length(legend_label))) +
         scale_color_manual(name = legend_title, labels = legend_label, values = rep("black", length(legend_label))) +
         guides(fill = "none", color = "none")
-    } else if (plot_type %in% c("errorplot") || nvar == 1) {
+    } else if (plot_type %in% c("errorplot","dotplot") || nvar == 1) {
       plot <- plot +
[…]

### 2022-05-15 tests/testthat/test_ggmatplot_parameters.R

@@ -161,7 +161,7 @@ test_that("invalid plot types are not allowed", {
   skip_on_os("windows")
   expect_error(
     ggmatplot(x, y, plot_type = "scatterplot"),
-    "'arg' should be one of “point”, “line”, “both”, “density”, “histogram”, “boxplot”, “dotplot”, “errorplot”, “violin”, “ecdf”"
+    "'arg' should be one of \"point\", \"line\", \"both\", \"density\", \"histogram\", \"boxplot\", \"dotplot\", \"errorplot\", \"violin\", \"ecdf\""
   )
 })
 

### 2022-05-15 tests/testthat/test_pivotlonger.R

@@ -30,9 +30,10 @@ test_that("NA values are preserved in both id columns and pivoted columns", {
 })
 
 test_that("id columns can have different data types", {
-  df <- data.frame(x = 1:3, y = c("a", "b", "c"), z = 31:33)
+  df <- data.frame(x = 1:3, y = c("a", "b", "c"), z = 31:33, stringsAsFactors = FALSE)
   pv <- pivotlonger(df,
-    names_to = "Group", values_to = "Value",
+    names_to = "Group",
+    values_to = "Value",
     id_cols = c(1, 2)
   )
 

### 2022-05-15 README.Rmd
Test plot
modify readme

@@
 ```{r eval=FALSE}
-# install.packages("remotes")
-remotes::install_github("xuan-liang/ggmatplot")
+# install.packages("devtools")
+devtools::install_github("xuan-liang/ggmatplot")
 ```
@@
 
-If two matrices with equal number of columns are used, the corresponding columns of the matrices will be plotted against each other.
-i.e.
-column 1 of matrix `x` will be plotted against column 1 of matrix `y`, column 2 of matrix `x` will be plotted against column 2 of matrix `y`, etc.
+If two matrices with equal number of columns are used, the corresponding columns of the matrices will be plotted against each other, i.e.
+column 1 of matrix `y` will be plotted against column 1 of matrix `x`, column 2 of matrix `y` will be plotted against column 2 of matrix `x`, etc.
 
 The next example uses the iris dataset, with matrices `x` and `y` as shown below.
-The `Sepal.Length` is plotted against `Sepal.Width` and the `Petal.Length` is plotted against `Petal.Width`.
+The `Sepal.Width` is plotted against `Sepal.Length` and the `Petal.Width` is plotted against `Petal.Length`.
 Therefore the groups 'Column 1' and 'Column 2' can be interpreted as 'Sepal' and 'Petal' respectively.
-To make the plot more meaningful, we can further add the legend label and axis names by `legend_label`, `xlab` and `ylab`.
+To make the plot more meaningful, we can further modify the legend label and axis names by `legend_label`, `xlab` and `ylab`.
 
@@
 
-Next is plot of the US personal expenditure over 5 categories and 5 years, and is a simple example of how wide format data can be used with `ggmatplot()`.
+Next is the plot of the US personal expenditure over 5 categories and 5 years, and is a simple example of how wide format data can be used with `ggmatplot()`.
 Note how the expenditure categories to be used on the x axis is used as vector `x`, and the expenditure values is used in wide format as matrix `y` - with its columns corresponding to the grouping structure.
 
-The plot specifies the plot type as `plot_type = "both"`, which is a combination of 'point' and 'line' plots. It customized using `ggmatplot()` parameters and a `ggplot` theme as well.
+The plot specifies the plot type as `plot_type = "both"`, which is a combination of 'point' and 'line' plots. It is further customized by using `ggmatplot()` parameters and a `ggplot` theme as well.
 
@@
 Boxplots accept only a single matrix or data frame as well, and uses its columns as individual groups.
-While `ggmatplot` plots are filled by default, the fill color can be made transparent by using `alpha = 0`.
+By default, the fill color is white.
+But it is easy to customize and the transparency can be modified by the `alpha` parameter.
 
@@
   plot_type = "boxplot",
-  alpha = 0, # removing fill values
+  color = 'black', 
+  fill = 'grey',
+  alpha = 0.8,
   xlab = "", ylab = ""
@@
 
-Violin plots too accepts a single matrix or data frame input, and behaves similar to density plots and boxplots.
+Violin plots accept a single matrix or data frame input, and behave similar to density plots and boxplots.
 
@@
 
-Dotplots too accept a single matrix input and plots the distribution of each of its columns.
-
-The next example uses the `plot_type = "dotplot"` to visualize the distribution of the data, and also customizes the position of the legend using a `ggplot` theme as well.
+Dotplots too accept a single matrix input and plot the distribution of each of its columns.
+The next example uses the `plot_type = "dotplot"` to visualize the distribution of the data with the custom color and binwidth value.
+Note that the default setting for binwidth is 1/30 of the range of the data.
 
@@
   color = c("#00AFBB", "#E7B800"),
+  binwidth = 0.1,
   xlab = "", ylab = ""
-) +
-  theme(legend.position = "bottom")
+)
 ```
 
-Similar to density, violin, dotplots, and box plots, histograms too accept a single matrix or data frame input and groups the plot using its columns.
+Similar to density, violin, dotplots, and box plots, histograms too accept a single matrix or data frame input and group the plot using its columns.
 The histogram in the following example uses a matrix of 4 columns, and therefore groups the plots based on these 4 columns.
-The plot is also faceted by group.
+The plot is also faceted by group and the legend is removed by a `ggplot` theme setting.
 
@@
   plot_type = "histogram",
-  xlab = "Group",
+  xlab = "",
   color = "black",
@@
 ) +
-  facet_wrap(~Group, scales = "free")
+  facet_wrap(~Group, scales = "free") + 
+  theme(legend.position = 'none')
 ```
@@
   plot_type = "ecdf",
-  xlab = "Group",
+  xlab = "",
+  ylab = 'Empirical CDF',
   size = 1
@@
 
-The next example construct out an `errorplot` using the medians and interquartile ranges of each variable.
+The next example, plots out an `errorplot` using the medians and interquartile ranges of each variable.
 
@@
   desc_stat = "median_iqr",
-  xlab = "Group",
+  xlab = "",
   size = 1

### 2022-05-15 Paper/paper.Rmd
update paper with package and vignette source

@@
 The package is inspired by the function `matplot()` in the core R `graphics` system -- as such, `ggmatplot` may be considered as a `ggplot` version of `matplot`, with the benefits of customising the plots as any other `ggplot` objects via `ggplot2` functions, as well as offering several other plotting types that are not immediately available from `matplot` directly e.g., comparative violin plots.
+The `ggmatplot` package is available on the Comprehensive R Archive Network (CRAN) with the latest developmental source code available at \url{https://github.com/xuan-liang/ggmatplot}.
 
@@
 Future developments of the package could benefit from using a grammar approach, like in @Wilkinson2005-oz and @Wickham2010-kt, where plot types can be extensible.
+The latest developmental source code can be found at  Further examples can be found at \url{https://xuan-liang.github.io/ggmatplot/}.
 

### 2022-05-15 Paper/paper.tex
update paper with package and vignette source

@@
 types that are not immediately available from \texttt{matplot} directly
-e.g., comparative violin plots.
+e.g., comparative violin plots. The \texttt{ggmatplot} package is
+available on the Comprehensive R Archive Network (CRAN) with the latest
+developmental source code available at
+\url{https://github.com/xuan-liang/ggmatplot}.
 
@@
 approach, like in Wilkinson (2005) and Wickham (2010), where plot types
-can be extensible.
+can be extensible. The latest developmental source code can be found at
+Further examples can be found at
+\url{https://xuan-liang.github.io/ggmatplot/}.
 

### 2022-05-15 tests/testthat/test_errorplotstats.R
update tests to skip ci and covr

@@ -22,6 +22,8 @@ test_that("format of resultant data frame is as expected", {
 
 test_that("column means are calculated as expected", {
   skip_on_cran()
+  skip_on_ci()
+  skip_on_covr()
   skip_on_os("windows")
   suppressWarnings(expect_setequal(
     round(errorstats_mean_se$y,3),
@@ -31,6 +33,8 @@ test_that("column means are calculated as expected", {
 
 test_that("column medians are calculated as expected", {
   skip_on_cran()
+  skip_on_ci()
+  skip_on_covr()
   skip_on_os("windows")
   suppressWarnings(expect_setequal(
     round(errorstats_median_iqr$y,3),

### 2022-05-15 tests/testthat/test_ggmatplot_baseplots.R

@@ -5,6 +5,8 @@ y <- iris[, 3:4]
 
 test_that("scatterplot draws correctly", {
   skip_on_cran()
+  skip_on_ci()
+  skip_on_covr()
   skip_on_os("windows")
   expect_doppelganger(
     "basic scatterplot",
@@ -14,6 +16,8 @@ test_that("scatterplot draws correctly", {
 
 test_that("line plot draws correctly", {
   skip_on_cran()
+  skip_on_ci()
+  skip_on_covr()
   skip_on_os("windows")
   expect_doppelganger(
     "basic line plot",
@@ -23,6 +27,8 @@ test_that("line plot draws correctly", {
 
 test_that("point + line plot draws correctly", {
   skip_on_cran()
+  skip_on_ci()
+  skip_on_covr()
   skip_on_os("windows")
   expect_doppelganger(
     "basic point + line plot",
@@ -32,6 +38,8 @@ test_that("point + line plot draws correctly", {
 
 test_that("density plot draws correctly", {
   skip_on_cran()
+  skip_on_ci()
+  skip_on_covr()
   skip_on_os("windows")
   expect_doppelganger(
     "basic density plot",
@@ -41,6 +49,8 @@ test_that("density plot draws correctly", {
 
 test_that("histogram draws correctly", {
   skip_on_cran()
+  skip_on_ci()
+  skip_on_covr()
   skip_on_os("windows")
   expect_doppelganger(
     "basic histogram",
@@ -50,6 +60,8 @@ test_that("histogram draws correctly", {
 
 test_that("violin plot draws correctly", {
   skip_on_cran()
+  skip_on_ci()
+  skip_on_covr()
   skip_on_os("windows")
   expect_doppelganger(
     "basic violin plot",
@@ -59,6 +71,8 @@ test_that("violin plot draws correctly", {
 
 test_that("boxplot draws correctly", {
   skip_on_cran()
+  skip_on_ci()
+  skip_on_covr()
   skip_on_os("windows")
   expect_doppelganger(
     "basic boxplot",
@@ -68,6 +82,8 @@ test_that("boxplot draws correctly", {
 
 test_that("dotplot draws correctly", {
   skip_on_cran()
+  skip_on_ci()
+  skip_on_covr()
   skip_on_os("windows")
   expect_doppelganger(
     "basic dotplot",
@@ -77,6 +93,8 @@ test_that("dotplot draws correctly", {
 
[…]

### 2022-05-15 tests/testthat/test_ggmatplot_parameters.R

@@ -5,6 +5,8 @@ y <- iris[, 2:4]
 
 test_that("plot colors by color values", {
   skip_on_cran()
+  skip_on_ci()
+  skip_on_covr()
   skip_on_os("windows")
   # single color value
   expect_doppelganger(
@@ -20,6 +22,8 @@ test_that("plot colors by color values", {
 
 test_that("plot colors by fill values", {
   skip_on_cran()
+  skip_on_ci()
+  skip_on_covr()
   skip_on_os("windows")
   # single fill value
   expect_doppelganger(
@@ -35,6 +39,8 @@ test_that("plot colors by fill values", {
 
 test_that("plot colors by color and fill values simultaneously", {
   skip_on_cran()
+  skip_on_ci()
+  skip_on_covr()
   skip_on_os("windows")
   # single color value, fill values = number of groups
   expect_doppelganger(
@@ -48,6 +54,8 @@ test_that("plot colors by color and fill values simultaneously", {
 
 test_that("point shapes updated based on shape values", {
   skip_on_cran()
+  skip_on_ci()
+  skip_on_covr()
   skip_on_os("windows")
   # single shape value
   expect_doppelganger(
@@ -63,6 +71,8 @@ test_that("point shapes updated based on shape values", {
 
 test_that("line types updated based on linetype values", {
   skip_on_cran()
+  skip_on_ci()
+  skip_on_covr()
   skip_on_os("windows")
   # single linetype value
   expect_doppelganger(
@@ -88,6 +98,8 @@ test_that("line types updated based on linetype values", {
 
 test_that("plot axis limits are updated based on xlim and ylim parameters", {
   skip_on_cran()
+  skip_on_ci()
+  skip_on_covr()
   skip_on_os("windows")
   # x axis limits only
   suppressWarnings(expect_doppelganger(
@@ -109,6 +121,8 @@ test_that("plot axis limits are updated based on xlim and ylim parameters", {
 test_that("plot axes are transformed to a log scale based on log parameter
           values", {
   skip_on_cran()
+  skip_on_ci()
+  skip_on_covr()
   skip_on_os("windows")
   # log(x)
   expect_doppelganger(
@@ -129,6 +143,8 @@ test_that("plot axes are transformed to a log scale based on log parameter
 
 test_that("plot title is updated based on main paramter", {
   skip_on_cran()
+  skip_on_ci()
+  skip_on_covr()
[…]

### 2022-05-17 Paper/paper.Rmd
minor modification

@@
 Future developments of the package could benefit from using a grammar approach, like in @Wilkinson2005-oz and @Wickham2010-kt, where plot types can be extensible.
-The latest developmental source code can be found at  Further examples can be found at \url{https://xuan-liang.github.io/ggmatplot/}.
+The latest developmental source code can be found at \url{https://github.com/xuan-liang/ggmatplot}.
+Further examples can be found at \url{https://xuan-liang.github.io/ggmatplot/}.
 

### 2022-05-17 Paper/paper.tex
minor modification

--- 2022-05-09 Paper/paper.tex
@@
 \renewcommand{\headrulewidth}{0pt}
-\fancyhead[L]{\hspace{-0.75cm}\includegraphics[width=5.5cm]{/Library/Frameworks/R.framework/Versions/4.1/Resources/library/rticles/rmarkdown/templates/joss/resources/JOSS-logo.png}}
+\fancyhead[L]{\hspace{-0.75cm}\includegraphics[width=5.5cm]{/Library/Frameworks/R.framework/Versions/4.2/Resources/library/rticles/rmarkdown/templates/joss/resources/JOSS-logo.png}}
 \fancyhead[C]{}
@@
 approach, like in Wilkinson (2005) and Wickham (2010), where plot types
-can be extensible.
+can be extensible. The latest developmental source code can be found at
+\url{https://github.com/xuan-liang/ggmatplot}. Further examples can be
+found at \url{https://xuan-liang.github.io/ggmatplot/}.
 

### 2022-05-17 Paper/paper.Rmd
update ggplot2 codes

@@
                values_to = "rating") %>% 
+  mutate(rating_type = fct_inorder(rating_type))%>%
   ggplot(aes(x = overall_rating, y = rating, color = rating_type)) + 
@@
   ggplot(aes(Month, Snowfall)) + 
-  geom_boxplot() + 
+  geom_boxplot(alpha = 0.5) + 
   ggtitle("Grand Rapids, Michigan, 1893-2011")

### 2022-05-17 Paper/paper.tex
update ggplot2 codes

--- 2022-05-15 Paper/paper.tex
@@
                \AttributeTok{values\_to =} \StringTok{"rating"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{rating\_type =} \FunctionTok{fct\_inorder}\NormalTok{(rating\_type))}\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ overall\_rating, }\AttributeTok{y =}\NormalTok{ rating, }\AttributeTok{color =}\NormalTok{ rating\_type)) }\SpecialCharTok{+}
@@
   \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(Month, Snowfall)) }\SpecialCharTok{+} 
-  \FunctionTok{geom\_boxplot}\NormalTok{() }\SpecialCharTok{+} 
+  \FunctionTok{geom\_boxplot}\NormalTok{(}\AttributeTok{alpha =} \FloatTok{0.5}\NormalTok{) }\SpecialCharTok{+} 
   \FunctionTok{ggtitle}\NormalTok{(}\StringTok{"Grand Rapids, Michigan, 1893{-}2011"}\NormalTok{)}
@@
 can be extensible. The latest developmental source code can be found at
-Further examples can be found at
-\url{https://xuan-liang.github.io/ggmatplot/}.
+\url{https://github.com/xuan-liang/ggmatplot}. Further examples can be
+found at \url{https://xuan-liang.github.io/ggmatplot/}.
 

### 2022-05-17 Paper/paper.Rmd
update plots

@@
                values_to = "rating") %>% 
-  mutate(rating_type = fct_inorder(rating_type))%>%
+  mutate(rating_type = fct_inorder(rating_type)) %>%
   ggplot(aes(x = overall_rating, y = rating, color = rating_type)) + 

### 2022-05-17 Paper/paper.tex
update plots

@@
                \AttributeTok{values\_to =} \StringTok{"rating"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{rating\_type =} \FunctionTok{fct\_inorder}\NormalTok{(rating\_type))}\SpecialCharTok{\%\textgreater{}\%}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{rating\_type =} \FunctionTok{fct\_inorder}\NormalTok{(rating\_type)) }\SpecialCharTok{\%\textgreater{}\%}
   \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ overall\_rating, }\AttributeTok{y =}\NormalTok{ rating, }\AttributeTok{color =}\NormalTok{ rating\_type)) }\SpecialCharTok{+}

### 2022-06-03 README.Rmd
Add DOI and update websites

@@
 ```
+
+[![DOI](https://zenodo.org/badge/195039918.svg)](https://zenodo.org/badge/latestdoi/195039918)
+
 # ggmatplot

### 2024-05-01 R/ggmatplot.R
change aes_string to aes (aes_string is deprecated)

@@ -125,61 +125,61 @@ ggmatplot <- function(x = NULL, y = NULL,
     do.call(geom, c(list(mapping = aes), params))
   }
 
-  p <- ggplot(data = data, mapping = aes_string(group = GROUP_NAME))
+  p <- ggplot(data = data, mapping = aes(group = .data[[GROUP_NAME]]))
   p <- p + switch(plot_type,
                   "point" = plot_geom(geom_point,
-                                      aes_string(x = xname,
-                                                 y = yname,
-                                                 shape = GROUP_NAME,
-                                                 color = GROUP_NAME)),
+                                      aes(x = .data[[xname]],
+                                          y = .data[[yname]],
+                                          shape = .data[[GROUP_NAME]],
+                                          color = .data[[GROUP_NAME]])),
                   "line" = plot_geom(geom_line,
-                                     aes_string(x = xname,
-                                                y = yname,
-                                                linetype = GROUP_NAME,
-                                                color = GROUP_NAME)),
+                                     aes(x = .data[[xname]],
+                                         y = .data[[yname]],
+                                         linetype = .data[[GROUP_NAME]],
+                                         color = .data[[GROUP_NAME]])),
                   "both" = list(plot_geom(geom_point,
-                                          aes_string(x = xname,
-                                                     y = yname,
-                                                     shape = GROUP_NAME,
-                                                     color = GROUP_NAME)),
+                                          aes(x = .data[[xname]],
+                                              y = .data[[yname]],
+                                              shape = .data[[GROUP_NAME]],
+                                              color = .data[[GROUP_NAME]])),
                                 plot_geom(geom_line,
-                                          aes_string(x = xname,
-                                                     y = yname,
-                                                     linetype = GROUP_NAME,
-                                                     color = GROUP_NAME))),
+                                          aes(x = .data[[xname]],
+                                              y = .data[[yname]],
+                                              linetype = .data[[GROUP_NAME]],
+                                              color = .data[[GROUP_NAME]]))),
                   "density" = plot_geom(geom_density,
-                                        aes_string(x = yname,
-                                                   color = GROUP_NAME,
-                                                   fill = GROUP_NAME)),
+                                        aes(x = .data[[yname]],
+                                            color = .data[[GROUP_NAME]],
+                                            fill = .data[[GROUP_NAME]])),
                   "histogram" = plot_geom(geom_histogram,
-                                          aes_string(x = yname,
-                                                     color = GROUP_NAME,
[…]

### 2024-05-01 R/pivotlonger.R
make pivotlonger more efficient

@@ -22,15 +22,16 @@ pivotlonger <- function(data, names_to = "", values_to = "", id_cols = NULL) {
   n_widecols <- ncol(data) - length(id_cols)
   ifelse(!is.null(id_cols), widecols <- data[-1 * id_cols], widecols <- data)
   for (i in 0:nrow(data) - 1) {
-    for (j in 1:n_widecols) {
-      if (!is.null(id_cols)) {
+    if (!is.null(id_cols)) {
+      for (j in 1:n_widecols) {
         for (k in 1:length(id_cols)) {
           long_df[n_widecols * i + j, k] <- data[i + 1, id_cols[k]]
         }
       }
-      long_df[n_widecols * i + j, names_to] <- colnames(widecols)[j]
-      long_df[n_widecols * i + j, values_to] <- widecols[i + 1, j]
     }
+    #browser()
+    long_df[n_widecols * i + 1:n_widecols, names_to] <- colnames(widecols)
+    long_df[n_widecols * i + 1:n_widecols, values_to] <- do.call(c, as.list(widecols[i + 1, ]))
   }
   return(long_df)
 }

### 2024-05-01 R/ggmatplot.R
fixed to preserve class when using pivotlonger

@@ -104,7 +104,6 @@ ggmatplot <- function(x = NULL, y = NULL,
                       ylim = c(NA, NA), log = NULL, main = NULL, xlab = NULL,
                       ylab = NULL, legend_label = NULL, legend_title = NULL,
                       desc_stat = "mean_se", asp = NA, ...) {
-
   # binding global variables to objects
   Variable <- ymin <- ymax <- NULL
   params_input <- list(...)

### 2024-05-01 R/pivotlonger.R

@@ -16,7 +16,7 @@
 #' pivotlonger(data = iris_sub, names_to = "Measurement", values_to = "Value",
 #' id_cols = c(1, 2))
 #'
-pivotlonger <- function(data, names_to = "", values_to = "", id_cols = NULL) {
+pivotlonger <- function(data, names_to = "name", values_to = "value", id_cols = NULL) {
   long_df <- data.frame(matrix(ncol = 2 + length(id_cols), nrow = 0))
   colnames(long_df) <- c(colnames(data[id_cols]), names_to, values_to)
   n_widecols <- ncol(data) - length(id_cols)
@@ -29,9 +29,10 @@ pivotlonger <- function(data, names_to = "", values_to = "", id_cols = NULL) {
         }
       }
     }
-    #browser()
+
     long_df[n_widecols * i + 1:n_widecols, names_to] <- colnames(widecols)
     long_df[n_widecols * i + 1:n_widecols, values_to] <- do.call(c, as.list(widecols[i + 1, ]))
   }
+  class(long_df[[values_to]]) <- class(widecols[[1]])
   return(long_df)
 }

### 2024-05-01 tests/testthat/test_pivotlonger.R

@@ -43,3 +43,10 @@ test_that("id columns can have different data types", {
   expect_equal(pv$Group, rep(names(df[, 3, drop = FALSE]), nrow(df)))
   expect_equal(pv$Value, c(31, 32, 33))
 })
+
+
+test_that("Expect date class to be preserved.", {
+  dates <- seq(as.Date("2023-01-01"), as.Date("2023-01-05"), by = "1 day")
+  pv <- pivotlonger(data.frame(date = dates))
+  expect_equal(pv$value, dates)
+})
